# Get Started {#sec-start}

This chapter provides installation instructions and introduces the datasets used in the tutorial.

```{r clear-environment, echo = FALSE}
rm(list = ls())
```

## Installation

To install **fect** from CRAN, run the code chunk below:

```{r install-cran, eval = FALSE, message = FALSE, warning = FALSE}
install.packages("fect")
```

We recommend users to install the most up-to-date, stable version of **fect** from Github using:

```{r install-github-main, eval = FALSE, message = FALSE, warning = FALSE, cache = FALSE,}
devtools::install_github("xuyiqing/fect")
```

We fix bugs in the `dev` branch before merging into the main branch; therefore, it is often more up to date.

```{r install-github-dev, eval = FALSE, message = FALSE, warning = FALSE, cache = FALSE,}
devtools::install_github("xuyiqing/fect@dev")
```

After installation, check **fect** version to make sure the package is up-to-date.

```{r check-version}
installed.packages()["fect", "Version"]
```

**panelView** for panel data visualization is highly recommended and will be used in the tutorial:

```{r install-panelview, eval=FALSE}
devtools::install_github('xuyiqing/panelView')
```

**fect** depends on the following packages, which should be installed automatically when **fect** is being installed. You can also install them manually.

```{r install-dependencies, eval=FALSE}
install_all <- function(packages) {
  installed_pkgs <- installed.packages()[, "Package"]
  for (pkg in packages) {
    if (!pkg %in% installed_pkgs) {
      install.packages(pkg)
    }
  }
}
packages <- c("abind", "doParallel", "doRNG", "fixest", "foreach", "future", 
              "GGally", "ggplot2", "grid", "gridExtra", "MASS", 
              "panelView", "Rcpp")
install_all(packages)
```

## Datasets

The **fect** package ships several datasets. All simulated and empirical datasets are loaded together with `data(fect)`.

```{r load-data, message = FALSE, warning = FALSE}
library(fect)
data(fect)
ls()
```

### Simulated datasets

The package includes two simulated panel datasets generated from the data-generating process (DGP) in @LWX2022. Both have $N = 200$ units and $T = 35$ time periods. Treatment switches on and off over time (99 of 150 treated units experience at least one reversal), reflecting a general treatment pattern rather than simple staggered adoption.

The full DGP for `simdata` is:
$$Y_{it} = \mu + 3\alpha_i + \xi_t + X_{1,it} + 3 X_{2,it} + \lambda_i' f_t + \tau_{it} D_{it} + \varepsilon_{it}$$
where $\alpha_i \sim N(0,1)$ are unit fixed effects, $\xi_t$ follows an AR(1) process with drift (time fixed effects), $X_{1,it}$ and $X_{2,it} \sim N(0,1)$ are observed covariates with coefficients 1 and 3, $\lambda_i \in \mathbb{R}^2$ are unit-specific factor loadings drawn from $N(0.5, 1)$, $f_t \in \mathbb{R}^2$ are latent time factors (one trending, one white noise), and $\varepsilon_{it} \sim N(0,2)$. The treatment effect is heterogeneous: $\tau_{it} \sim N(0.4 \cdot \text{tr\_cum}_{it}/T,\; 0.2)$, where $\text{tr\_cum}_{it}$ counts cumulative treatment periods. Treatment assignment depends on the latent factors and fixed effects, creating confounding that violates parallel trends. The grand mean is $\mu = 5$.

-   `simdata`: The main simulated dataset. The outcome includes two latent factors ($r = 2$), so the parallel trends assumption is violated and the FE estimator is biased. Used in [Chapter @sec-ife-mc] and [Chapter @sec-cfe] to demonstrate factor-augmented approaches.
-   `simdata1`: A simplified version of `simdata` in which the latent factor contributions ($\lambda_i' f_t$) are removed from the outcome. The parallel trends assumption holds, and the FE estimator is consistent. Used in [Chapter @sec-fect] to demonstrate the imputation estimator. Treatment assignment, covariates, fixed effects, and errors are identical to `simdata`.

The scripts that generate these datasets from the original @LWX2022 DGP code are in `data-raw/`.

### Empirical datasets

-   `simgsynth` & `turnout`: Based on @Xu2017. Used in [Chapter @sec-gsynth] to demonstrate the never-treated estimation regime.
-   `gs2020`: Based on @GS2020, who examine the effect of minority candidate presence on the proportion of coethnic donations in U.S. House elections. Used in [Chapter @sec-plots] and [Chapter @sec-panel].
-   `hh2019`: Based on @HH2019, who study the effect of indirect versus direct democracy on naturalization rates in Switzerland. Used in [Chapter @sec-plots], [Chapter @sec-panel], and [Chapter @sec-panel-sens].
