# Get Started {#sec-start}

This chapter provides installation instructions and introduces the datasets used in the tutorial.

```{r echo = FALSE}
rm(list = ls())
```

## Installation

To install **fect** from CRAN, run the code chunk below:

```{r eval = FALSE, message = FALSE, warning = FALSE}
install.packages("fect")
```

We recommend users to install the most up-to-date version of **fect** from Github using:

```{r eval = FALSE, message = FALSE, warning = FALSE, cache = FALSE,}
devtools::install_github("xuyiqing/fect")
```

After installation, check **fect** version to make sure the package is up-to-date.

```{r}
installed.packages()["fect", "Version"]
```

**panelView** for panel data visualization is highly recommended and will be used in the tutorial:

```{r eval=FALSE}
devtools::install_github('xuyiqing/panelView')
```

**fect** depends on the following packages, which should be installed automatically when **fect** is being installed. You can also install them manually.

```{r eval=FALSE}
install_all <- function(packages) {
  installed_pkgs <- installed.packages()[, "Package"]
  for (pkg in packages) {
    if (!pkg %in% installed_pkgs) {
      install.packages(pkg)
    }
  }
}
packages <- c("abind", "doParallel", "doRNG", "fixest", "foreach", "future", 
              "GGally", "ggplot2", "grid", "gridExtra", "Mass", 
              "panelView", "Rcpp")
install_all(packages)
```

## Datasets

The **fect** package ships five datasets.

```{r message = FALSE, warning = FALSE}
library(fect)
data(fect)
ls()
```

Below is a brief explanation of the datasets used in this book:

-   `simdata`: Based on @LWX2022. Used in [Chapter @sec-fect] to demonstrate various counterfactual estimators.\
-   `gs2020`: Based on @GS2020, who examine the effect of minority candidate presence on the proportion of coethnic donations in U.S. House elections. Used in [Chapter @sec-plots] and [Chapter @sec-panel].\
-   `simgsynth` & `turnout`: Based on @Xu2017. Used in [Chapter @sec-gsynth] to demonstrate Gsynth.
-   `hh2019`: Based on @HH2019, who study the effect of indirect versus direct democracy on naturalization rates in Switzerland. Used in [Chapter @sec-panel] and [Chapter @sec-panel-sens].
