# Get Started {#sec-start}

This chapter provides installation instructions and introduces the datasets used in the tutorial.

```{r clear-environment, echo = FALSE}
rm(list = ls())
```

## Installation

To install **fect** from CRAN, run the code chunk below:

```{r install-cran, eval = FALSE, message = FALSE, warning = FALSE}
install.packages("fect")
```

We recommend users to install the most up-to-date, stable version of **fect** from Github using:

```{r install-github-main, eval = FALSE, message = FALSE, warning = FALSE, cache = FALSE,}
devtools::install_github("xuyiqing/fect")
```

We fix bugs in the `dev` branch before merging into the main branch; therefore, it is often more up to date.

```{r install-github-dev, eval = FALSE, message = FALSE, warning = FALSE, cache = FALSE,}
devtools::install_github("xuyiqing/fect@dev")
```

After installation, check **fect** version to make sure the package is up-to-date.

```{r check-version}
installed.packages()["fect", "Version"]
```

**panelView** for panel data visualization is highly recommended and will be used in the tutorial:

```{r install-panelview, eval=FALSE}
devtools::install_github('xuyiqing/panelView')
```

**fect** depends on the following packages, which should be installed automatically when **fect** is being installed. You can also install them manually.

```{r install-dependencies, eval=FALSE}
install_all <- function(packages) {
  installed_pkgs <- installed.packages()[, "Package"]
  for (pkg in packages) {
    if (!pkg %in% installed_pkgs) {
      install.packages(pkg)
    }
  }
}
packages <- c("abind", "doParallel", "doRNG", "fixest", "foreach", "future", 
              "GGally", "ggplot2", "grid", "gridExtra", "MASS", 
              "panelView", "Rcpp")
install_all(packages)
```

## Datasets

The **fect** package ships several datasets. All simulated and empirical datasets are loaded together with `data(fect)`.

```{r load-data, message = FALSE, warning = FALSE}
library(fect)
data(fect)
ls()
```

### Simulated datasets

The package includes two simulated panel datasets with 200 units, 35 time periods, and treatment that switches on and off. Both share the same unit fixed effects, time fixed effects, covariates, treatment assignment, and error structure. They differ only in whether latent factors are present:

-   `simdata`: The main simulated dataset from @LWX2022. The outcome includes two latent factors ($r = 2$), so that the parallel trends assumption is violated and the FE estimator is biased. Used in [Chapter @sec-ife-mc] and [Chapter @sec-cfe] to demonstrate factor-augmented approaches.
-   `simdata1`: A simplified version of `simdata` in which the latent factors are removed. The parallel trends assumption holds, and the outcome follows a standard two-way fixed effects model ($Y_{it} = \alpha_i + \xi_t + \beta_1 X_{1,it} + \beta_2 X_{2,it} + \tau_{it} D_{it} + \epsilon_{it}$). Used in [Chapter @sec-fect] to demonstrate the imputation estimator.

### Empirical datasets

-   `simgsynth` & `turnout`: Based on @Xu2017. Used in [Chapter @sec-gsynth] to demonstrate the never-treated estimation regime.
-   `gs2020`: Based on @GS2020, who examine the effect of minority candidate presence on the proportion of coethnic donations in U.S. House elections. Used in [Chapter @sec-plots] and [Chapter @sec-panel].
-   `hh2019`: Based on @HH2019, who study the effect of indirect versus direct democracy on naturalization rates in Switzerland. Used in [Chapter @sec-plots], [Chapter @sec-panel], and [Chapter @sec-panel-sens].
