# The Imputation Estimator {#sec-fect}

In this chapter, we illustrate how to use the **fect** package to implement counterfactual estimators (or imputation estimators) and conduct diagnostic tests proposed by @LWX2024 [<a href="https://onlinelibrary.wiley.com/doi/full/10.1111/ajps.12723" target="_blank">Paper</a>]. R script used in this chapter can be downloaded [here](https://raw.githubusercontent.com/xuyiqing/fect/dev/vignettes/rscript/02-fect.R).

## Simulated data

In this chapter, we use `sim_base`, a simulated panel dataset in which the parallel trends assumption holds. The dataset has 200 units, 35 time periods, and treatment that switches on and off. The data are simulated with treatment switching on and off, capturing the general case of treatment reversal under strict exogeneity. The outcome is generated from a two-way fixed effects model with two covariates: $$Y_{it} = \tau_{it} D_{it} + \beta_1 X_{1,it} + \beta_2 X_{2,it} + \alpha_i + \xi_t  + \epsilon_{it}$$

Since there are no latent factors, the `"fe"` method (two-way fixed effects counterfactual estimator, or FEct) is correctly specified for this DGP. For settings where latent factors are present and the FE estimator is biased, see [Chapter @sec-ife-mc].

```{r .common, include = FALSE}
source("_common.R")
```

```{r setup-seed, echo = FALSE}
set.seed(1234)
```

```{r load-packages, message = FALSE, warning = FALSE}
data(sim_base)
data(sim_gsynth)
```

We use the **panelView** package to visualize the treatment and outcome variables:

```{r panelview-treatment, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 4.5}
library(panelView)
panelview(Y ~ D, data = sim_base, index = c("id","time"),
  axis.lab = "time", xlab = "Time", ylab = "Unit",
  gridOff = TRUE, by.timing = TRUE,
  background = "white", main = "Simulated Data: Treatment Status")
```

In the figure below, blue and gray represent treatment and control conditions.

```{r panelview-outcome, fig.width = 6, fig.height = 4.5, warning = FALSE}
panelview(Y ~ D, data = sim_base, index = c("id","time"),
  axis.lab = "time", xlab = "Time", ylab = "Unit",
  theme.bw = TRUE, type = "outcome", 
  main = "Simulated Data: Outcome")
```

------------------------------------------------------------------------

## The FEct estimator

The two-way fixed effects counterfactual (FEct) estimator imputes the counterfactual outcome $\hat{Y}(0)$ for treated observations using a two-way fixed effects model estimated on control observations. It is also independently proposed by @BJS2024 and @gardnertwo, who refer to it as the "imputation method" and "two-stage DID," respectively.

### Estimation

We estimate the average treatment effect on the treated (ATT) using the following information: the outcome variable $Y$, binary treatment variable $D$, two observed covariates $X_{1}$ and $X_{2}$, and the unit and time indicators $id$ and $time$, respectively.

The first variable on the right hand side of the formula is the treatment indicator $D$; the rest of the right-hand-side variables serve as controls. The `index` option specifies the unit and time indicators. The `force` option ("none", "unit", "time", and "two-way") specifies the additive component(s) of the fixed effects included in the model. The default option is "two-way" (including both unit and time fixed effects).

```{r simdata_fect_nose, eval=TRUE, cache = TRUE, message = FALSE, results = 'hide'}
out.fect <- fect(Y ~ D + X1 + X2, data = sim_base, index = c("id","time"),
                 method = "fe", force = "two-way")
```

::: {.callout-tip appearance="simple"}
### Model specification

The key parameters that control the **model** are: `method` (estimator choice), `force` (fixed effects structure), and `r` (number of factors for IFE/MC methods; see [Chapter @sec-ife-mc]; by default, `r` is set to 0 when `method = "fe"`). These determine how $\hat{Y}(0)$ is imputed.
:::

We can use the `plot` function to visualize the estimation results. By default, it produces a "gap" plot --- `plot(out.fect, type = "gap")` --- which shows the estimated period-wise ATT (dynamic treatment effects). The true population average effects in `sim_base` go from 1 to 3 from the 1st to the 10th post-treatment period.

```{r fect_plot_nose, fig.width = 6, fig.height = 4.5}
plot(out.fect, main = "Estimated ATT (FEct)", ylab = "Effect of D on Y",
  cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8)
```

In the gap plot, pre-treatment estimates appear in gray (in-sample) while post-treatment estimates appear in black (out-of-sample). This visual distinction highlights that pre-treatment "effects" should be near zero if the model is well-specified, while post-treatment effects are the quantities of interest. When `loo = TRUE` is used in the plot call, all points appear in black because both pre- and post-treatment estimates are out-of-sample. The uncertainty estimates are unavailable in the plot above because, by default, `se = FALSE` to save computational power. See [Chapter @sec-plots] for more visualization options.

### Inference

The package can produce uncertainty estimates when `se = TRUE`. The default is the non-parametric cluster-bootstrap (`vartype = "bootstrap"`), which works well when the number of units is relatively large and many units experience the treatment condition. The number of bootstrap runs can be set by `nboots`.

::: {.callout-tip appearance="simple"}
### Inference parameters

The key parameters that control **uncertainty estimation** are: `se` (enable standard errors), `vartype` (`"bootstrap"`, `"jackknife"`, or `"parametric"`), `nboots` (number of bootstrap replications), and `cl` (clustering variable). These parameters do not affect point estimates. `"parametric"` is only available in the Synth setting, i.e., when `time.component.from = "nevertreated"`.
:::

::: {.callout-note appearance="simple"}
**Parallel computing** will speed up both cross-validation and uncertainty estimation significantly. We recommend that users manually set the number of cores using the `cores` option. If this is not supplied or is `NULL`, we will automatically select the smaller of `8` and the number of usable system cores minus `2`, to prevent excessive use of system resources.

The `parallel` argument accepts five forms, allowing users to enable parallelism for cross-validation and the bootstrap independently:

| Value | CV | Bootstrap |
|:------|:---|:----------|
| `TRUE` (default) | Parallel (subject to auto-threshold) | Parallel |
| `FALSE` | Serial | Serial |
| `"cv"` | Parallel (overrides threshold) | Serial |
| `"boot"` | Serial | Parallel |
| `c("cv", "boot")` | Parallel | Parallel |

When `parallel = TRUE`, CV parallelism auto-engages only on panels above a method-specific size threshold (see [Chapter @sec-ife-mc] for the CV thresholds). The string forms `"cv"` and `"boot"` bypass the threshold — useful when the user knows their panel will benefit from parallelism even at smaller sizes.
:::

::: {.callout-note appearance="simple"}
**Clustering.**

-   By default, `fect()` uses cluster-bootstrap at the `unit` level when `se = TRUE`, that is `vartype = "bootstrap"`. The uncertainty estimates thus account for arbitrary serial correlation within a unit over time---commonly understood as being "clustered" at the unit level. In the example above, the unit is `id`, hence, by default, `cl = "id"`.
-   Alternatively, users can obtain uncertainty estimates using the cluster-jackknife method by specifying `vartype = "jackknife"`, also at the unit level. In this case, the algorithm calculates standard errors by iteratively dropping one unit (i.e., the entire time series) from the dataset. This can be particularly useful when the number of treated units is small.
-   To cluster standard errors at a different, usually higher, level, users can specify the clustering variable using the `cl` option.
:::

### Parametric bootstrap: valid regimes {#sec-parametric-regimes}

The parametric bootstrap (`vartype = "parametric"`) uses a two-stage pseudo-treated resampling procedure to produce uncertainty estimates. It is well-suited for the synthetic control setting, where the number of treated units is small and nonparametric bootstrap is unreliable. However, the procedure rests on two key assumptions: (a) the time components (factors and loadings) are estimated from a **never-treated** control pool that is independent of the treated units' data, and (b) the model residuals on control units accurately reflect the true error distribution. When either assumption is violated, the parametric bootstrap underestimates standard errors and produces confidence intervals that are too narrow.

To prevent users from invoking the parametric bootstrap in settings where these assumptions fail, the package enforces three hard restrictions:

::: {.callout-important appearance="simple"}
### Three-gate system

| Gate | Condition | What it blocks |
|---|---|---|
| A | `method %in% c("mc", "both")` + `vartype = "parametric"` | MC + parametric |
| B | treatment reversal present + `vartype = "parametric"` | reversal + parametric |
| C | `time.component.from = "notyettreated"` + `vartype = "parametric"` | notyettreated + parametric |

In all three cases the package raises an informative error pointing to the valid alternatives.
:::

The motivation for Gate C comes from both theory and evidence. The `notyettreated` path uses EM imputation to estimate factors jointly across all not-yet-treated observations, including the treated units' pre-period data. EM enlarges the effective factor space by imputing masked treated-post cells, which deflates the observed-cell residuals that the parametric bootstrap feeds into its covariance estimator. In a Monte Carlo study (N = 50, T = 20, r = 2 factors, 500 replications), the blocked `ife + notyettreated + parametric` combination produced **80.6% coverage at nominal 95%** (SE ratio ≈ 0.67). In contrast, all tested nevertreated paths produced near-nominal coverage (0.93–0.96, SE ratio ≈ 1.0). The architectural details of all three gates are documented in `ARCHITECTURE.md`.

If your current code uses `vartype = "parametric"` with the default `time.component.from = "notyettreated"`, migrate as follows:

```r
# Before (now raises an error under Gate C):
fect(data, Y = "Y", D = "D", index = c("id", "time"),
     method = "ife", se = TRUE, vartype = "parametric")

# After — option 1: switch to nevertreated imputation
# (requires that never-treated control units exist in the data)
fect(data, Y = "Y", D = "D", index = c("id", "time"),
     method = "ife", time.component.from = "nevertreated",
     se = TRUE, vartype = "parametric")

# After — option 2: use nonparametric bootstrap (safe default)
fect(data, Y = "Y", D = "D", index = c("id", "time"),
     method = "ife",
     se = TRUE, vartype = "bootstrap")
```

Option 1 switches to the gsynth-style estimation regime and keeps parametric bootstrap. This is appropriate when never-treated control units exist and the treatment does not reverse. Option 2 leaves the estimator unchanged and uses cluster-bootstrap, which is appropriate when the number of treated units is moderate to large. See [Chapter @sec-gsynth] for guidance.

```{r simdata_fect, eval=TRUE, cache = TRUE, message = FALSE, results = 'hide'}
out.fect <- fect(Y ~ D + X1 + X2, data = sim_base, index = c("id","time"),
  method = "fe", force = "two-way", se = TRUE,
  parallel = TRUE, cores = 16, nboots = 1000)
```

The `plot()` function can visualize the estimated period-wise ATTs as well as their uncertainty estimates. `stats = "F.p"` shows the p-value for the F test of no-pretrend.

```{r fect_plot_nse, fig.width = 6, fig.height = 4.5}
plot(out.fect, main = "Estimated ATT (FEct)", ylab = "Effect of D on Y",
  cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")
```

### Exiting the treatment

**fect** allows the treatment to switch back and forth and provides diagnostic tools for this setting. After the estimation, we can visualize the period-wise ATTs relative to the `exit` of treatments by setting `type = "exit"`. The x-axis is then realigned based on the timing of the treatment's exit, not onset, e.g., 1 represents 1 period after the treatment ends. In the exit plot, the color convention is reversed: pre-exit estimates appear in black (out-of-sample) and post-exit estimates appear in gray (in-sample).

```{r exit_fect, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.fect, type = "exit", main = "Exit Plot (FEct)")
```

### Save estimates

Users can use the `print` function to view a summary of the estimation results or retrieve relevant statistics by directly accessing the fect object. Specifically, `est.avg` and `est.avg.unit` show the ATT averaged over all periods -- the former weights each treated observation equally while the latter weights each treated unit equally. `beta` reports the coefficients of the time-varying covariates. `est.att` reports the average treatment effect on the treated (ATT) by period.

```{r print-fect}
print(out.fect)
```

To save space, results are not shown here.

```{r extract-estimates, eval = FALSE}
out.fect$est.att
out.fect$est.avg
out.fect$beta
```

After estimation with `se = TRUE`, bootstrap treatment effect estimates from each run are stored in `eff.boot`, an array whose dimension = (#time periods \* #treated \* #bootstrap runs). Standard errors for the period-wise ATTs are available in the output object alongside the point estimates.

```{r extract-bootstrap, eval = FALSE}
out.fect$eff.boot
```

For IFE and MC methods, see [Chapter @sec-ife-mc].

------------------------------------------------------------------------

## Diagnostics

Once we have point estimates and uncertainty estimates, we conduct diagnostic checks before interpreting results. This section covers the placebo test, the carryover test, and the leave-one-out approach.

### Placebo test

We provide a placebo test for a settled model by setting `placeboTest = TRUE`. We specify a range of pretreatment periods as "placebo periods" in option `placebo.period` to remove observations in the specified range for model fitting, and then test whether the estimated ATT in this range is significantly different from zero. Below, we set `c(-2, 0)` as the placebo periods.

```{r fect_placebo, eval=TRUE, cache=TRUE, message=FALSE, results='hide', fig.width=6, fig.height=4.5}
out.fect.placebo <- fect(Y ~ D + X1 + X2, data = sim_base, index = c("id","time"),
  force = "two-way", method = "fe",
  se = TRUE, nboots = 1000, parallel = TRUE, cores = 16,
  placeboTest = TRUE, placebo.period = c(-2, 0))
plot(out.fect.placebo, cex.text = 0.8)
```

For more detail on placebo tests, including IFE and MC examples, see [Chapter @sec-ife-mc].

### Carryover test

The idea of the placebo test can be extended to testing the presence of carryover effects. Instead of hiding a few periods right before the treatment starts, we hide a few periods right after the treatment ends. If carryover effects do not exist, we would expect the average prediction error in those periods to be close to zero.

To perform the carryover test, we set `carryoverTest = TRUE` and specify the range of exit-treatment periods in `carryover.period`. Below, we set `carryover.period = c(1, 3)`. Since `sim_base` is simulated without carryover effects, we expect the test to pass.

```{r fect_carryover, eval=TRUE, cache=TRUE, message=FALSE, results='hide'}
out.fect.carry <- fect(Y ~ D + X1 + X2, data = sim_base, index = c("id","time"),
  force = "two-way", method = "fe",
  se = TRUE, nboots = 1000, parallel = TRUE, cores = 16,
  carryoverTest = TRUE, carryover.period = c(1, 3))
```

```{r fect_carryover_plot, eval=TRUE, cache=TRUE, warning=FALSE, fig.width=6, fig.height=5}
plot(out.fect.carry, type = "exit", cex.text = 0.8, main = "Carryover Effects (FEct)")
```

For more detail on carryover tests with IFE and MC, see [Chapter @sec-ife-mc].

### Leave-one-out approach

@li2025benchmarking show that in some applications, pre-trend estimates based on in-sample model fit can lead to the mistaken belief that no pre-trend exists, even when a non-parallel pre-trend is present. A simple fix is to use a leave-one-out method by setting `loo = TRUE` to obtain these estimates, although it is significantly more time-consuming.

::: {.callout-note appearance="simple"}
We recommend setting `loo = TRUE` when (i) the event-study plot is intended as a critical piece of evidence to support the parallel trends assumption, which is often the case, or (ii) when implementing an equivalence test for the pre-trend estimates. For more discussion on the LOO pre-trend test, see [Chapter @sec-ife-mc].

Our most preferred tests, however, are the placebo test described above and the sensitivity analysis discussed in [Chapter @sec-panel-sens], which combines out-of-sample placebo estimates with post-treatment ATT estimates, but it requires a lot of power.
:::

We can implement the leave-one-out pre-trend test by setting `loo = TRUE`.

```{r fect_loo, eval=TRUE, cache = TRUE, message = FALSE}
out.fect.loo <- fect(Y ~ D + X1 + X2, data = sim_base, index = c("id","time"),
  method = "fe", force = "two-way", se = TRUE, loo = TRUE,
  parallel = TRUE, cores = 16, nboots = 1000)
```

The event study plot utilizing leave-one-out for pretreatment estimates is shown below. This graph is fairly similar to the graphics we presented earlier without using leave-one-out. However, this is not always true.

```{r plot-gap-loo, fig.width = 6, fig.height = 4.5}
plot(out.fect.loo,main = "Estimated ATT (FEct) -- LOO",
  cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8)
```

------------------------------------------------------------------------

## Cumulative effects

Users can use `effect()` to calculate cumulative treatment effects. The behavior of `effect()` is similar to the function of the same name in `gsynth`. Calculation of cumulative effects will need unit-time level bootstrap results. Choose the option `keep.sims=TRUE` to record them.

::: {.callout-note appearance="simple"}
The example below uses `method = "ife"` with `time.component.from = "nevertreated"`, which is equivalent to `method = "gsynth"` (see [Chapter @sec-gsynth]). We use the explicit form here for clarity.
:::

```{r cumu_effect, cache = TRUE}
out <- fect(Y ~ D + X1 + X2, data = sim_gsynth, index = c("id","time"),
                        method = "ife", time.component.from = "nevertreated",
                        force = "two-way", CV = TRUE, r = c(0, 5),
                        se = TRUE, nboots = 1000, vartype = 'bootstrap',
                        parallel = TRUE, cores = 16, keep.sims=TRUE)
cumu.out <- effect(out)
```

Print and plot cumulative effects

```{r cumu_effect_plot, cache = TRUE}
print(cumu.out)
plot(cumu.out)
```

Users can choose to calculate by-period average effects by setting `cumu=FALSE`.

```{r cumu_effect_byperiod, cache = TRUE}
effect(out, cumu=FALSE)
```

Calculate the cumulative effect of certain units at certain periods.

```{r cumu_effect_subset, cache = TRUE}
effect(out, cumu=TRUE, id=c(101,102,103), period=c(1,5))
```

`effect()` also accepts results from other estimation and inference methods. For example, we can use matrix completion:

```{r effect-mc, cache = TRUE}
out_mc <- fect(Y ~ D + X1 + X2, data = sim_gsynth, index = c("id","time"),
                        method = "mc", force = "two-way", CV = TRUE, r = c(0, 5),
                        se = TRUE, nboots = 1000, vartype = 'bootstrap',
                        parallel = TRUE, cores = 16, keep.sims=TRUE)
plot(effect(out_mc))
```

We can also use jackknife instead of bootstrap for inference:

```{r effect-jackknife, cache = TRUE}
out_jack <- fect(Y ~ D + X1 + X2, data = sim_gsynth, index = c("id","time"),
                        method = "mc", force = "two-way", CV = TRUE, r = c(0, 5),
                        se = TRUE, nboots = 1000, vartype = 'jackknife',
                        parallel = TRUE, cores = 16, keep.sims=TRUE)
plot(effect(out_jack))
```

------------------------------------------------------------------------

## Other estimands

After obtaining the individual treatment effects using one of the counterfactual estimators, we can weight these estimates using a constructed balanced treated sample or other user-supplied weighting schemes.

### Balanced treated sample

**fect** also provides the option `balance.period`, which allows the calculation of the average treatment effects only for *treated* units that exhibit complete data in specified pre- and post-treatment periods. For instance, if the option is set to `balance.period = c(-3,4)`, the algorithm will calculate the average treatment effects for units that have at least four consecutive non-missing observations in the pre-treatment periods `(-3, -2, -1, 0)` and at least four consecutive non-missing observations in the post-treatment periods `(1, 2, 3, 4)`. Note that this option does not affect whether a never-treated unit enters estimation.

```{r simdata_bal, eval=TRUE, cache = TRUE}
out.bal <- fect(Y ~ D + X1 + X2, data = sim_base, index = c("id","time"),
  balance.period = c(-3, 4), force = "two-way", method = "ife",
  CV = FALSE, r = 2, se = TRUE, nboots = 1000, parallel = TRUE, cores = 16)
```

We can then visualize the dynamic treatment effects using the inbuilt function `plot`. By default, it displays the dynamic treatment effects of the "balanced" sample.

```{r plot-balanced-att, fig.width = 6, fig.height = 4.5}
plot(out.bal, main = "Estimated ATT (Balanced Sample)")
```

The usual plotting options can be used to adjust the balanced plot as well.

```{r plot-balanced-custom, fig.width = 6, fig.height = 4.5}
plot(out.bal, main = "Estimated ATT (Balanced Sample)",
  post.color = "red", count.color = "blue")
```

### Average cohort treatment effect

**fect** allows us to estimate and visualize the ATTs for sub-groups of treated units. For example, it can draw the gap plot for units that adopt the treatment at the same time under staggered adoption, which is defined as "Cohort" in Sun & Abraham (2021). Our simulated dataset is not ideal to demonstrate this functionality because the treatment switches on and off. To improve feasibility, we define a cohort as a group of treated units that first adopt the treatment at the same time.

```{r simdata_panelview_cohort, fig.width = 6, fig.height = 4.5, warning = FALSE}
panelview(Y ~ D, data = sim_base, index = c("id","time"), by.timing = TRUE,
  axis.lab = "time", xlab = "Time", ylab = "Unit",
  background = "white", main = "Simulated Data: Treatment Status")
```

The `get.cohort()` function can generate a new variable "Cohort" based on the timing when treated units first get treated.

```{r get_cohort}
sim_base.cohort <- get.cohort(data = sim_base,D = 'D',index = c("id","time"))
print(table(sim_base.cohort[,'Cohort']))
```

We can also pass a list of **intervals** for first get-treated time into the `entry.time` option of `get.cohort()`. For example, we can categorize all treated units into the group that adopts the treatment between time 21 and 27, and the group that adopts the treatment in time 30 and 33.

```{r get_cohort2}
sim_base.cohort2 <- get.cohort(data = sim_base,D = 'D',index = c("id","time"),
                               entry.time = list(c(21,27),c(30,33)))
print(table(sim_base.cohort2[,'Cohort']))
```

By setting the option `group = "Cohort"`, **fect** estimates the ATT for each specified sub-group and saves it for further visualization.

```{r simdata_fe_cohort, eval = TRUE, cache = TRUE, message = FALSE, results='hide'}
out.fe.g <- fect(Y ~ D + X1 + X2, data = sim_base.cohort, index = c("id","time"),
          force = "two-way", method = "fe",
          se = TRUE, nboots = 1000, parallel = TRUE, cores = 16, group = 'Cohort')
```

Then one can draw the gap plot for each sub-group. Here we present the gap plot for Cohort 22.

```{r cohort_plot1, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.fe.g, show.group = "Cohort:22",
          xlim = c(-15, 10), ylim = c(-10, 10))
```

### User-supplied weights

The package offers the option `W` to calculate the weighted average treatment effects. The weighting variable does not affect the estimation of fixed effects or factors. Only the weighted average treatment effects or weighted dynamic treatment effects are obtained by aggregating the treatment effects using the weight `W`.

```{r simdata_w, eval=TRUE, cache = TRUE}
sim_base$Weight <- abs(rnorm(n = dim(sim_base)[1]))
out.w <- fect(Y ~ D + X1 + X2, data = sim_base, index = c("id","time"),
  force = "two-way", method = "ife", W = 'Weight',
  CV = FALSE, r = 2, se = TRUE, nboots = 1000, parallel = TRUE, cores = 16)
```

We can then visualize the weighted dynamic treatment effects using the inbuilt function `plot`, it by default shows the weighted dynamic treatment effects.

```{r plot-weighted-att, fig.width = 6, fig.height = 4.5}
plot(out.w, main = "Estimated Weighted ATT")
```


## Additional notes

1.  By default, the program will drop the units that have no larger than 5 observations under control, which is the reason why sometimes there are less available units in the placebo test or carryover test than in the original estimation. We can specify a preferred criterion in the option `min.T0` (default to 5). As a rule of thumb for the IFE estimator, the minimum number of observations under control for a unit should be larger than the specified number of factor `r`.

2.  We can get replicable results by setting the option `seed` to a certain integer, no matter whether the parallel computing is used.

3.  When `na.rm = FALSE` (default), the program allows observations to have missing outcomes $Y$ but not $X$ or treatment statuses $D$. When `na.rm = TRUE` the program will drop all observations that have missing values in outcomes, treatments, or covariates.


## How to Cite

If you find these methods helpful, you can cite @LWX2024.

```bibtex
@article{LWX2024,
  title = {A Practical Guide to Counterfactual Estimators for Causal Inference with Time-Series Cross-Sectional Data},
  author = {Liu, Licheng and Wang, Ye and Xu, Yiqing},
  journal = {American Journal of Political Science},
  volume = {68},
  number = {1},
  pages = {160--176},
  year = {2024}
}
```
