# The Imputation Estimator {#sec-fect}

In this chapter, we illustrate how to use the **fect** package to implement counterfactual estimators (or imputation estimators) and conduct diagnostic tests proposed by @LWX2022--\[<a href="https://onlinelibrary.wiley.com/doi/full/10.1111/ajps.12723" target="_blank">Paper</a>\].

## Simulated data

In this chapter, we use `simdata1`, a simulated panel dataset in which the parallel trends assumption holds. The dataset has 200 units, 35 time periods, and treatment that switches on and off. The data are simulated with treatment switching on and off, capturing the general case of treatment reversal under strict exogeneity. The outcome is generated from a two-way fixed effects model with two covariates:
$$Y_{it} = \alpha_i + \xi_t + \beta_1 X_{1,it} + \beta_2 X_{2,it} + \tau_{it} D_{it} + \epsilon_{it}$$

Since there are no latent factors, the `"fe"` method (two-way fixed effects counterfactual estimator) is correctly specified for this DGP. For settings where latent factors are present and the FE estimator is biased, see [Chapter @sec-ife-mc].

```{r echo = FALSE}
set.seed(1234)
rm(list = ls())
```

```{r message = FALSE, warning = FALSE}
library(fect)
data(fect)
```

We use the **panelView** package to visualize the treatment and outcome variables:

```{r message = FALSE, warning = FALSE, fig.width = 6, fig.height = 4.5}
library(panelView)
panelview(Y ~ D, data = simdata1, index = c("id","time"),
  axis.lab = "time", xlab = "Time", ylab = "Unit",
  gridOff = TRUE, by.timing = TRUE,
  background = "white", main = "Simulated Data: Treatment Status")
```

In the figure below, blue and gray represent treatment and control conditions.

```{r fig.width = 6, fig.height = 4.5}
panelview(Y ~ D, data = simdata1, index = c("id","time"),
  axis.lab = "time", xlab = "Time", ylab = "Unit",
  theme.bw = TRUE, type = "outcome",
  main = "Simulated Data: Outcome")
```

------------------------------------------------------------------------

## The FEct estimator

The two-way fixed effects counterfactual estimator (FEct) imputes the counterfactual outcome $\hat{Y}(0)$ for treated observations using a two-way fixed effects model estimated on control observations. It is also independently proposed by @BJS2024 and @gardnertwo, who refer to it as the "imputation method" and "two-stage DID," respectively.

### Estimation

We estimate the average treatment effect on the treated (ATT) using the following information: the outcome variable $Y$, binary treatment variable $D$, two observed covariates $X_{1}$ and $X_{2}$, and the unit and time indicators $id$ and $time$, respectively.

The first variable on the right hand side of the formula is the treatment indicator $D$; the rest of the right-hand-side variables serve as controls. The `index` option specifies the unit and time indicators. The `force` option ("none", "unit", "time", and "two-way") specifies the additive component(s) of the fixed effects included in the model. The default option is "two-way" (including both unit and time fixed effects).

```{r simdata_fect_nose, eval=TRUE, cache = TRUE, message = FALSE, results = 'hide'}
out.fect <- fect(Y ~ D + X1 + X2, data = simdata1, index = c("id","time"),
                 method = "fe", force = "two-way")
```

::: {.callout-tip appearance="simple"}
### Estimation parameters
The key parameters that control the **model** are: `method` (estimator choice), `force` (fixed effects structure), and `r` (number of factors for IFE/MC methods; see [Chapter @sec-ife-mc]). These determine how $\hat{Y}(0)$ is imputed.
:::

We can use the `plot` function to visualize the estimation results. By default, it produces a "gap" plot --- `plot(out.fect, type = "gap")` --- which shows the estimated period-wise ATT (dynamic treatment effects). The true population average effects in `simdata1` go from 1 to 3 from the 1st to the 10th post-treatment period.

```{r fect_plot_nose, fig.width = 6, fig.height = 4.5}
plot(out.fect, main = "Estimated ATT (FEct)", ylab = "Effect of D on Y",
  cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8)
```

In the gap plot, pre-treatment estimates appear in gray while post-treatment estimates appear in black. This visual distinction highlights that pre-treatment "effects" should be near zero if the model is well-specified, while post-treatment effects are the quantities of interest. The uncertainty estimates are unavailable in the plot above because, by default, `se = FALSE` to save computational power. See [Chapter @sec-plots] for more visualization options.

### Save estimates

Users can use the `print` function to view a summary of the estimation results or retrieve relevant statistics by directly accessing the fect object. Specifically, `est.avg` and `est.avg.unit` show the ATT averaged over all periods -- the former weights each treated observation equally while the latter weights each treated unit equally. `beta` reports the coefficients of the time-varying covariates. `est.att` reports the average treatment effect on the treated (ATT) by period.

```{r}
print(out.fect)
```

To save space, results are not shown here.

```{r eval = FALSE}
out.fect$est.att
out.fect$est.avg
out.fect$beta
```

------------------------------------------------------------------------

## Visualization and diagnostics

Once we have point estimates, we typically want uncertainty estimates and diagnostic checks before interpreting results. This section covers inference, pre-trend diagnostics, and the leave-one-out approach.

::: {.callout-tip appearance="simple"}
### Inference parameters
The key parameters that control **uncertainty estimation** are: `se` (enable standard errors), `vartype` (`"bootstrap"`, `"jackknife"`, or `"parametric"`), `nboots` (number of bootstrap replications), and `cl` (clustering variable). These do not affect point estimates.
:::

### Inference

The package can produce uncertainty estimates when `se = TRUE`. The default is the non-parametric cluster-bootstrap (`vartype = "bootstrap"`), which works well when the number of units is relatively large and many units experience the treatment condition. The number of bootstrap runs can be set by `nboots`.

::: {.callout-note appearance="simple"}
**Parallel computing** will speed up both cross-validation and uncertainty estimation significantly. We recommend that users manually set the number of cores using the `cores` option. If this is not supplied or is `NULL`, we will automatically select the smaller of `8` and the number of usable system cores minus `2`, to prevent excessive use of system resources.
:::

```{r simdata_fect, eval=TRUE, cache = TRUE, message = FALSE, results = 'hide'}
out.fect <- fect(Y ~ D + X1 + X2, data = simdata1, index = c("id","time"),
  method = "fe", force = "two-way", se = TRUE,
  cores = 8, parallel = TRUE, nboots = 1000)
```

::: {.callout-note appearance="simple"}
**Clustering.**

-   By default, `fect()` uses cluster-bootstrap at the `unit` level when `se = TRUE`, that is `vartype = "bootstrap"`. The uncertainty estimates thus account for arbitrary serial correlation within a unit over time---commonly understood as being "clustered" at the unit level. In the example above, the unit is `id`, hence, by default, `cl = "id"`.
-   Alternatively, users can obtain uncertainty estimates using the cluster-jackknife method by specifying `vartype = "jackknife"`, also at the unit level. In this case, the algorithm calculates standard errors by iteratively dropping one unit (i.e., the entire time series) from the dataset. This can be particularly useful when the number of treated units is small.
-   To cluster standard errors at a different, usually higher, level, users can specify the clustering variable using the `cl` option.
:::

The `plot()` function can visualize the estimated period-wise ATTs as well as their uncertainty estimates. `stats = "F.p"` shows the p-value for the F test of no-pretrend.

```{r fect_plot_nse, fig.width = 6, fig.height = 4.5}
plot(out.fect, main = "Estimated ATT (FEct)", ylab = "Effect of D on Y",
  cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")
```

### Diagnostics

A basic sanity check is the placebo test: remove some pre-treatment periods from estimation and verify the model does not find spurious effects there.

```{r fect_placebo, eval=TRUE, cache=TRUE, message=FALSE, results='hide', fig.width=6, fig.height=4.5}
out.fect.placebo <- fect(Y ~ D + X1 + X2, data = simdata1, index = c("id","time"),
  force = "two-way", method = "fe",
  se = TRUE, cores = 8, nboots = 200, parallel = TRUE,
  placeboTest = TRUE, placebo.period = c(-2, 0))
plot(out.fect.placebo, cex.text = 0.8)
```

Similarly, the carryover test checks whether treatment effects persist after treatment ends. Both tests are covered in detail in [Chapter @sec-ife-mc].

### Leave-one-out approach

@li2025benchmarking show that in some applications, pre-trend estimates based on in-sample model fit can lead to the mistaken belief that no pre-trend exists, even when a non-parallel pre-trend is present. A simple fix is to use a leave-one-out method by setting `loo = TRUE` to obtain these estimates, although it is significantly more time-consuming.

::: {.callout-note appearance="simple"}
We recommend setting `loo = TRUE` when (i) the event-study plot is intended as a critical piece of evidence to support the parallel trends assumption, which is often the case, or (ii) when implementing an equivalence test for the pre-trend estimates. For more discussion on the LOO pre-trend test, see [Chapter @sec-ife-mc].

Our most preferred test, however, is the sensitivity analysis discussed in [Chapter @sec-panel-sens], which combines out-of-sample placebo estimates with post-treatment ATT estimates.
:::

We can implement the leave-one-out pre-trend test by setting `loo = TRUE`.

```{r fect_loo, eval=TRUE, cache = TRUE, message = FALSE}
out.fect.loo <- fect(Y ~ D + X1 + X2, data = simdata1, index = c("id","time"),
  method = "fe", force = "two-way", se = TRUE, loo = TRUE,
  cores = 8, parallel = TRUE, nboots = 200)
```

The event study plot utilizing leave-one-out for pretreatment estimates is shown below. This graph is fairly similar to the graphics we presented earlier without using leave-one-out. However, this is not always true.

```{r, fig.width = 6, fig.height = 4.5}
plot(out.fect.loo,main = "Estimated ATT (FEct) -- LOO",
  cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8)
```

### Save estimates

After estimation with `se = TRUE`, bootstrap treatment effect estimates from each run are stored in `eff.boot`, an array whose dimension = (#time periods \* #treated \* #bootstrap runs). Standard errors for the period-wise ATTs are available in the output object alongside the point estimates.

```{r eval = FALSE}
out.fect$eff.boot
```

For IFE and MC methods, see [Chapter @sec-ife-mc].

------------------------------------------------------------------------

## Other estimands

After obtaining the individual treatment effects using one of the counterfactual estimators, we can weight these estimates using a constructed balanced treated sample or other user-supplied weighting schemes.

### Balanced treated sample

**fect** also provides the option `balance.period`, which allows the calculation of the average treatment effects only for *treated* units that exhibit complete data in specified pre- and post-treatment periods. For instance, if the option is set to `balance.period = c(-3,4)`, the algorithm will calculate the average treatment effects for units that have at least four consecutive non-missing observations in the pre-treatment periods `(-3, -2, -1, 0)` and at least four consecutive non-missing observations in the post-treatment periods `(1, 2, 3, 4)`. Note that this option does not affect whether a never-treated unit enters estimation.

```{r simdata_bal, eval=TRUE, cache = TRUE}
out.bal <- fect(Y ~ D + X1 + X2, data = simdata1, index = c("id","time"),
  balance.period = c(-3, 4), force = "two-way", method = "ife",
  CV = FALSE, r = 2, se = TRUE, nboots = 200, parallel = TRUE)
```

We can then visualize the dynamic treatment effects using the inbuilt function `plot`. By default, it displays the dynamic treatment effects of the "balanced" sample.

```{r fig.width = 6, fig.height = 4.5}
plot(out.bal, main = "Estimated ATT (Balanced Sample)")
```

The usual plotting options can be used to adjust the balanced plot as well.

```{r fig.width = 6, fig.height = 4.5}
plot(out.bal, main = "Estimated ATT (Balanced Sample)",
  color = "red", count.color = "blue")
```

### User-supplied weights

The package offers the option `W` to calculate the weighted average treatment effects. The weighting variable does not affect the estimation of fixed effects or factors. Only the weighted average treatment effects or weighted dynamic treatment effects are obtained by aggregating the treatment effects using the weight `W`.

```{r simdata_w, eval=TRUE, cache = TRUE}
simdata1$Weight <- abs(rnorm(n = dim(simdata1)[1]))
out.w <- fect(Y ~ D + X1 + X2, data = simdata1, index = c("id","time"),
  force = "two-way", method = "ife", W = 'Weight',
  CV = FALSE, r = 2, se = TRUE, nboots = 200, parallel = TRUE)
```

We can then visualize the weighted dynamic treatment effects using the inbuilt function `plot`, it by default shows the weighted dynamic treatment effects.

```{r fig.width = 6, fig.height = 4.5}
plot(out.w, main = "Estimated Weighted ATT")
```

### Average cohort treatment effect

**fect** allows us to estimate and visualize the ATTs for sub-groups of treated units. For example, it can draw the gap plot for units that adopt the treatment at the same time under staggered adoption, which is defined as "Cohort" in Sun & Abraham (2021). Our simulated dataset is not ideal to demonstrate this functionality because the treatment switches on and off. To improve feasibility, we define a cohort as a group of treated units that first adopt the treatment at the same time.

```{r simdata_panelview_cohort, fig.width = 6, fig.height = 4.5}
panelview(Y ~ D, data = simdata1, index = c("id","time"), by.timing = TRUE,
  axis.lab = "time", xlab = "Time", ylab = "Unit",
  background = "white", main = "Simulated Data: Treatment Status")
```

The `get.cohort()` function (originally from the **paneltools** package) can generate a new variable "Cohort" based on the timing when treated units first get treated. The new version of **fect** incorporates the feature: users no longer need to install any complementary packages to replicate the tutorial.

```{r get_cohort}
#  devtools:: install_github("xuyiqing/paneltools" if not already installed
simdata1.cohort <- get.cohort(data = simdata1,D = 'D',index = c("id","time"))
print(table(simdata1.cohort[,'Cohort']))
```

We can also pass a list of **intervals** for first get-treated time into the `entry.time` option of `get.cohort()`. For example, we can categorize all treated units into the group that adopts the treatment between time 21 and 27, and the group that adopts the treatment in time 30 and 33.

```{r get_cohort2}
simdata1.cohort2 <- get.cohort(data = simdata1,D = 'D',index = c("id","time"),
                               entry.time = list(c(21,27),c(30,33)))
print(table(simdata1.cohort2[,'Cohort']))
```

By setting the option `group = "Cohort"`, **fect** estimates the ATT for each specified sub-group and saves it for further visualization.

```{r simdata_fe_cohort, eval = TRUE, cache = TRUE, message = FALSE, results='hide'}
out.fe.g <- fect(Y ~ D + X1 + X2, data = simdata1.cohort, index = c("id","time"),
          force = "two-way", method = "fe",
          se = TRUE, nboots = 200, parallel = TRUE, group = 'Cohort')
```

Then one can draw the gap plot for each sub-group. Here we present the gap plot for Cohort 22.

```{r cohort_plot1, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.fe.g, show.group = "Cohort:22",
          xlim = c(-15, 10), ylim = c(-10, 10))
```

------------------------------------------------------------------------

## Additional notes

1.  By default, the program will drop the units that have no larger than 5 observations under control, which is the reason why sometimes there are less available units in the placebo test or carryover test than in the original estimation. We can specify a preferred criteria in the option `min.T0` (default to 5). As a rule of thumb for the IFE estimator, the minimum number of observations under control for a unit should be larger than the specified number of factor `r`.

2.  We can get replicable results by setting the option `seed` to a certain integer, no matter whether the parallel computing is used.

3.  When `na.rm = FALSE` (default), the program allows observations to have missing outcomes $Y$ or covariates $X$ but decided treatment statuses $D$. Otherwise the program will drop all observations that have missing values in outcomes, treatments, or covariates.
