# Alternative Estimands {#sec-estimands}

The default output of `fect()` is the per-event-time average treatment effect on the treated (ATT). This chapter introduces the estimand interface added in v2.4.0. It has two functions: `estimand()`, which computes alternative summaries directly from a fitted model, and `imputed_outcomes()`, which returns the cell-level imputed potential outcomes so you can build summaries that `estimand()` does not include.

`estimand()` covers four common requests in one line each: cumulative ATT, average proportional treatment effect on the treated [APTT, @chen2024logs], log-scale ATT, and ATT restricted to a window of event times. Each call returns a data frame ready for plotting with fect's `esplot()`. For anything else, `imputed_outcomes()` returns the cell-level data in long format and the rest is standard data manipulation.

R script used in this chapter can be downloaded [here](https://raw.githubusercontent.com/xuyiqing/fect/dev/vignettes/rscript/03-estimands.R).

## Beyond the default ATT

Use `estimand()` when you want any of the following: the per-event-time level ATT as a data frame (instead of the legacy matrix); the cumulative ATT through each event time, or summed over a fixed window; APTT [@chen2024logs]; log-scale ATT; or a level ATT restricted to a window of event times --- say, the first five post-treatment periods.

For summaries that `estimand()` does not cover, `imputed_outcomes()` returns the cell-level data in long format and you do the aggregation yourself. 

## Setup

```{r .common, include = FALSE}
source("_common.R")
```

The examples below use a small synthetic panel with positive outcomes so that log-scale calculations are well defined. The fit sets `keep.sims = TRUE` so that the cell-level imputed counterfactuals from each bootstrap replicate are saved on the fit object. APTT and log-scale ATT need these saved values to recompute their non-linear formulas inside each replicate. The level ATT and cumulative ATT use the pre-aggregated bootstrap that fect always keeps, so the default `keep.sims` is fine for those.

```{r setup-estimand, eval = TRUE, message = FALSE, results = 'hide'}
library(dplyr)

set.seed(1)
N <- 60; TT <- 25
df <- expand.grid(id = 1:N, time = 1:TT)
treat_start <- sample(c(NA, 8:18), N, replace = TRUE)
df$D <- ifelse(is.na(treat_start[df$id]) | df$time < treat_start[df$id],
               0, 1)
## Higher intercept (2.0) and smaller noise (sd = 0.1) keep Y safely
## positive throughout, so the v2.4.2+ cell-drop hard-error in
## log.att / aptt does not fire on benign bootstrap noise.
df$Y <- exp(2.0 + 0.05 * df$time + 0.3 * df$D + rnorm(nrow(df), sd = 0.1))

fit <- fect(Y ~ D, data = df, index = c("id", "time"),
            method = "fe", force = "two-way",
            se = TRUE, nboots = 1000, parallel = FALSE,
            keep.sims = TRUE)
```

## Per-event-time level ATT

```{r est-att-event-time, eval = TRUE}
estimand(fit, "att", "event.time") |>
  head(8)
```

Each row is one event time. The columns hold the point estimate, the standard error, the bounds of the confidence interval, the number of treated cells that contributed at that event time, and the variance type used for inference. The numbers match the legacy `fit$est.att` exactly. The only difference is the data-frame format, which works directly with downstream plotting and aggregation tools --- including fect's own `esplot()`:

```{r plot-att-event-time, eval = TRUE, fig.width = 6, fig.height = 4.5}
est <- estimand(fit, "att", "event.time")
esplot(est, Period = "event.time",
       Estimate = "estimate", CI.lower = "ci.lo", CI.upper = "ci.hi",
       Count = "n_cells",
       main = "Per-event-time ATT")
```

## Cumulative ATT

```{r est-cumu-event-time, eval = TRUE}
estimand(fit, "att.cumu", "event.time") |>
  head(8)
```

The cumulative ATT through each event time. Each row reports the running sum of the per-period average effect, computed cell by cell. This call replaces the legacy `effect()` function and produces identical numbers.

```{r plot-cumu-event-time, eval = TRUE, fig.width = 6, fig.height = 4.5}
cumu <- estimand(fit, "att.cumu", "event.time")
esplot(cumu, Period = "event.time",
       Estimate = "estimate", CI.lower = "ci.lo", CI.upper = "ci.hi",
       Count = "n_cells",
       main = "Cumulative ATT",
       ylab = "Cumulative Effect")
```

For a single overall window, request the overall grouping axis instead:

```{r est-cumu-overall, eval = TRUE}
estimand(fit, "att.cumu", "overall", window = c(1, 5))
```

The result is the cumulative effect over the first five post-treatment periods. It matches the final row of the legacy `att.cumu(fit, period = c(1, 5))` output.

## APTT

```{r est-aptt, eval = TRUE}
estimand(fit, "aptt", "event.time") |>
  head(8)
```

The average proportional treatment effect on the treated [@chen2024logs]:

$$
\text{APTT}_t \;=\; \frac{\mathbb{E}[\,Y - \widehat{Y}(0) \mid D = 1, t\,]}{\mathbb{E}[\,\widehat{Y}(0) \mid D = 1, t\,]}.
$$

Inside each bootstrap replicate, both the numerator and the denominator are recomputed and then divided. The bootstrap distribution is therefore a distribution of ratios, not a ratio of two distributions of means. This requires the cell-level imputed counterfactuals, which is why APTT needs `keep.sims = TRUE` at fit time.

```{r plot-aptt-event-time, eval = TRUE, fig.width = 6, fig.height = 4.5}
aptt <- estimand(fit, "aptt", "event.time")
esplot(aptt, Period = "event.time",
       Estimate = "estimate", CI.lower = "ci.lo", CI.upper = "ci.hi",
       Count = "n_cells",
       main = "APTT by event time",
       ylab = "Average Proportional TE on the Treated")
```

## Log-scale ATT

```{r est-log-att, eval = TRUE}
estimand(fit, "log.att", "event.time") |>
  head(8)
```

The mean log-scale treatment effect: at each event time, the average over treated cells of $\log Y_{it} - \log \widehat{Y}_{it}(0)$. Cells where either the observed outcome or the imputed counterfactual is non-positive are dropped from the point estimate, since the logarithm is undefined there.

When a cell used in the point estimate has $\widehat{Y}_{it}(0)_b \le 0$ in more than 5% of bootstrap replicates, `estimand("log.att", ...)` errors with guidance on what to do (filter the unstable cells, transform the outcome, or use a different estimand). The example data above uses an exponential-link outcome with an intercept high enough that no cell's bootstrap distribution crosses zero.

```{r plot-log-att-event-time, eval = TRUE, fig.width = 6, fig.height = 4.5}
log_att <- estimand(fit, "log.att", "event.time")
esplot(log_att, Period = "event.time",
       Estimate = "estimate", CI.lower = "ci.lo", CI.upper = "ci.hi",
       Count = "n_cells",
       main = "Log-scale ATT",
       ylab = "Mean log(Y) − log(Y0)")
```

## Placebo tests

Starting in v2.4.2, `estimand()` takes a `test = c("none", "placebo", "carryover")` argument that lets you compute placebo and carryover versions of any alternative estimand, not just the default `att.placebo` / `att.carryover` scalars (see issue #131). The placebo case requires a fit produced with `placeboTest = TRUE` and a placebo period.

```{r setup-placebo, eval = TRUE, message = FALSE, results = 'hide'}
## Re-fit with placeboTest = TRUE so that estimand() can recompute
## APTT / log-ATT at the placebo cells.
fit_placebo <- fect(Y ~ D, data = df, index = c("id", "time"),
                    method = "fe", force = "two-way",
                    se = TRUE, nboots = 200, parallel = FALSE,
                    keep.sims = TRUE,
                    placeboTest = TRUE, placebo.period = c(-2, 0))
```

The placebo APTT series should be statistically indistinguishable from zero across the placebo window if the parallel-counterfactual assumption holds; non-zero placebo estimates are evidence against identification.

```{r plot-aptt-placebo, eval = TRUE, fig.width = 6, fig.height = 4.5}
aptt_placebo <- estimand(fit_placebo, "aptt", "event.time",
                         test = "placebo")
esplot(aptt_placebo, Period = "event.time",
       Estimate = "estimate", CI.lower = "ci.lo", CI.upper = "ci.hi",
       Count = "n_cells",
       main = "APTT placebo (pre-treatment)",
       ylab = "Average Proportional TE on the Treated")
```

The call gives a warning that BCa confidence intervals with `nboots = 200` may be noisy at the endpoints. This is a caution, not an error. BCa uses the 2.5% and 97.5% quantiles of the bootstrap distribution to form the CI, and these tail quantiles can vary with only 200 bootstrap runs. Efron (1987) and DiCiccio and Efron (1996) recommend at least 1000 runs for publication-quality tail CIs. The point estimate and standard error are not affected by `nboots`. This example uses 200 to keep the chapter fast. For more stable CIs in your own analysis, refit with `fect(..., nboots = 1000)`.

The same logic carries over to log-scale ATT under placebo:

```{r plot-log-att-placebo, eval = TRUE, fig.width = 6, fig.height = 4.5}
log_att_placebo <- estimand(fit_placebo, "log.att", "event.time",
                            test = "placebo")
esplot(log_att_placebo, Period = "event.time",
       Estimate = "estimate", CI.lower = "ci.lo", CI.upper = "ci.hi",
       Count = "n_cells",
       main = "Log-ATT placebo (pre-treatment)",
       ylab = "Mean log(Y) − log(Y0)")
```

## Carryover tests

The carryover case requires `carryoverTest = TRUE` and a panel with treatment reversals. The carryover series uses cells in `fit$carryover.period`---the early post-reversal window where, if the treatment effect dissipates immediately, estimates should be zero. Non-zero carryover estimates suggest the effect persists past treatment removal. We first build a panel with treatment reversals (each treated unit's spell ends after a uniformly-drawn 5--10 periods) and visualize the on/off pattern with `panelview()`:

```{r setup-carryover, eval = TRUE, message = FALSE, results = 'hide'}
## Build a panel with treatment reversals for the carryover demo.
set.seed(2)
df_rev <- df
treat_end <- pmin(treat_start[df_rev$id] + sample(5:10, N, replace = TRUE),
                  TT + 1L)
df_rev$D <- ifelse(is.na(treat_start[df_rev$id]) |
                   df_rev$time < treat_start[df_rev$id] |
                   df_rev$time >= treat_end[df_rev$id],
                   0, 1)
df_rev$Y <- exp(2.0 + 0.05 * df_rev$time + 0.3 * df_rev$D +
                rnorm(nrow(df_rev), sd = 0.1))
```

```{r panelview-carryover, eval = TRUE, fig.width = 7, fig.height = 4.5}
panelView::panelview(Y ~ D, data = df_rev, index = c("id", "time"),
                     by.timing = TRUE, axis.lab = "time",
                     main = "Treatment-reversal panel for carryover demo")
```

```{r fit-carryover, eval = TRUE, message = FALSE, results = 'hide'}
fit_carry <- fect(Y ~ D, data = df_rev, index = c("id", "time"),
                  method = "fe", force = "two-way",
                  se = TRUE, nboots = 1000, parallel = FALSE,
                  keep.sims = TRUE,
                  carryoverTest = TRUE, carryover.period = c(1, 2))
```

```{r plot-aptt-carryover, eval = TRUE, fig.width = 6, fig.height = 4.5}
aptt_carry <- estimand(fit_carry, "aptt", "event.time",
                       test = "carryover")
esplot(aptt_carry, Period = "event.time",
       Estimate = "estimate", CI.lower = "ci.lo", CI.upper = "ci.hi",
       Count = "n_cells",
       main = "APTT carryover (post-reversal)",
       ylab = "Average Proportional TE on the Treated")
```

```{r plot-log-att-carryover, eval = TRUE, fig.width = 6, fig.height = 4.5}
log_att_carry <- estimand(fit_carry, "log.att", "event.time",
                          test = "carryover")
esplot(log_att_carry, Period = "event.time",
       Estimate = "estimate", CI.lower = "ci.lo", CI.upper = "ci.hi",
       Count = "n_cells",
       main = "Log-ATT carryover (post-reversal)",
       ylab = "Mean log(Y) − log(Y0)")
```

`type = "att.cumu"` is rejected with a clear error when `test != "none"`. Cumulative ATT is defined relative to treatment onset, so it has no meaning at placebo or post-reversal cells.

## Window-restricted ATT

Pass a `window` argument to focus on a subset of event times. The function keeps cells whose event time falls inside the closed interval and computes the requested estimand over them.

```{r est-window-overall, eval = TRUE}
estimand(fit, "att", "overall", window = c(1, 5))
```

The result is a single overall ATT averaged over treated cells with event time between one and five. Use this when the question is "what is the average ATT for the first five post-treatment years?" and you do not need a per-event-time series. The answer is one number with one confidence interval, so a series plot is not relevant; if you want to compare multiple windows, build a small data frame of `(window_label, estimate, ci.lo, ci.hi)` and pass it to `esplot()` with `Period = "window_label"`.

## Custom estimands

If the summary you need is not one of the four built-in types, you can compute it yourself from the cell-level data:

```{r imp-out-glance, eval = TRUE}
po <- imputed_outcomes(fit)
head(po)
```

Each row is one treated cell. The columns are the unit and time coordinates, the observed outcome, the imputed counterfactual, the cell-level effect (the cell's contribution to the ATT estimator), and the aggregation weight. Setting `replicates = TRUE` expands the data so there is one row per cell per bootstrap replicate.

```{r imp-out-rep, eval = TRUE}
po_rep <- imputed_outcomes(fit, replicates = TRUE)
nrow(po_rep) == nrow(po) * 200    # one row per (cell, replicate)
```

An example: the per-event-time standard deviation of the cell-level treatment effect, a simple heterogeneity diagnostic.

```{r custom-estimand, eval = TRUE}
po |>
  group_by(event.time) |>
  summarise(sd_eff = sd(eff, na.rm = TRUE),
            n     = dplyr::n(),
            .groups = "drop") |>
  head(8)
```

To get bootstrap confidence intervals for a custom summary, group the replicate-expanded data by event time and replicate, compute the summary inside each group, then take the appropriate quantiles across replicates. ggplot2 is the most flexible option for plotting; the data-frame format keeps the plot code as short as the aggregation code.

## Save bootstrap runs

Setting `keep.sims = TRUE` at fit time saves the cell-level imputed counterfactuals from every bootstrap replicate on the fit object --- a three-dimensional array of dimension $T \times N \times \text{nboots}$. Some estimands need it, others do not.

The level ATT and the cumulative ATT (over either the event-time axis or an overall window) work without it, because they read the pre-aggregated bootstrap that fect always keeps. The other types --- APTT, log-scale ATT, and any call that filters cells or groups by something other than event time --- recompute their aggregations over the cell-level array and therefore need the saved runs.

A back-of-envelope memory cost:

| Panel size (T × N) | nboots | Memory cost |
| --- | --- | --- |
| 1,000 | 200 | 1.6 MB |
| 10,000 | 200 | 16 MB |
| 50,000 | 200 | 80 MB |
| 50,000 | 500 | 200 MB |
| 200,000 | 500 | 800 MB |

If you fit without `keep.sims = TRUE` and then call an estimand that needs it, the function errors with a message pointing back to the option.

```{r keep-sims-error, eval = FALSE}
# No bootstrap/jackknife results available. Choose keep.sims = TRUE in fect().
```

## Parametric variance

Parametric bootstrap (`vartype = "parametric"`) is one of fect's
fit-time variance options alongside `"bootstrap"` (the default) and
`"jackknife"`. The parametric option resamples residuals from a
fitted factor model rather than resampling units; it requires
`time.component.from = "nevertreated"` (see
[Section @sec-parametric-regimes]).

`estimand()` treats parametric replicates the same as bootstrap or
jackknife replicates: they are stored in `fit$eff.boot` and all four
`type` values use them through the same code. In v2.4.1, the
`vartype` argument of `estimand()` accepts `"parametric"` explicitly
so callers can name the method they expect.

For details on how `vartype`, `ci.method`, and the parametric `para.error`
sub-option combine to form the bootstrap distribution, see
[Chapter @sec-inference].

```{r parametric-att, eval = FALSE}
fit_para <- fect(Y ~ D, data = sim_linear, index = c("id", "time"),
                 method = "ife", force = "two-way",
                 r = 2, CV = FALSE, se = TRUE, nboots = 200,
                 keep.sims = TRUE,
                 vartype = "parametric",
                 time.component.from = "nevertreated",
                 parallel = FALSE)

est <- estimand(fit_para, "att", "event.time")
head(est)
#>   event.time     estimate        se      ci.lo     ci.hi n_cells    vartype
#> 1        -39 -0.019958112 0.2427538 -0.4957469 0.4558307      80 parametric
#> 2        -38 -0.006695018 0.1693391 -0.3385935 0.3252035      80 parametric
#> ...
```

The `vartype` column reports `"parametric"` --- the method actually
used at fit time --- regardless of what you pass as the `vartype`
argument. The argument is informational; it does not re-aggregate
replicates. The output matches `fit_para$est.att` exactly for the
per-event-time path.

For the other types (`"att.cumu"`, `"aptt"`, `"log.att"`, and
`"att"` with `by = "overall"`), `estimand()` aggregates quantiles
over `fit$eff.boot`. Under parametric, the simulated distribution is
centered on the parametric DGP's mean rather than on the point
estimate, so quantile-based confidence intervals reflect the
parametric sampling distribution rather than a reflection around the
point estimate. This matches what `fit$est.att` reports under
parametric.

## Migration

The legacy functions `effect()` and `att.cumu()` still work and produce identical results, but now emit a one-time-per-session deprecation message pointing here. Direct replacements:

| Today | Replacement |
| --- | --- |
| `effect(fit, cumu = TRUE)` | `estimand(fit, "att.cumu", "event.time")` |
| `effect(fit, cumu = FALSE)` | `estimand(fit, "att", "event.time")` |
| `att.cumu(fit, period = c(L, R))` | `estimand(fit, "att.cumu", "overall", window = c(L, R))` |

Outputs from `estimand()` and the legacy functions match to all decimals (verified by package tests). The only user-facing difference is that `estimand()` returns a data frame instead of the legacy matrix.
