# Factor-Augmented Approaches {#sec-ife-mc}

```{r echo = FALSE, message = FALSE, warning = FALSE}
set.seed(1234)
library(fect)
data(fect)
```

When the parallel trends assumption is violated due to latent common factors with heterogeneous loadings, the FE estimator from [Chapter @sec-fect] is biased. This chapter introduces two methods that account for such latent factors: the **interactive fixed effects** (IFE) method, which explicitly models unit-specific factor loadings, and the **matrix completion** (MC) method, which uses nuclear-norm regularization to recover the low-rank structure of the untreated potential outcomes.

We use `simdata2`, which includes two latent factors ($r = 2$). The FE estimator is biased on this dataset, while IFE and MC recover the correct ATT.

## Interactive fixed effects

In addition to FEct, **fect** supports the interactive fixed effects counterfactual (IFEct) method proposed by @Gobillon2016 and @Xu2017 and the matrix completion (MC) method proposed by @Athey2021---`method = "ife"` and `method = "mc"`, respectively. The EM algorithm is used to impute the counterfactuals of treated observations.

### IFE

For the IFE approach, we need to specify the number of factors using option `r`. By default, the algorithm will select an optimal hyper-parameter via a built-in cross-validation procedure. For a detailed treatment of cross-validation options, see [Chapter @sec-model-selection].

We specify an interval of candidate number of unobserved factors in option `r` like `r=c(0,5)`. When cross-validation is switched off, the first element in `r` will be set as the number of factors. Below we use the MSPE criterion and search the number of factors from 0 to 5.

```{r simdata_ife, eval=TRUE, cache = TRUE}
out.ife <- fect(Y ~ D + X1 + X2, data = simdata2, index = c("id","time"),
          force = "two-way", method = "ife", CV = TRUE, r = c(0, 5),
          se = TRUE, cores = 8, nboots = 1000, parallel = TRUE)
print(out.ife)
```

The figure below shows the estimated ATT using the IFE method. The cross-validation procedure selects the correct number of factors ($r=2$).

```{r  fig.width = 6, fig.height = 4.5}
plot(out.ife, main = "Estimated ATT (IFEct)")
```

### MC

For the MC method, we need to specify the tuning parameter in the penalty term using option `lambda`. If users don't have any prior knowledge to set candidate tuning parameters, a number of candidate tuning parameters can be generated automatically based on the information from the outcome variable. We specify the number in option `nlambda`, e.g. `nlambda = 10`.

```{r simdata_mc, eval=TRUE, cache = TRUE}
out.mc <- fect(Y ~ D + X1 + X2, data = simdata2, index = c("id","time"),
          force = "two-way", method = "mc", CV = TRUE,
          se = TRUE, cores = 8, nboots = 1000, parallel = TRUE)

print(out.mc)
```

```{r fig.width = 6, fig.height = 4.5}
plot(out.mc, main = "Estimated ATT (MC)")
```

::: {.callout-note appearance="simple"}
### The `em` parameter
By default, `em = TRUE` and the EM algorithm is used to estimate the factor model when the estimation sample has missing entries. This is always the case in the default DID setting (`factors.from = "notyettreated"`), where treated post-treatment cells are unobserved under control. For the never-treated setting, see [Chapter @sec-gsynth].
:::

------------------------------------------------------------------------

## Quick diagnostic check

We verify the IFE model passes basic pre-trend checks using the equivalence plot:

```{r ife_quick_equiv, eval=TRUE, cache=TRUE, fig.width=6, fig.height=4.5}
plot(out.ife, type = "equiv", ylim = c(-4, 4),
     cex.legend = 0.6, main = "Testing Pre-Trend (IFEct)", cex.text = 0.8)
```

For a comprehensive treatment of diagnostic tests including placebo, carryover, and LOO pre-trend tests, see [Chapter @sec-diagnostics]. For cross-validation details and model selection, see [Chapter @sec-model-selection].
