# Fect Plot Options {#sec-plots}

In this chapter, we explore various visualization options available in the **fect** package using data from @GS2020. Download the R code used in this chapter [here](rscript/03-plot.R).

------------------------------------------------------------------------

`plot.fect` is an S3 method that offers various options for customizing data and results visualization. Below is a brief summary of the most commonly used options.

-   **Starting Period:**
    -   `start0`: If `TRUE`, shifts the time axis so that treatment begins at Period 0 instead of Period 1.
-   **Confidence Intervals:**
    -   `plot.ci`: Options include `"none"`, `"0.9"`, or `"0.95"` to hide confidence intervals or display 90% or 95% confidence intervals.
-   **Axis and Legend Customization:**
    -   `xlim` / `ylim`: Set the x- and y-axis ranges.\
    -   `xlab` / `ylab`: Customize axis labels.\
    -   `xbreaks` / `ybreaks`: Specify tick marks.\
    -   `xangle` / `yangle`: Adjust the rotation angle of axis text.\
    -   `legend.pos`, `legend.nrow`, `legend.labs`: Control legend placement, number of rows, and labels.
-   **Theme and Text:**
    -   `theme.bw`: If `TRUE`, applies a black-and-white theme.
    -   `preset`: If `NULL`, will be the default color preset, which is mostly black and white with a bit of color. Other options include `"vibrant"` and `"grayscale"`.\
    -   `cex.main`, `cex.axis`, `cex.lab`, `cex.text`: Adjust text sizes for the title, tick labels, axis labels, and annotations.
-   **Lines and Bounds:**
    -   `color` / `est.lwidth`: Define the color and width of main lines.\
    -   `lcolor` / `lwidth` / `ltype`: Set the color, width, and line type for the axes. Takes a vector, where the first value is applied the horizontal axis and the second is applied the vertical axis. If only one value is given, both axes will take on the same value.

While these customization options are demonstrated using the default `gap` plot, they can be applied universally, with only a few exceptions.

## Load Data

We will be using two datasets in this chapter. As explained in @sec-panel, @GS2020 examines the mobilizing effect of minority candidates on coethnic support in U.S. congressional elections. The treatment variable indicates the presence of an Asian candidate, and the outcome variable represents the proportion of general election contributions from Asian donors. @HH2019 study the effects of indirect democracy versus direct democracy (treatment) on naturalization rates (outcome) in Switzerland using municipality-year panel data from 1991 to 2009.

First, we load the required packages. The datasets, `hh2019` and `gs2020`, are included with the **fect** package and can be loaded using `data(fect)`.

```{r load, message=FALSE}
# load libraries and data
library(ggplot2)
library(panelView)
library(fect)
data(fect)
ls()
```

## Gap Plot

To create the gap plot, also known as the event study plot, we first apply `fect`, the fixed effects counterfactual estimator. For details, see @sec-fect.

```{r est, cache = TRUE}
out <- fect(Y = "general_sharetotal_A_all", 
            D = "cand_A_all", 
            X = c("cand_H_all", "cand_B_all"), 
            index = c("district_final", "cycle"), 
            data = gs2020, method = "fe", 
            force = "two-way", se = TRUE, 
            parallel = TRUE, nboots = 1000)

out.hh <- fect(nat_rate_ord ~ indirect, 
               data = hh2019,
               index = c("bfs","year"),
               method = 'fe', se = TRUE, 
               parallel = TRUE, nboots = 1000,
               keep.sims = TRUE)
```

After running the model, we can plot the dynamic treatment effects over (relative) time, including confidence intervals if `se = TRUE` is specified in the estimation. Note that `type = "gap"` is the default option, so we omit it here.

```{r}
plot(out) # the effect co-ethnic mobilization
plot(out.hh) # the effect of indirect democrazy on naturalization rate
```

### Starting Period

By default, the first post-treatment period is set to 0, and the last pre-treatment period is set to -1. However, some researchers prefer to designate the former as 1 and the latter as 0. To achieve this, set `start0 = TRUE`.

```{r begin-post-customization}
plot(out, start0 = TRUE,  # Shift time so treatment begins at 0
     main = "Custom Starting Period")
```

### Axis and Legend

Below, we customize the x- and y-axis ranges, labels, tick breaks, and legend. The x-axis labels are rotated for clarity. Moreover, by setting `xlim = c(-10, 1)`, the x-axis is restricted to time periods -8 to 1, with the treatment shifted to begin at period 0 instead of period 1. We also remove grid lines by setting `gridOff = TRUE`.

```{r axis-legend-customization}
plot(out, 
     xlim = c(-10, 1),  # only show time periods -8 to 1
     ylim = c(-0.15, 0.30),  # set y-range
     xlab = "Custom Time Axis", # x-axis label
     ylab = "Estimated ATT", # y-axis label
     xangle = 90, # rotate x-axis labels by 90°
     xbreaks = seq(-10, 1, by = 2), 
     gridOff = TRUE,
     # Label x-axis from -12 to 1 with a break of 2
     main = "Axis and Legend Customization")
```

### Confidence Intervals

Below, we plot the treatment effect with 90% confidence intervals instead of 95%.

```{r ci-raw-customization}
plot(out, plot.ci = "0.9", 
     main = "90% confidence intervals")
```

### Text and ggplot2 Theme

This plot adjusts text sizes with a series of `cex` options and turns off the `theme.bw` option.

```{r text-customization}
plot(out, 
     ylim = c(-0.15, 0.3), # set yrange
     theme.bw = FALSE,      # Change the color theme
     cex.main = 1.25,     # Scale for the main title
     cex.axis = 1.2,      # Axis tick label size
     cex.lab = 1.2,       # Axis label size
     cex.legend = 1,    # Legend text size
     cex.text = 1.2,        # Annotation text size
     main = "Text and Theme Customization")
```

### Presets

For convenience, we can use the `preset` argument to apply preset colors. The default is `"default"`, which is mostly black and white with a bit of color. Other options include `"vibrant"` and `"grayscale"`, which can be used to create more colorful or monochromatic plots, respectively.

```{r preset-vibrant}
plot(out, 
     preset = "vibrant", # Use vibrant colors
     main = "Vibrant Preset Colors: Grumbach and Sahn (2020)")
plot(out.hh, 
     preset = "vibrant", # Use vibrant colors
     main = "Vibrant Preset Colors: Hainmueller and Hangartner (2019)")
```

We can change the color of the estimates (and their confidence intervals) using the `color` option.

```{r preset-vibrant2}
plot(out.hh, 
     preset = "vibrant", # Use vibrant colors
     color = "green4", # Color of the estimates and CIs
     main = "Change Estimates' Color: Hainmueller and Hangartner (2019)")
```

```{r preset-grayscale}
plot(out, 
     preset = "grayscale", # Use grayscale colors
     main = "Grayscale Preset Colors")
```

### Connected Estimates

By default, the estimates are plotted as points. To connect the points with lines, set `connected = TRUE`. The width of the line and size of the points can be adjusted with `est.lwidth` and `est.pointsize`, respectively.

```{r connected-estimates}
plot(out, 
     color = "green4", # color of the estimates and CIs
     connected = TRUE, # Connect the points with lines
     est.lwidth = 1.2, # Makes the lines thicker
     est.pointsize = 3 # Makes the points larger
)
```

Moreover, in any plot that uses a shaded band to represent the CIs, we can set `ci.outline` to `TRUE` to draw an outline around the shaded band to improve visibility.

```{r ci-outline}
plot(out, 
     connected = TRUE, 
     ci.outline = TRUE,
     main = "The Effect of Coethnic Mobilization") # Outline the confidence interval band
plot(out.hh, 
     preset = "vibrant",
     ci.outline = TRUE,
     main = "The Effect of Indirect Democracy") # Outline the confidence interval band
```

### Line and Point Customization

Here, we demonstrate how to change main and the horizontal reference lines, as well as the points and lines in the plot.

```{r line-bound-customization}
plot(out, 
     est.lwidth = 1.5, # Makes the confidence intervals thicker
     est.pointsize = 3, # Makes the points larger
     lcolor = c("red","skyblue"),      # Color for horizontal and vertical lines
     lwidth = 2,                # Widths of the horizontal and vertical lines
     main = "Line Customization")
```

### Count Histogram Customization

The count histogram on the bottom of the graph shows the number of observed units in each period. To customize its color, outline color, and opacity, we can use the `count.color`, `count.outline.color`, and `count.alpha` options, respectively.

```{r count-histogram-customization}
plot(out, 
     count.color = "lightblue", # Color of the histogram bars
     count.outline.color = "darkblue", # Outline color of the histogram bars
     count.alpha = 0.2, # Opacity of the histogram bars
     main = "Count Histogram Customization")
```

## Counterfactual Plot

While the gap plot shows the *difference* (ATT) over time, researchers often want to see the **levels**: the observed outcome for the treated unit(s) and the model‑predicted counterfactual path side‑by‑side. To do this, set `type = "counterfactual"`:

```{r counterfactual}
plot(out, type = "counterfactual",
     main = "Grumbach & Sahn (2020): Treated vs. Counterfactuals",
     ylab = "Proportion of Asian Donation",
     legend.pos = "bottom")
```

```{r counterfactual_hh}
plot(out.hh, type = "counterfactual",
     main = "Hainmueller & Hangartner (2019): Treated vs. Counterfactuals",
     ylab = "Naturalization Rate",
     legend.pos = "top")
```

We can change the color of the lines in this plot using `color`, which sets the color of the main line, and `counterfactual.color`, which sets the color of the counterfactual line (as well as the color of the confidence band but with more transparency). Additionally, we can add an outline to the CIs with `ci.outline = TRUE`.

```{r counterfactual_colors}
plot(out.hh, type = "counterfactual",
     main = "Hainmueller & Hangartner (2019): Treated vs. Counterfactuals",
     ylab = "Naturalization Rate",
     legend.pos = "bottom",
     ci.outline = TRUE, # Outline the confidence interval band
     color = "red3", # Color for the main line
     counterfactual.color = "green4") # Color for the counterfactual line
```

We can also visualize the paths of the individual units by setting `raw = "all"`

```{r counterfactual_rawall}
plot(out, type = "counterfactual", raw = "all")
```

Setting `raw = "band` displays the the 5-95 interpercentile range of the treated and control units. When adoption is staggered, only the band around the treated units is shown.

```{r counterfactual_rawband}
plot(out, type = "counterfactual", raw = "band")
```

We can also change the colors in this plot using the same options as in the gap plot, as well as the `counterfactual.color`, `counterfactual.raw.controls.color`, `counterfactual.raw.treated.color`, and `counterfactual.linetype` options.

```{r counterfactual_colors2}
plot(out, type = "counterfactual",
     count.color = "black", # Color for the count histogram
     count.alpha = 1, # Opacity for the count histogram
     color = "red", # Color for the main line
     counterfactual.color = "purple", # Color for the counterfactual line
     counterfactual.raw.treated.color = "orange", # Color for the treated units
     counterfactual.linetype = "dotted", # Line type for the counterfactual line
     raw = "all", 
     main = "Counterfactual Plot with Custom Colors")
```

## Cumulative Effects

We can also plot cumulative effects by plotting the output of the `effect()` function. Note that this is only well-defined when there are no treatment reversals, that is, all treated units remain treated for the duration of the study. Additionally, we must set `keep.sims = TRUE` to keep the unit-level bootstrap results. We first apply it to `hh2019`.

```{r}
plot(effect(out.hh), main = "Cumulative Effect of Indirect Democracy",
     ylab = "Cumulative Effect on Naturalization Rate")
```

Since the `gs2020` datset has treatment reversals, we will first subset the units that remained treated throughout the study period. We do this by checking for any instances where the treatment variable changes from 1 to 0 within a unit.

```{r}
# flag units that ever have a 1 to 0 change in d
rev_flag <- tapply(gs2020[["cand_A_all"]],
                   gs2020[["district_final"]],
                   function(x) any(diff(x) < 0))

# units with no reversals
good_units <- names(rev_flag)[!rev_flag]

# subset the desired rows
gs2020_no_reversals <- gs2020[gs2020[["district_final"]] %in% good_units, ]

```

Next we will estimate the function again on these units only.

```{r no-reversals-est, cache = TRUE}
out_no_reversals <- fect(Y = "general_sharetotal_A_all", 
                         D = "cand_A_all" , 
                         X = c("cand_H_all", "cand_B_all") ,
                         index = c("district_final", "cycle"), 
                         data = gs2020_no_reversals,
                         method = "fe", 
                         force =  "two-way", 
                         se = TRUE, parallel = TRUE,
                         nboots = 100, 
                         keep.sims = TRUE)
```

Finally, we will plot the cumulative effects.

```{r cumulative-effects}
plot(effect(out_no_reversals), xlim = c(1, 2))
```

## Pretrend Tests

We can conduct several tests to shed light on (not directly test) the parallel trends (PT) assumption, including the equivalence test and the placebo test. For details, see @sec-fect or @LWX2022.

### Equivalence Test

In the equivalence plot (`type = "equiv"`), the equivalence bound is defined by the two-one-sided test (TOST) threshold. For example, in the plot below, the bound is set by `tost.threshold = 0.1`, with lines at -0.1 and 0.1. This threshold should be set based on the magnitude of the ATT or the standard deviation of the outcome (or residualized outcome).

The `bound` option has four choices: `"none"`, `"min"`, `"equiv"`, or `"both"`. When set to `"none"`, no bound is displayed.

```{r}
plot(out, type = "equiv", bound = "equiv", tost.threshold = 0.1, 
     ylim = c(-0.15, 0.15))
```

The `"min"` displays the minimum range bound based on the maximum absolute pre‐treatment residual (e.g., if the largest pre-treatment estimate is 0.3, lines at -0.3 and 0.3).

```{r}
plot(out, type = "equiv", bound = "min", ylim = c(-0.15, 0.15))
```

We can plot both the minimum range and the equivalence bound with `bound = "both"`, which is also the default option.

```{r}
plot(out, type = "equiv", tost.threshold = 0.1, ylim = c(-0.15, 0.15))
```

We use the `stats` argument to select which results to display, label them with `stats.labs`, and position the legend with `stats.pos`. Setting `show.stats = FALSE` hides the test results entirely.

```{r stats-customization}
plot(out, type = "equiv",
     ylim = c(-0.25, 0.25),
     stats = c("F.p", "equiv.p"),
     stats.labs = c("F Test P-value", "Equivalence P-value"),
     stats.pos = c(-8, 0.2),   # (x, y) position for the stats text
     show.stats = TRUE,     # Can be switched off to hide all test stats
     main = "Statistical Test Annotations")
```

### Placebo Test

A placebo test evaluates whether the "fake" ATT is statistically distinguishable in a placebo period. It artificially assigns treatment during placebo periods and estimates the "placebo effect" in those periods.

Therefore, the model must be re-run. Below, we set `placebo.period = c(-2, 0)`, specifying the pre-treatment periods used for the placebo test.

```{r placebo, cache = TRUE}
out_fe_placebo <- fect(Y = "general_sharetotal_A_all", D = "cand_A_all", X = c("cand_H_all", "cand_B_all"), data = gs2020, 
                       index = c("district_final", "cycle"), force = "two-way",
                       method = "fe", CV = FALSE, parallel = TRUE,
                       se = TRUE, nboots = 1000, placeboTest = TRUE,
                       placebo.period = c(-2, 0))

plot(out_fe_placebo)
```

The plot shows the estimated "effects" in the pre-treatment periods (placebo effects). The blue lines in the pre-treatment period suggest that we do not observe significant effects of the treatment in the pre-periods.

We can also change the color of the placebo periods by using the `placebo.color` argument. Colors for many other plot types can also be adjusted in a similar way, including, but not limited to, the carryover and box plots.

```{r}
plot(out_fe_placebo, placebo.color = "green4")
```

## Carryover Effects

One type of plot rarely seen in the empirical literature is how the difference between treatment and control groups evolves after treatment ends. We call it the `"exit"` plot, where the x-axis represents time relative to treatment exit. In contrast, the `"gap"` plot focuses on treatment entry. The `"exit"` plot is essential for assessing potential carryover effects.

```{r}
plot(out_fe_placebo, type = "exit")
```

The test for carryover effects examines whether the treatment effect persists after treatment ends. It artificially labels several post-treatment periods as treated and estimates the "placebo effect" in those periods. By setting `carryover.period = c(1, 3)`, we specify a placebo period that includes three post-treatment periods. If the treatment effect is purely contemporaneous (i.e., there are no carryover effects), the test will not reject the null hypothesis. In this application, the average carryover effect is close to zero and statistically indistinguishable from zero.

```{r carryover, cache = TRUE}
out_fe_carryover <- fect(Y = "general_sharetotal_A_all", D = "cand_A_all", X = c("cand_H_all", "cand_B_all"), data = gs2020, 
                       index = c("district_final", "cycle"), force = "two-way",
                         parallel = TRUE, se = TRUE, CV = FALSE,
                         nboots = 1000, carryoverTest = TRUE,
                         carryover.period = c(1, 3))
plot(out_fe_carryover)
```

## Status Plot

The status plot (`type = "status"`) displays the treatment status by period for all units in a similar fashion to `panelView`. Each of the indicator colors can be customized using the `status.*.color` options.

```{r status}
plot(out_fe_carryover, type = "status",
     status.treat.color      = "#D55E00",  # Color for treated units
     status.control.color    = "#0072B2",  # Color for control units
     status.carryover.color  = "#CC79A7",  # Color for carryover units
     status.missing.color    = "#009E73",  # Color for missing data
     status.background.color = "#F3EAD2",  # Background color
     main = "Status Plot")
```

## Effect Heterogeneity

We provide two ways to visualize treatment effect heterogeneity: a `"box"` plot, which shows the distribution of individual treatment effects, and a `"calendar"` plot, which depicts the ATT conditional on calendar time. We plan to expand this functionality to allow for more pre-treatment covariates soon.

In the box plot, the box in each period represents the range of the middle 50% of the individual effects, while the whiskers show the 2.5%–95% quantiles and the horizontal line represents the median.

```{r}
plot(out, type = "box", xlim = c(-12, 3))
```

In the calendar plot, the blue ribbon represents a loess fit of the conditional ATT, with 95% confidence intervals.

```{r}
plot(out, type = "calendar", main = "The Effect of Coethnic Mobilization")
plot(out.hh, type = "calendar", xlim = c(1995, 2009),
     main = "The Effect of Indirect Democracy")
```
