# Complex Fixed Effects {#sec-cfe}

The **Complex Fixed Effects (CFE)** estimator extends the standard two-way fixed effects counterfactual by incorporating additional model components: extra additive fixed effects, time-invariant covariates with time-varying coefficients, unit-specific loadings on known time trends, and interactive fixed effects (latent factors). Each component helps to relax specific assumptions about the data-generating process, and this chapter introduces them one at a time. R script used in this chapter can be downloaded [here](https://raw.githubusercontent.com/xuyiqing/fect/dev/vignettes/rscript/05-cfe.R).

In [Chapter @sec-ife-mc], we introduced factor-based methods (IFE and MC) that model latent common factors. CFE generalizes this framework by allowing researchers to incorporate additional observed structure alongside latent factors.

```{r .common, include = FALSE}
source("_common.R")
```

```{r setup-cfe, echo = FALSE}
set.seed(1234)
```

```{r load-packages-cfe, message = FALSE, warning = FALSE}
data(simdata)
data(sim_region)
data(sim_linear)
data(sim_trend)
```

------------------------------------------------------------------------

## When to use CFE

The standard IFEct estimator (`method = "ife"`) assumes that the untreated potential outcome is driven by observed time-varying covariates, interactive fixed effects, plus unit and time fixed effects:

$$Y_{it}^{0} = \mu + X_{it}'\beta + \sum_{m=1}^{r} \lambda_{im} f_{tm} + \alpha_i + \xi_t + e_{it}$$

This may be insufficient when:

1.  Units belong to groups with group-level time shocks (additional FEs needed),
2.  Time-invariant characteristics have effects that change over time ($Z'\gamma$),
3.  Units follow known time trends with heterogeneous intensity ($\kappa' Q$),

The CFE estimator models the untreated potential outcome as:

$$Y_{it}^{0} = \mu + X_{it}'\beta + \sum_{m=1}^{r} \lambda_{im} f_{tm} + \alpha_i + \xi_t + \omega_{k(i)}  + Z_i'\gamma_{g(t)} + \kappa_i' Q_t  + e_{it}$$

where:

-   $\mu$ is the grand mean,
-   $X_{it}'\beta$ captures time-varying covariates with constant coefficients,
-   $\lambda_{im} f_{tm}$ are interactive fixed effects (latent factors),
-   $\alpha_i$ and $\xi_t$ are unit and time fixed effects,
-   $\omega_{k(i)}$ are additional group-level fixed effects,
-   $Z_i'\gamma_{g(t)}$ captures time-invariant covariates with coefficients that vary by time group,
-   $\kappa_i' Q_t$ captures unit-specific loadings on known time trends,
-   $e_{it}$ is the idiosyncratic error.

::: {.callout-tip appearance="simple"}
Most of the above elements, except additional group-level fixed effects $\omega_{k(i)}$, can be captured by an interactive fixed effects model. Knowing $Z_i$ or $Q_t$, however, can improve efficiency. In practice, we often do not know the exact functional form of $Q$, and $Z$ may suffer from severe measurement error. Therefore, including these terms may not yield the expected gains and may instead inflate variance.

In general, we recommend a data-driven approach to model selection, such as using `fect_mspe()`, which is based on cross-validation.
:::

### Observed vs. unobserved

| Component | What the researcher provides | What **fect** estimates |
|:----------------|:-----------------------------|:------------------------|
| $X_{it}$ (time-varying covariates) | Covariates in the formula (right-hand side) | $\beta$ --- constant coefficients |
| $Z_i$ (time-invariant covariates) | Column names via `Z` argument | $\gamma_{g(t)}$ --- coefficients that vary by time group |
| $Q_t$ (known time basis) | Either column names via `Q`, or auto-generated via `Q.type` | $\kappa_i$ --- unit-specific loadings |
| Group indicators | Extra elements in `index` | $\omega_k$ --- group fixed effects |
| Latent factors | Number of factors via `r` | Both $\lambda_i$ (loadings) and $f_t$ (factors) |

### Algorithm

The model is estimated via an EM-style block coordinate descent algorithm. The algorithm iterates between imputing missing or treated entries with current fitted values, updating each component block --- covariates ($\beta$), time-invariant covariate coefficients ($\gamma$), unit-specific trend loadings ($\kappa$), additive fixed effects ($\alpha$, $\xi$, $\omega$), and interactive fixed effects via SVD ($\lambda$, $f$) --- and checking convergence. The entire procedure is implemented in C++ for computational efficiency.

::: {.callout-tip appearance="simple"}
The CFE estimator is activated by `method = "cfe"`. The key arguments that control the model components are: `index` (additional FEs), `Z` and `gamma` (time-invariant covariates), `Q.type` or `Q` and `kappa` (time trends), and `r` (latent factors). The rest of this chapter introduces each component individually.
:::

------------------------------------------------------------------------

## Additional fixed effects

When units belong to groups --- such as regions, industry sectors, or cohorts --- group-level time shocks can affect outcomes and correlate with treatment assignment. In unbalanced panels where units enter or exit at different times, the group structure becomes especially important: without accounting for group effects, the counterfactual imputation is biased and placebo tests fail.

### Data-generating process

The untreated potential outcome in this DGP follows:

$$Y_{it}^{0} = \alpha_i + \xi_t + \delta_{g(i),t} + e_{it}$$

where $\alpha_i$ is a unit fixed effect, $\xi_t$ is a time fixed effect, and $\delta_{g(i),t}$ is a region-specific time effect that serves as the confounding source. Because treatment assignment and timing both depend on region, $\delta_{g(i),t}$ correlates with treatment status --- creating exactly the kind of bias that additional fixed effects are designed to absorb.

We generate an unbalanced panel with 300 units belonging to 5 regions. Each region has a distinct linear time trend, and treatment probability and timing both depend on region. Units in higher-numbered regions enter the panel later, creating the kind of unbalancedness common in real datasets (e.g., firms entering markets, countries joining surveys).

The `sim_region` dataset ships with the package. It is an unbalanced panel with 500 units belonging to 5 regions. The DGP generation script is in `data-raw/sim_region.R`.

```{r cfe-42-load, eval = TRUE}
head(sim_region)
```

### Without additional fixed effects

We first try the standard FE estimator, which ignores the region-level shocks:

```{r cfe-42-fe-only, eval = TRUE, cache = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
out.fe.only <- fect(Y ~ D, data = sim_region,
  index = c("id", "time"),
  method = "fe", force = "two-way",
  se = TRUE, parallel = TRUE, cores = 16, nboots = 200,
  placeboTest = TRUE, placebo.period = c(-2, 0))
```

```{r cfe-42-fe-only-plot, fig.width = 6, fig.height = 4.5}
plot(out.fe.only, 
     stats = c("placebo.p", "equiv.p"),
     main = "FE Only — Placebo Test")
```

The placebo test detects significant pre-trends (low p-value) because the model fails to account for region-level confounding. In the unbalanced panel, group-level time trends $\delta_{g(i),t}$ interact with the differential entry times across regions to create spurious pre-trends that the standard two-way fixed effects model cannot absorb.

### With additional fixed effects

Now we interact region with time to create group×period fixed effects, which absorb the region-specific time shocks $\delta_{g(i),t}$. A simple region intercept FE cannot capture these shocks because they vary across both regions and time periods. By passing `region_time` as the third index element, the CFE estimator absorbs the full set of region-by-period effects:

```{r cfe-42-with-region, eval = TRUE, cache = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
out.cfe.region <- fect(Y ~ D, data = sim_region,
  index = c("id", "time", "region_time"),
  method = "cfe", force = "two-way",
  se = TRUE, parallel = TRUE, cores = 16, nboots = 200,
  placeboTest = TRUE, placebo.period = c(-2, 0))
```

```{r cfe-42-with-region-plot, fig.width = 6, fig.height = 4.5}
plot(out.cfe.region, 
     stats = c("placebo.p", "equiv.p"),
     main = "CFE with Region×Time FE — Placebo Test")
```

The placebo test now passes (high p-value), confirming that the region×time fixed effects absorb the group-level confounding. By including the interacted `region_time` variable in the `index` argument, the CFE estimator absorbs the full set of region-specific time shocks $\delta_{g(i),t}$ during counterfactual imputation, removing the bias that was driving the spurious pre-trends.

::: {.callout-note appearance="simple"}
In the `index` argument, elements beyond the first two (unit, time) are treated as additional fixed-effect grouping variables. You can include multiple additional groupings: `index = c("id", "time", "region_time", "sector_time")`.
:::

------------------------------------------------------------------------

## Time-invariant covariates with time-varying coefficients

When time-invariant unit characteristics $Z_i$ (e.g., baseline GDP, initial population) have effects that change over time, a simple additive control is insufficient. The CFE model allows $Z_i'\gamma_{g(t)}$, where $\gamma$ varies by time group, capturing the time-varying nature of these effects.

### The challenge with interactive structures

We use the existing `simdata` dataset. In the true DGP, $Y_{it}^{0}$ includes the interactive term $L_{1,i} \cdot F_{1,t}$ --- a product of the unit-level factor loading $L_1$ and the time-varying factor $F_1$. If we observe $L_1$ as a covariate but estimate it with a constant coefficient, we miss the time-varying nature of its effect. The CFE $Z/\gamma$ mechanism handles this.

First, we estimate the FE model, which ignores the interactive structure entirely:

```{r cfe-43-fe-baseline, eval = TRUE, cache = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
out.fe.base <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time"),
  method = "fe", force = "two-way",
  se = TRUE, parallel = TRUE, cores = 16, nboots = 200,
  placeboTest = TRUE, placebo.period = c(-2, 0))
```

```{r cfe-43-fe-baseline-plot, fig.width = 6, fig.height = 4.5}
plot(out.fe.base, 
     stats = c("placebo.p", "equiv.p"),
     main = "FE Only (simdata) — Placebo Test")
```

The FE estimator fails the placebo test because it cannot account for the interactive term $L_1 \cdot F_1$. Now we include $L_1$ as a time-invariant covariate with time-varying coefficients. We create a `gamma` variable --- here we use `time` itself as the grouping variable, which gives the most flexible specification (one coefficient per period):

```{r cfe-43-gamma-setup, eval = TRUE}
simdata$gamma_t <- simdata$time
```

```{r cfe-43-with-z, eval = TRUE, cache = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
out.cfe.z <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time"),
  method = "cfe", force = "two-way",
  Z = "L1", gamma = "gamma_t",
  se = TRUE, parallel = TRUE, cores = 16, nboots = 200,
  placeboTest = TRUE, placebo.period = c(-2, 0))
```

```{r cfe-43-with-z-plot, fig.width = 6, fig.height = 4.5}
plot(out.cfe.z, 
     stats = c("placebo.p", "equiv.p"),
     main = "CFE with Z = L1 — Placebo Test")
```

By allowing $L_1$'s coefficient to vary across time periods, CFE approximates the true interactive structure $L_{1,i} \cdot F_{1,t}$. The placebo test shows improvement. However, there is still a second latent factor $L_2 \cdot F_2$ that is not captured --- we will address this in Section 4.5.

::: {.callout-note appearance="simple"}
The `Z` argument takes a character vector of column names for time-invariant covariates. The `gamma` argument specifies a column that defines the time grouping for the $Z$ coefficients. Using `time` as the gamma variable gives the most flexible specification; coarser groupings (e.g., early/middle/late) reduce the number of parameters.
:::

------------------------------------------------------------------------

## Unit-specific time trends

When all units follow the same temporal pattern but the amplitude differs across units --- and units with stronger trends are more likely to be treated --- the counterfactual imputation from a standard two-way FE model is biased. The CFE estimator can model unit-specific loadings on known time bases via the `Q.type` argument. Options include `"linear"`, `"quadratic"`, `"cubic"`, and `"bspline"`. We illustrate with two examples of increasing complexity.

### Linear time trends

The simplest case is a linear time trend with unit-specific slopes. We generate a block DID panel with 200 units (80 treated, 120 control), where treated units have systematically steeper linear trends.

The `sim_linear` dataset ships with the package. It has 200 units (80 treated, 120 control), 50 time periods, with block treatment at period 41. The DGP generation script is in `data-raw/sim_linear.R`.

```{r cfe-44-linear-load, eval = TRUE}
head(sim_linear)
```

Without accounting for the unit-specific linear trends, the FE estimator fails the placebo test:

```{r cfe-44-lin-fe-only, eval = TRUE, cache = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
out.fe.lin <- fect(Y ~ D, data = sim_linear,
  index = c("id", "time"),
  method = "fe", force = "two-way",
  se = TRUE, parallel = TRUE, cores = 16, nboots = 200,
  placeboTest = TRUE, placebo.period = c(-2, 0))
```

```{r cfe-44-lin-fe-only-plot, fig.width = 6, fig.height = 4.5}
plot(out.fe.lin, 
     stats = c("placebo.p", "equiv.p"),
     main = "FE Only (Linear Trend DGP) — Placebo Test")
```

Now we use `Q.type = "linear"` to allow unit-specific loadings on a linear time basis. Because the true DGP is exactly linear, this specification matches perfectly:

```{r cfe-44-lin-cfe, eval = TRUE, cache = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
out.cfe.lin <- fect(Y ~ D, data = sim_linear,
  index = c("id", "time"),
  method = "cfe", force = "two-way",
  Q.type = "linear",
  se = TRUE, parallel = TRUE, cores = 16, nboots = 200,
  placeboTest = TRUE, placebo.period = c(-2, 0))
```

```{r cfe-44-lin-cfe-plot, fig.width = 6, fig.height = 4.5}
plot(out.cfe.lin, 
     stats = c("placebo.p", "equiv.p"),
     main = "CFE with Linear Trend — Placebo Test")
```

The placebo test passes because the linear basis exactly matches the true trend shape, and the CFE estimator correctly recovers each unit's slope $\kappa_i$.

### Nonlinear time trends

When the true trend is nonlinear, a linear basis is insufficient. We now generate a DGP with a sinusoidal time trend --- a half-cycle over the panel --- and unit-specific amplitudes:

The `sim_trend` dataset ships with the package. It has the same block DID structure as `sim_linear` but with a sinusoidal (half-cycle) time trend instead of a linear one. The DGP generation script is in `data-raw/sim_trend.R`.

```{r cfe-44-sin-load, eval = TRUE}
head(sim_trend)
```

The FE estimator fails the placebo test because it cannot capture the nonlinear trend:

```{r cfe-44-sin-fe-only, eval = TRUE, cache = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
out.fe.trend <- fect(Y ~ D, data = sim_trend,
  index = c("id", "time"),
  method = "fe", force = "two-way",
  se = TRUE, parallel = TRUE, cores = 16, nboots = 200,
  placeboTest = TRUE, placebo.period = c(-2, 0))
```

```{r cfe-44-sin-fe-only-plot, fig.width = 6, fig.height = 4.5}
plot(out.fe.trend, 
     stats = c("placebo.p", "equiv.p"),
     main = "FE Only (Sin Trend DGP) — Placebo Test")
```

Now we use `Q.type = "bspline"`, which generates a B-spline basis that can approximate smooth nonlinear functions:

```{r cfe-44-sin-bspline, eval = TRUE, cache = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
out.cfe.bs <- fect(Y ~ D, data = sim_trend,
  index = c("id", "time"),
  method = "cfe", force = "two-way",
  Q.type = "bspline",
  se = TRUE, parallel = TRUE, cores = 16, nboots = 200,
  placeboTest = TRUE, placebo.period = c(-2, 0))
```

```{r cfe-44-sin-bspline-plot, fig.width = 6, fig.height = 4.5}
plot(out.cfe.bs, 
     stats = c("placebo.p", "equiv.p"),
     main = "CFE with B-spline Trend — Placebo Test")
```

The B-spline specification should pass or nearly pass the placebo test. Some bias may remain because B-splines approximate the true sinusoidal trend rather than matching it exactly.

::: {.callout-note appearance="simple"}
The `Q.type` argument accepts `"linear"`, `"quadratic"`, `"cubic"`, or `"bspline"`. Multiple types can be combined: `Q.type = c("linear", "quadratic")`. For B-splines, `Q.bspline.degree` controls the polynomial degree (default chosen automatically based on the number of time periods).
:::

------------------------------------------------------------------------

## CFE with factors

Sometimes the data contains both observed structure (known covariates whose effects vary over time) and unobserved latent factors. The CFE estimator can combine $Z'\gamma$ and interactive fixed effects $\lambda' f$ in a single model.

We use `simdata`, which has two latent factors ($L_1, F_1$ and $L_2, F_2$). We pretend $L_1$ is an observed time-invariant covariate and model the remaining factor structure with $r = 1$ latent factor.

```{r cfe-45-gamma-setup, eval = TRUE}
simdata$gamma_t <- simdata$time
```

### Model comparison via `fect_mspe`

We fit four models and compare their out-of-sample prediction accuracy:

1.  FE only (baseline),
2.  CFE with $Z = L_1$ only (one observed loading, no factors),
3.  CFE with $Z = L_1$ + 1 latent factor (correct specification),
4.  IFE with 2 latent factors (correct for IFE, but less efficient than CFE).

```{r cfe-45-fit-models, eval = TRUE, cache = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
# Model 1: FE only
out.fe <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time"),
  method = "fe", force = "two-way", se = FALSE)

# Model 2: CFE with Z = L1 only
out.cfe.z.only <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time"),
  method = "cfe", force = "two-way",
  Z = "L1", gamma = "gamma_t",
  se = FALSE, max.iteration = 20000)

# Model 3: CFE with Z = L1 + 1 factor
out.cfe.z.f1 <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time"),
  method = "cfe", force = "two-way",
  Z = "L1", gamma = "gamma_t",
  r = 1, se = FALSE, max.iteration = 20000)

# Model 4: IFE with 2 factors
out.ife.r2 <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time"),
  method = "ife", force = "two-way",
  r = 2, se = FALSE, max.iteration = 20000)
```

```{r cfe-45-mspe, eval = TRUE, cache = FALSE}
mspe.out <- fect_mspe(list(FE = out.fe,
       CFE_Z = out.cfe.z.only,
       CFE_Z_F1 = out.cfe.z.f1,
       IFE_r2 = out.ife.r2),
  seed = 1234)
print(mspe.out$summary[, c("Model", "MSPE", "RMSE", "MAD")])
```

All three metrics consistently rank the CFE model with observed $Z$ and one latent factor as the best specification. CFE with $Z + 1$ factor has the lowest MSPE because it uses observed information ($L_1$) efficiently and estimates only the remaining unobserved factor ($L_2$). The IFE with $r=2$ should also do well but slightly worse, since it estimates both factors entirely from the data.

### Best model: placebo test

```{r cfe-45-best-placebo, eval = TRUE, cache = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
out.cfe.best <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time"),
  method = "cfe", force = "two-way",
  Z = "L1", gamma = "gamma_t",
  r = 1,
  se = TRUE, parallel = TRUE, cores = 16, nboots = 200,
  placeboTest = TRUE, placebo.period = c(-2, 0))
```

```{r cfe-45-best-placebo-plot, fig.width = 6, fig.height = 4.5}
plot(out.cfe.best, 
     stats = c("placebo.p", "equiv.p"),
     main = "CFE (Z + 1 Factor) — Placebo Test")
```

The placebo test passes, confirming that the CFE model with one observed loading and one latent factor adequately captures the DGP.

::: {.callout-note appearance="simple"}
`fect_mspe()` compares out-of-sample prediction accuracy across different model specifications. A lower MSPE indicates better counterfactual prediction. This is useful for selecting among CFE configurations when the true model is unknown.
:::

------------------------------------------------------------------------

## Grouped coefficients

In some applications, the $Z$-$\gamma$ or $\kappa$-$Q$ coefficients should not all share the same grouping structure. For example:

-   **`Z.param`**: Suppose you have two time-invariant covariates --- baseline GDP and initial population --- and their effects change at different rates. GDP effects may shift by decade, while population effects shift by political era. `Z.param` allows you to assign different gamma groupings to different $Z$ variables.
-   **`Q.param`**: Similarly, if you specify multiple $Q$ time bases, `Q.param` controls which unit-grouping variable is used for each $Q$ basis.

### `Z.param` syntax

```{r cfe-46-zparam-example, eval = FALSE}
# Example with Z.param (not run — requires appropriate data)
# out <- fect(Y ~ D, data = mydata,
#   index = c("unit", "time"),
#   method = "cfe", force = "two-way",
#   Z = c("baseline_gdp", "baseline_pop"),
#   gamma = c("decade", "political_era"),
#   Z.param = list(decade = "baseline_gdp",
#                  political_era = "baseline_pop"))
```

This means: `baseline_gdp` uses `decade` for its time grouping, while `baseline_pop` uses `political_era`.

### `Q.param` syntax

```         
Q.param = list(sector = c("Q_linear", "Q_quadratic"), region = "Q_spline")
```

This means: the linear and quadratic trend bases are grouped by sector, while the spline basis is grouped by region.

### When to use grouped coefficients

-   Use `Z.param` when different covariates' effects change at different timescales.
-   Use `Q.param` when different trend components should vary by different unit groupings.
-   Without `Z.param`/`Q.param`, all $Z$ variables share all gamma groupings (fully crossed), and all $Q$ variables share all kappa groupings --- this may overparameterize the model.

::: {.callout-note appearance="simple"}
`Z.param` and `Q.param` are optional. Without them, all $Z$ (or $Q$) variables are assigned to all gamma (or kappa) groupings. Use them to impose structure when you have domain knowledge about which covariates should share which groupings.
:::

------------------------------------------------------------------------

## Cross-validation

The number of latent factors `r` in a CFE specification can be selected by cross-validation through the main `fect()` dispatcher --- set `CV = TRUE` and pass `cv.method = "rolling"` to use rolling-window CV (recommended on macro panels with serially correlated residuals; see [Chapter @sec-ife-mc]). CFE-specific structural arguments (`Z`, `gamma`, `Q`, `Q.type`, `kappa`, plus extra index columns for additional grouping FEs) are held fixed at the user-supplied spec; CV varies only `r`.

```{r cfe-cv-dispatcher, eval = FALSE}
fit.cfe <- fect(Y ~ D, data = sim_region,
  index   = c("id", "time", "region_time"),
  method  = "cfe", force = "two-way",
  CV      = TRUE, r = c(0, 3),
  cv.method = "rolling",
  cv.buffer = 1, cv.nobs = 3, k = 20, cv.prop = 0.1,
  cv.rule = "1se",
  se = TRUE, parallel = TRUE, cores = 16, nboots = 200
)
fit.cfe$r.cv  # selected r
```

Block CV (`cv.method = "block"`) is also available as a drop-in replacement; see the [Cross-Validation section of the cheatsheet](aa-cheatsheet.html#sec-cheatsheet-cv) for the full parameter reference and the deprecation plan for the legacy `"all_units"` / `"treated_units"` values.

------------------------------------------------------------------------

## Summary of CFE-specific arguments

| Argument | Type | Description |
|:-----------------------|:----------------|:------------------------------|
| `method = "cfe"` | character | Activates the CFE estimator |
| `index` | character vector | `c(unit, time, ...)` --- elements beyond the first two are extra additive FEs |
| `Z` | character vector | Names of time-invariant covariate columns |
| `gamma` | character vector | Names of columns defining time grouping for $Z$ coefficients |
| `Q` | character vector | Names of known time trend columns (user-supplied) |
| `kappa` | character vector | Names of columns defining unit/group assignment for $Q$ loadings |
| `Q.type` | character vector | Auto-generate $Q$: `"linear"`, `"quadratic"`, `"cubic"`, `"bspline"` |
| `Q.bspline.degree` | integer | Degree of B-spline when `Q.type = "bspline"` (default: auto) |
| `Z.param` | named list | Block assignment: which $Z$ columns share which $\gamma$ grouping |
| `Q.param` | named list | Block assignment: which $Q$ columns share which $\kappa$ grouping |
| `r` | integer or vector | Number of latent factors; use `c(0, 5)` with `CV = TRUE` to select |
| `force` | character | Additive FE: `"none"`, `"unit"`, `"time"`, `"two-way"` |

The CFE estimator combines all these components into a single unified model. Each component relaxes a specific assumption about the data-generating process. In practice, researchers should use `fect_mspe()` (Section 4.5) to compare specifications and placebo tests to validate the chosen model.
