# Diagnostics {#sec-diagnostics}

This chapter covers diagnostic tests in **fect**: placebo tests, pre-trend tests (F test and equivalence/TOST), leave-one-out pre-trend tests, exit-treatment plots, carryover effect tests, and cumulative effect estimation. These tools help assess whether the identifying assumptions are plausible for a given model and dataset.

The fitted objects `out.fect`, `out.ife`, and `out.mc` from earlier chapters are re-created below so that this chapter can be rendered independently.

```{r diagnostics_setup, eval=TRUE, cache=TRUE, results='hide', message=FALSE}
library(fect)
data(fect)

out.fect <- fect(Y ~ D + X1 + X2, data = simdata2, index = c("id","time"),
  method = "fe", force = "two-way", se = TRUE,
  cores = 8, parallel = TRUE, nboots = 1000)

out.ife <- fect(Y ~ D + X1 + X2, data = simdata2, index = c("id","time"),
          force = "two-way", method = "ife", CV = TRUE, r = c(0, 5),
          se = TRUE, cores = 8, nboots = 1000, parallel = TRUE)

out.mc <- fect(Y ~ D + X1 + X2, data = simdata2, index = c("id","time"),
          force = "two-way", method = "mc", CV = TRUE,
          se = TRUE, cores = 8, nboots = 1000, parallel = TRUE)
```

------------------------------------------------------------------------

## Diagnostic tests

We provide three types of diagnostic tests: (1) a placebo test, (2) a test for (no) pretrend, and (3) a test for (no) carry-over effects. For each test, we support both the difference-in-means (DIM) approach and the equivalence approach. The details are provided in the paper.

### Placebo tests

We provide a placebo test for a settled model---hence, cross-validation is not allowed---by setting `placeboTest = TRUE`. We specify a range of pretreatment periods as "placebo periods" in option `placebo.period` to remove observations in the specified range for model fitting, and then test whether the estimated ATT in this range is significantly different from zero. Below, we set `c(-2, 0)` as the placebo periods.

```{r placebo, eval = TRUE, cache = TRUE, message = FALSE, results='hide'}
out.fect.p <- fect(Y ~ D + X1 + X2, data = simdata2, index = c("id", "time"),
  force = "two-way", parallel = TRUE, se = TRUE, CV = 0,
  nboots = 200, placeboTest = TRUE, placebo.period = c(-2, 0))

out.ife.p <- fect(Y ~ D + X1 + X2, data = simdata2, index = c("id", "time"),
  force = "two-way", method = "ife",  r = 2, CV = 0,
  parallel = TRUE, se = TRUE,
  nboots = 200, placeboTest = TRUE, placebo.period = c(-2, 0))

out.mc.p <- fect(Y ~ D + X1 + X2, data = simdata2, index = c("id", "time"),
  force = "two-way", method = "mc",  lambda = out.mc$lambda.cv,
  CV = 0, parallel = TRUE, se = TRUE,
  nboots = 200, placeboTest = TRUE, placebo.period = c(-2, 0))
```

The placebo test conducts two types of tests:

**t test.** If t-test p-value is smaller than a pre-specified threshold (e.g. 5%), we reject the null of no-differences. Hence, the placebo test is deemed failed.

**TOST.** The TOST checks whether the 90% confidence intervals for estimated ATTs in the placebo period exceed a pre-specified range (defined by a threshold), or the equivalence range. A TOST p-value smaller than a pre-specified threshold suggests that the null of difference bigger than the threshold is rejected; hence, the placebo test is passed.

By default, the plot will display the p-value of the $t$-test (`stats = "placebo.p"`). Users can also add the p-value of a corresponding TOST test by setting `stats = c("placebo.p","equiv.p")`. A larger placebo p-value from a t-test and a smaller placebo TOST p-value are preferred.

```{r placebo_1, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.fect.p, cex.text = 0.8, stats = c("placebo.p","equiv.p"),
     main = "Estimated ATT (TWFE)")
```

```{r placebo_2, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.ife.p, ylab = "Effect of D on Y", main = "Estimated ATT (IFE)",
     cex.text = 0.8, stats = c("placebo.p","equiv.p"))
```

```{r placebo_3, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.mc.p, cex.text = 0.8, stats = c("placebo.p","equiv.p"),
     main = "Estimated ATT (MC)")
```

The results in the placebo test confirm that IFEct is a better model than MC for this particular DGP.

### Tests for (no) pre-trend

We introduce two statistical tests for the presence of a pre-trend (or the lack thereof). The first test is an $F$ test for zero residual averages in the pretreatment periods. The second test is a two-one-sided $t$ (TOST) test, a type of equivalence tests.

**F test.** We offer a goodness-of-fit test (a variant of the $F$ test) and to gauge the presence of pretreatment (differential) trends. A larger F-test p-value suggests a better pre-trend fitting. Users can specify a test range in option `pre.periods`. For example, `pre.periods = c(-4,0)` means that we test pretreatment trend of the last 5 periods prior to the treatment (from period -4 to period 0). If `pre.period = NULL` (default), all pretreatment periods in which the number of treated units exceeds the total number of treated units \* `proportion` will be included in the test.

**TOST.** The TOST checks whether the 90% confidence intervals for estimated ATTs in the pretreatment periods (again, subject to the `proportion` option) exceed a pre-specified range, or the equivalence range. A smaller TOST p-value suggests a better pre-trend fitting. While users can check the values of confidence intervals, we give a visualization of the equivalence test. We can plot the pretreatment residual average with the equivalence confidence intervals by setting `type = "equiv"`. Option `tost.threshold` sets the equivalence range (the default is $0.36\sigma_{\epsilon}$ in which $\sigma_{\epsilon}$ is the standard deviation of the outcome variable after two-way fixed effects are partialed out). By setting `range = "both"`, both the minimum range (in gray) and the equivalence range (in red) are drawn.

On the topleft corner of the graph, we show several statistics of the user's choice. User can choose which statistics to show by setting `stats = c("none", "F.stat", "F.p", "F.equiv.p", "equiv.p")` which corresponds to not showing any, the $F$ statistic, the p-value for the $F$ test, the p-value for the equivalence $F$ test, the (maximum) p-value for the the TOST tests, respectively. For the gap plot, the default is `stats = "none"`. For the equivalence plot, the default is `stats = c("equiv.p, F.p")`. Users can also change the labels of statistics using the `stats.labs` options. Users can adjust its position using the `stats.pos` option, for example `stats.pos = c(-30, 4)`. To turn off the statistics, set `stats = "none"`.

Below, we visualize the result of the equivalence test for each of the three estimators using our simulated data. These figures show that both the IFE and MC methods pass the equivalence test while the FE method does not.

```{r pretrend1, eval = TRUE, cache = TRUE, fig.width = 6, fig.height = 4.5}
plot(out.fect, type = "equiv", ylim = c(-4,4),
     cex.legend = 0.6, main = "Testing Pre-Trend (FEct)", cex.text = 0.8)
```

```{r pretrend2, eval = TRUE, cache = TRUE, fig.width = 6, fig.height = 4.5}
plot(out.ife, type = "equiv", ylim = c(-4,4),
     cex.legend = 0.6, main = "Testing Pre-Trend (IFEct)", cex.text = 0.8)
```

```{r pretrend3, eval = TRUE, cache = TRUE, fig.width = 6, fig.height = 4.5}
plot(out.mc, type = "equiv", ylim = c(-4,4),
     cex.legend = 0.6, main = "Testing Pre-Trend (MC)", cex.text = 0.8)
```

From the above plots, we see that FEct fails both tests; IFEct passes both tests using a conventional test size (5%); and MC fails the F tests, but passes the TOST (equivalence) test. Hence, we may conclude that IFEct is a more suitable model.

### LOO pre-trend test {#loo-pre-trend-test}

Instead of using estimated ATTs for periods prior to the treatment to test for pre-trends, we recommend users employ a leave-one-out (LOO) approach (`loo = TRUE`) to consecutively hide one pretreatment period (relative to the timing of the treatment) and repeatedly estimate the pseudo treatment effects for that pretreatment period. The LOO approach can be understood as an extension of the placebo test. It has the benefit of providing users with a more holistic view of whether the identifying assumptions likely hold. However, as the program needs to conduct uncertainty estimates for each turn, it is much more time-consuming than the original one.

```{r simdata_fect_loo, eval=FALSE, cache = TRUE, message = FALSE, results = 'hide'}
out.fect.loo <- fect(Y ~ D + X1 + X2, data = simdata2, index = c("id","time"),
  method = "fe", force = "two-way", se = TRUE, parallel = TRUE, nboots = 200, loo = TRUE)
out.ife.loo <- fect(Y ~ D + X1 + X2, data = simdata2, index = c("id","time"),
  method = "ife", force = "two-way", se = TRUE, parallel = TRUE, nboots = 200, loo = TRUE)
out.mc.loo <- fect(Y ~ D + X1 + X2, data = simdata2, index = c("id","time"),
  method = "mc", force = "two-way", se = TRUE, parallel = TRUE, nboots = 200, loo = TRUE)
```

After the LOO estimation, one can plot these LOO pre-trends in the gap plot or the equivalence plot by setting `loo = TRUE` in the `plot` function.

```{r pretrend_loo1, eval = FALSE, cache = TRUE, fig.width = 6, fig.height = 4.5}
plot(out.fect.loo, type = "equiv", ylim = c(-4,4), loo = TRUE,
     cex.legend = 0.6, main = "Testing Pre-Trend LOO (FEct)", cex.text = 0.8)

plot(out.ife.loo, type = "equiv", ylim = c(-4,4), loo = TRUE,
     cex.legend = 0.6, main = "Testing Pre-Trend LOO (IFEct)", cex.text = 0.8)

plot(out.mc.loo, type = "equiv", ylim = c(-4,4), loo = TRUE,
     cex.legend = 0.6, main = "Testing Pre-Trend LOO (MC)", cex.text = 0.8)
```

Note that the LOO test usually takes lots of computational power. For our example, we find that the IFE estimator still passes both the F test and the equivalence test based on its LOO pre-trends, while the MC estimator fails both tests.

### Exiting the treatment

**fect** allows the treatment to switch back and forth and provides diagnostic tools for this setting. After the estimation, we can visualize the period-wise ATTs relative to the `exit` of treatments by setting `type = "exit"` (one can still draw the classic gap plot by setting `type = "gap"`). The x-axis is then realigned based on the timing of the treatment's exit, not onset, e.g., 1 represents 1 period after the treatment ends.

```{r exit_1, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.fect, type = "exit", ylim = c(-2.5,4.5), main = "What Happens after the Treatment Switches Off?")
```

```{r exit_2, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.ife, type = "exit", ylim = c(-2.5,4.5), main = "Exit Plot (IFE)")
```

```{r exit_3, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.mc, type = "exit", ylim = c(-2.5,4.5), main = "Exit Plot (MC)")
```

### Tests for (no) carryover effects

The idea of the placebo test can be extended to testing the presence of carryover effects. Instead of hiding a few periods right before the treatment starts, we hide a few periods right after the treatment ends. If carryover effects do not exist, we would expect the average prediction error in those periods to be close to zero. To perform the carryover test, we set the option `carryoverTest = TRUE`. We can treat a range of exit-treatment periods in option `carryover.period` to remove observations in the specified range for model fitting, and then test whether the estimated ATT in this range is significantly different from zero.

Below, we set `carryover.period = c(1, 3)`. As we deduct the treatment effect from the outcome in `simdata2`, we expect the average prediction error for these removed periods to be close to zero.

```{r carryover, eval = TRUE, cache = TRUE, message = FALSE, results='hide'}
out.fect.c <- fect(Y ~ D + X1 + X2, data = simdata2, index = c("id", "time"),
  force = "two-way", parallel = TRUE, se = TRUE, CV = 0,
  nboots = 200, carryoverTest = TRUE, carryover.period = c(1, 3))

out.ife.c <- fect(Y ~ D + X1 + X2, data = simdata2, index = c("id", "time"),
  force = "two-way", method = "ife", r = 2, CV = 0,
  parallel = TRUE, se = TRUE,
  nboots = 200, carryoverTest = TRUE, carryover.period = c(1, 3))

out.mc.c <- fect(Y ~ D + X1 + X2, data = simdata2, index = c("id", "time"),
  force = "two-way", method = "mc",  lambda = out.mc$lambda.cv,
  CV = 0, parallel = TRUE, se = TRUE,
  nboots = 200, carryoverTest = TRUE, carryover.period = c(1, 3))
```

Like the placebo test, the plot will display the p-value of the carryover effect test (`stats = "carryover.p"`). Users can also add the p-value of a corresponding TOST test by setting `stats = c("carryover.p","equiv.p")`.

```{r carryover1, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.fect.c, type = "exit", ylim = c(-2.5,4.5),
          cex.text = 0.8, main = "Carryover Effects (FE)")
```

```{r carryover2, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.ife.c, type = "exit", ylim = c(-2.5,4.5),
          cex.text = 0.8, main = "Carryover Effects (IFE)")
```

```{r carryover3, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.mc.c, type = "exit", ylim = c(-2.5,4.5),
          cex.text = 0.8, main = "Carryover Effects (MC)")
```

Once again, the IFE estimator outperforms the other two.

Using real-world data, researchers will likely find that carryover effects exist. If such effects are limited, researchers can consider removing a few periods after the treatment ended for the treated units from the first-stage estimation (using the `carryover.period` option) and re-estimated the model (and re-conduct the test). We provide such an example in the paper. Here, we illustrate the option using `simdata2`.

```{r carryover_rm, eval = TRUE, cache = TRUE, message = FALSE, results='hide', fig.width = 6, fig.height = 4.5}
out.ife.rm.test <- fect(Y ~ D + X1 + X2, data = simdata2, index = c("id", "time"),
  force = "two-way", method = "ife", r = 2, CV = 0,
  parallel = TRUE, se = TRUE,  carryover.rm = 3,
  nboots = 200, carryoverTest = TRUE, carryover.period = c(1, 3))# remove three periods

plot(out.ife.rm.test, cex.text = 0.8, stats.pos = c(5, 2))
```

In the above plot, the three periods in blue are droppred from the first-stage estimation of the factor model while the periods in red are reserved for the (no) carryover effects test.

------------------------------------------------------------------------

## Cumulative effects

Users can use `effect()` to calculate cumulative treatment effects. The behavior of `effect()` is similar to the function of the same name in `gsynth`. - Calculation of cumulative effects will need unit-time level bootstrap results. Choose the option `keep.sims=TRUE` to record them.

::: {.callout-note appearance="simple"}
The example below uses `method = "ife"` with `factors.from = "nevertreated"`, which is equivalent to `method = "gsynth"` (see [Chapter @sec-gsynth]). We use the explicit form here for clarity.
:::

```{r effect, cache = TRUE}
out <- fect(Y ~ D + X1 + X2, data = simgsynth, index = c("id","time"),
                        method = "ife", factors.from = "nevertreated",
                        force = "two-way", CV = TRUE, r = c(0, 5),
                        se = TRUE, nboots = 200, vartype = 'bootstrap',
                        parallel = FALSE, keep.sims=TRUE)
cumu.out <- effect(out)
```

Print and plot cumulative effects

```{r effect.plot, cache = TRUE}
print(cumu.out)
plot(cumu.out)
```

Users can choose to calculate by-period average effects by setting `cumu=FALSE`.

```{r}
effect(out, cumu=FALSE)
```

Calculate the cumulative effect of certain units at certain periods.

```{r}
effect(out, cumu=TRUE, id=c(101,102,103), period=c(1,5))
```

`effect()` also accepts results from other estimation and inference methods. For example, we can use matrix completion:

```{r effect-mc, cache = TRUE}
out_mc <- fect(Y ~ D + X1 + X2, data = simgsynth, index = c("id","time"),
                        method = "mc", force = "two-way", CV = TRUE, r = c(0, 5),
                        se = TRUE, nboots = 200, vartype = 'bootstrap',
                        parallel = FALSE, keep.sims=TRUE)
plot(effect(out_mc))
```

We can also use jackknife instead of bootstrap for inference:

```{r effect-jackknife, cache = TRUE}
out_jack <- fect(Y ~ D + X1 + X2, data = simgsynth, index = c("id","time"),
                        method = "mc", force = "two-way", CV = TRUE, r = c(0, 5),
                        se = TRUE, nboots = 200, vartype = 'jackknife',
                        parallel = FALSE, keep.sims=TRUE)
plot(effect(out_jack))
```
