# Effect Heterogeneity {#sec-hte}

```{r echo = FALSE, message = FALSE, warning = FALSE}
set.seed(1234)
library(fect)
data(fect)
```

We provide several methods for researchers to explore heterogeneous treatment effects (HTE). These work with any method; here we demonstrate with the FE estimator using `simdata1`.

```{r hte_setup, eval=TRUE, cache=TRUE, message=FALSE, results='hide'}
out.fect <- fect(Y ~ D + X1 + X2, data = simdata1, index = c("id","time"),
  method = "fe", force = "two-way", se = TRUE,
  cores = 8, parallel = TRUE, nboots = 1000)
```

## Box plot

One way to understand HTE is to use a series of box plots to visualize the estimated individualistic treatment effects of observations under the treatment condition (by setting `type = "box"`). Although these effects are not identified at the individual observation level, their level of dispersion is informative of treatment effects heterogeneity at different (relative) time periods, as well as model performance.

```{r hte, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.fect, type = "box", xlim = c(-15, 10))
```

## By calendar time

Another way to explore HTE is to investigate how the treatment effect evolves over time. In the plot below, the point estimates represents the ATTs by calendar time; the blue curve and band represent a lowess fit of the estimates and its 95% confidence interval, respectively; and the red horizontal dashed line represents the ATT (averaged over all time periods).

```{r hte_time, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.fect, type = "calendar", xlim = c(1, 35))
```

## By a covariate

By setting `type = "hte"` or `type = "heterogeneous`, we can also plot the HTE by arbitrary covariates that are unaffected by the treatment. As before, the blue curve and band represent a lowess fit of the estimates and its 95% confidence interval, respectively. The red dashed line represents the ATT. The histogram at the bottom of the figure illustrates the distribution of the covariates, and can be turned off using `show.count = FALSE`. In our simulated case, the effect size is unrelated to the values of covariate `X1`.

```{r hte_X1, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.fect, type = "hte", covariate = "X1")
```

We can also plot the CATT when a covariate is discrete. To demonstrate this, we artificially create a moderating variable `X3`, which must be included in the outcome model and then specified in the heterogeneous treatment effect plot.

```{r hte_discrete, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
simdata1$X3 <- sample(1:3, size = nrow(simdata1), replace = TRUE)
out.fect.X3 <- fect(Y ~ D + X1 + X2 + X3, data = simdata1, index = c("id","time"),
                   method = "fe", se = TRUE, seed = 123,
                   cores = 8, nboots = 1000, parallel = TRUE)
```

As expected, there is not much effect heterogeneity along `X3`. In the resulting figure, we can also assign labels to the discrete values in the moderator.

```{r, fig.width = 6, fig.height = 4.5}
plot(out.fect.X3, type="hte", covariate = "X3",
     xlab = "", ylab = "Effet of D on Y",
     covariate.labels = c("USA", "China", "UK"),
     ylim = c(-2, 6))
```

Our next update will accommodate time-invariant covariates and allow users to explore effect heterogeneity around them.
