# Effect Heterogeneity {#sec-hte}

```{r .common, include = FALSE}
source("_common.R")
```

```{r setup-hte, echo = FALSE, message = FALSE, warning = FALSE}
set.seed(1234)
data(sim_base)
```

We provide several methods for researchers to explore heterogeneous treatment effects (HTE). These methods help distinguish between *effect modification* --- how the treatment effect varies across subpopulations --- and *causal moderation* --- whether changing the moderator causally alters the treatment effect. This chapter demonstrates both descriptive HTE tools and the formal causal moderation framework. R script used in this chapter can be downloaded [here](https://raw.githubusercontent.com/xuyiqing/fect/dev/vignettes/rscript/05-hte.R).

## Basic HTE Visualization

We start with descriptive tools using `sim_base`. These work with any estimation method; here we demonstrate with the FE estimator.

```{r hte_setup, eval=TRUE, cache=TRUE, message=FALSE, results='hide'}
out.fect <- fect(Y ~ D + X1 + X2, data = sim_base, index = c("id","time"),
  method = "fe", force = "two-way", se = TRUE,
  parallel = TRUE, cores = 16, nboots = 1000)
```

### Box plot

One way to understand HTE is to use a series of box plots to visualize the estimated individualistic treatment effects of observations under the treatment condition (by setting `type = "box"`). Although these effects are not identified at the individual observation level, their level of dispersion is informative of treatment effects heterogeneity at different (relative) time periods, as well as model performance.

```{r hte, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.fect, type = "box", xlim = c(-15, 10))
```

### By calendar time

Another way to explore HTE is to investigate how the treatment effect evolves over time. In the plot below, the point estimates represent the ATTs by calendar time; the blue curve and band represent a lowess fit of the estimates and its 95% confidence interval, respectively; and the red horizontal dashed line represents the ATT (averaged over all time periods).

```{r hte_time, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.fect, type = "calendar", xlim = c(1, 35))
```

### By a covariate

By setting `type = "hte"` or `type = "heterogeneous"`, we can also plot the HTE by arbitrary covariates that are unaffected by the treatment. As before, the blue curve and band represent a lowess fit of the estimates and its 95% confidence interval, respectively. The red dashed line represents the ATT. The histogram at the bottom of the figure illustrates the distribution of the covariates, and can be turned off using `show.count = FALSE`. In our simulated case, the effect size is unrelated to the values of covariate `X1`.

```{r hte_X1, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.fect, type = "hte", covariate = "X1")
```

We can also plot the CATT when a covariate is discrete. To demonstrate this, we artificially create a moderating variable `X3`, which must be included in the outcome model and then specified in the heterogeneous treatment effect plot.

```{r hte_discrete, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
sim_base$X3 <- sample(1:3, size = nrow(sim_base), replace = TRUE)
out.fect.X3 <- fect(Y ~ D + X1 + X2 + X3, data = sim_base, index = c("id","time"),
                   method = "fe", se = TRUE, seed = 123,
                   nboots = 1000, parallel = TRUE, cores = 16)
```

As expected, there is not much effect heterogeneity along `X3`. In the resulting figure, we can also assign labels to the discrete values in the moderator.

```{r plot-hte-discrete, fig.width = 6, fig.height = 4.5}
plot(out.fect.X3, type="hte", covariate = "X3",
     xlab = "", ylab = "Effect of D on Y",
     covariate.labels = c("USA", "China", "UK"),
     ylim = c(-2, 6))
```

---

## Causal Moderation

A prevalent approach for examining treatment effect heterogeneity in political science is the two-way fixed effects (TWFE) model with a multiplicative interaction term. Despite its widespread use, this model conflates two conceptually distinct quantities:

1. **Effect modification**: How does the *average treatment effect* vary across subpopulations defined by the moderator? This is a *correlational* relationship.
2. **Causal moderation**: Does exogenously changing the moderator *causally alter* the treatment effect? This is a *mechanistic* relationship.

Under the ignorability of $M$ assumption --- i.e., there are no unobserved links between the moderator and confounders --- the two estimands are numerically identical. In practice, divergence between the two estimates provides a useful diagnostic.

### Estimation with `cm = TRUE`

The `cm` (causal moderation) option in `fect()` fits two separate imputation models: one for untreated potential outcomes $\hat{g}_0$ and one for treated potential outcomes $\hat{g}_1$. The causal moderation CME is then $\hat{\theta}(m) = \hat{g}_1(m) - \hat{g}_0(m)$.

Currently `cm` is available for the `"fe"` and `"ife"` methods.

```{r cm_fe, eval = TRUE, cache = TRUE, message = FALSE, results = 'hide'}
out.cm <- fect(Y ~ D + X1 + X2, data = sim_base, index = c("id", "time"),
               method = "fe", force = "two-way", se = TRUE,
               cm = TRUE, parallel = TRUE, cores = 16, nboots = 1000)
```

### Effect modification vs. causal moderation

The default HTE plot shows the effect modification estimate (projecting $\hat{\delta}_{it}$ onto the moderator):

```{r hte_em, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.cm, type = "hte", covariate = "X1",
     xlab = "Moderator (X1)", ylab = "Effect on Y")
```

By adding `cm = TRUE` to the plot call, we get the causal moderation estimate:

```{r hte_cm, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.cm, type = "hte", covariate = "X1", cm = TRUE,
     xlab = "Moderator (X1)", ylab = "Effect on Y", ylim = c(-0.5, 5))
```

When the ignorability assumption holds (as in the simulated data), the two curves should be similar.

### Scatter without smoothing

To inspect the raw relationship without loess smoothing, set `loess.fit = FALSE`:

```{r hte_scatter, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.cm, type = "hte", covariate = "X1", cm = TRUE, loess.fit = FALSE,
     xlab = "Moderator (X1)", ylab = "Effect on Y", ylim = c(-0.5, 5))
```

---

## Diagnostics

### Placebo test (pre-treatment HTE)

The placebo test applies the HTE estimation to *pre-treatment* periods only. Under the conditional parallel trends assumption (CPTA), there should be no relationship between the moderator and the "treatment effect" before treatment onset:

```{r hte_placebo, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.cm, type = "hte", covariate = "X1",
     pretreatment = TRUE, num.pretreatment = 3,
     xlab = "X1", ylab = "Placebo Effect")
```

If the curve is flat around zero, it supports CPTA. A significant non-zero pattern signals that the moderator co-varies with pre-existing trends. The same test can be applied to the causal moderation estimate:

```{r hte_placebo_cm, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.cm, type = "hte", covariate = "X1", cm = TRUE,
     pretreatment = TRUE, num.pretreatment = 3,
     xlab = "X1", ylab = "Placebo Effect", ylim = c(-0.5, 1.5))
```

---

## Over-Identification Test

The linear CM model assumes $M$ enters linearly. If the true relationship involves $M^2$ or $M \times X$, the linear model is misspecified. `fect_iden()` tests this by regressing residuals on nonlinear terms:
$$
H_0:\text{Nonlinear terms have no explanatory power} \quad \Rightarrow \quad n \times R^2 \sim \chi^2(\text{df})
$$

```{r iden_test, eval = TRUE, cache = TRUE, message = FALSE}
iden.test <- fect_iden(out.cm, moderator = "X1")
```

```{r iden_results, eval = TRUE}
cat("=== Treated cells (e1) ===\n")
cat("  n =", iden.test$e1$n, "\n")
cat("  R-squared =", round(iden.test$e1$r2, 4), "\n")
cat("  Test stat =", round(iden.test$e1$stat, 3), "\n")
cat("  df =", iden.test$e1$df, "\n")
cat("  p-value =", round(iden.test$e1$p, 4), "\n\n")

cat("=== Control cells (e0) ===\n")
cat("  n =", iden.test$e0$n, "\n")
cat("  R-squared =", round(iden.test$e0$r2, 4), "\n")
cat("  Test stat =", round(iden.test$e0$stat, 3), "\n")
cat("  df =", iden.test$e0$df, "\n")
cat("  p-value =", round(iden.test$e0$p, 4), "\n")
```

- **Large p-values** ($> 0.05$): No evidence against the linear specification.
- **Small p-values** ($< 0.05$): Nonlinear effects present. Consider more flexible specifications.

You can also test quadratic and interaction terms separately:

```{r iden_components, eval = TRUE, cache = TRUE, message = FALSE}
# Quadratic terms only (no interactions)
iden.quad <- fect_iden(out.cm, moderator = "X1", interaction = FALSE)
cat("Quadratic-only: p =",
    round(iden.quad$e1$p, 4), "(treated),",
    round(iden.quad$e0$p, 4), "(control)\n")

# Interactions only (no quadratics)
iden.inter <- fect_iden(out.cm, moderator = "X1", quadratic = FALSE)
cat("Interaction-only: p =",
    round(iden.inter$e1$p, 4), "(treated),",
    round(iden.inter$e0$p, 4), "(control)\n")
```

---

## Discrete Moderator with Causal Moderation

```{r discrete_cm_setup, eval = TRUE, cache = TRUE, message = FALSE, results = 'hide'}
sim_base$X3 <- sample(1:3, size = nrow(sim_base), replace = TRUE)
out.discrete <- fect(Y ~ D + X1 + X2 + X3, data = sim_base,
                     index = c("id", "time"),
                     method = "fe", force = "two-way", se = TRUE,
                     cm = TRUE, parallel = TRUE, cores = 16, nboots = 1000)
```

```{r discrete_em_plot, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.discrete, type = "hte", covariate = "X3",
     covariate.labels = c("USA", "China", "UK"),
     xlab = "", ylab = "Effect on Y", ylim = c(-0.5, 5))
```

```{r discrete_cm_plot, eval = TRUE, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 4.5}
plot(out.discrete, type = "hte", covariate = "X3", cm = TRUE,
     covariate.labels = c("USA", "China", "UK"),
     xlab = "", ylab = "Effect on Y", ylim = c(-0.5, 5))
```

Since `X3` is randomly assigned and unrelated to the DGP, both estimates should show no significant heterogeneity.

---

## Summary of Parameters

| Parameter | Description |
|:---|:---|
| `cm = TRUE` in `fect()` | Estimate causal moderation model (dual imputation) |
| `cm = TRUE` in `plot()` | Plot causal moderation effect $\hat{\theta}(m)$ |
| `type = "hte"` | HTE plot: individual effects vs. covariate |
| `covariate` | Specify the moderator variable |
| `pretreatment = TRUE` | Restrict to pre-treatment periods (placebo test) |
| `num.pretreatment` | Number of pre-treatment periods to include |
| `loess.fit = FALSE` | Scatter only, without loess smoothing |
| `covariate.labels` | Labels for discrete moderator categories |
| `fect_iden()` | Over-identification test for linearity |
