# Model Selection {#sec-model-selection}

```{r echo = FALSE, message = FALSE, warning = FALSE}
set.seed(1234)
library(fect)
data(fect)
```

When using `method = "ife"` or `method = "mc"`, we need to choose a tuning parameter --- the number of factors `r` (IFE) or the regularization strength `lambda` (MC). The `method = "gsynth"` (nevertreated) path also selects `r` via cross-validation. This chapter explains the cross-validation framework, masking strategies, scoring criteria, and the post-hoc model comparison function `fect_mspe()`.

## Cross-validation overview

Setting `CV = TRUE` activates the built-in cross-validation procedure. In each round, a subset of observations is masked (held out), the model is re-estimated on the remaining data, and the prediction error on the held-out set is scored. The tuning parameter that minimizes the chosen criterion is selected.

```{r cv_ife_demo, eval=TRUE, cache=TRUE, message=FALSE, results='hide'}
out.cv <- fect(Y ~ D + X1 + X2, data = simdata, index = c("id","time"),
               method = "ife", CV = TRUE, r = c(0, 5),
               se = FALSE, parallel = TRUE)
```

```{r}
cat("Selected r:", out.cv$r.cv, "\n")
```

------------------------------------------------------------------------

## The `cv.method` parameter

The `cv.method` parameter controls **which observations are masked** during cross-validation. Different masking strategies test different aspects of the model.

| `cv.method` | What is masked | What it tests |
|:------------|:--------------|:-------------|
| `"all_units"` | Random subsets of control observations across all units | Factor estimation quality on the full panel |
| `"treated_units"` | Random subsets from eventually-treated units' pre-treatment observations only | Counterfactual prediction quality for the target units |
| `"loo"` | One treated pre-treatment period at a time | Projection quality (legacy gsynth method) |

::: {.callout-note appearance="simple"}
### Which `cv.method` should I use?

`"all_units"` (default for `fect_cv` and `fect_mspe`) is the general-purpose choice — it tests whether the model can predict held-out control observations well, and is suitable when the panel is large and the factor structure is the main concern. `"treated_units"` (default for nevertreated/gsynth) directly tests counterfactual quality for the treated group, and is preferred when prediction accuracy for treated units is the priority. `"loo"` (nevertreated only) is the original gsynth leave-one-out method; it can be unstable with few pre-treatment periods, in which case `"treated_units"` is a more robust alternative.
:::

### Example: comparing `cv.method` choices

```{r cv_method_compare, eval=TRUE, cache=TRUE, message=FALSE, results='hide'}
out.all <- fect(Y ~ D + X1 + X2, data = simdata, index = c("id","time"),
                method = "ife", CV = TRUE, r = c(0, 5),
                cv.method = "all_units", se = FALSE, parallel = TRUE)

out.tr <- fect(Y ~ D + X1 + X2, data = simdata, index = c("id","time"),
               method = "ife", CV = TRUE, r = c(0, 5),
               cv.method = "treated_units", se = FALSE, parallel = TRUE)
```

```{r}
cat("cv.method = 'all_units':     r.cv =", out.all$r.cv, "\n")
cat("cv.method = 'treated_units': r.cv =", out.tr$r.cv, "\n")
```

Both strategies often agree, but they can select different `r` when the factor structure matters more for one group than the other.

------------------------------------------------------------------------

## Cross-validation sub-options

The k-fold masking procedure is controlled by several sub-options that apply to both `"all_units"` and `"treated_units"`:

| Option | Default | Description |
|:-------|:--------|:-----------|
| `k` | 10 | Number of cross-validation rounds |
| `cv.prop` | 0.1 | Proportion of eligible observations held out per round |
| `cv.nobs` | 3 | Block size for structured removal (consecutive observations) |
| `cv.donut` | 0 | Periods excluded around treatment onset in the evaluation set |
| `min.T0` | 5 | Minimum pre-treatment observations required per unit |

The block removal (`cv.nobs`) addresses serial correlation: instead of masking individual observations, we mask blocks of consecutive observations. The donut option (`cv.donut`) further excludes observations near the treatment boundary from the test set to avoid contamination.

------------------------------------------------------------------------

## Scoring criteria

The `criterion` parameter determines which scoring metric is used to select the best tuning parameter:

| `criterion` | Description | When to use |
|:------------|:-----------|:-----------|
| `"mspe"` (default) | Mean squared prediction error | General-purpose; best for prediction accuracy |
| `"gmspe"` | Geometric mean SPE | Robust to outlier prediction errors |
| `"pc"` | Information criterion (Bai & Ng) | Model selection based on penalized fit; does not require CV rounds |
| `"moment"` | Weighted period-wise residual averages | Selects for best pre-trend fitting rather than prediction |

```{r criterion_compare, eval=TRUE, cache=TRUE, message=FALSE, results='hide'}
out.mspe <- fect(Y ~ D + X1 + X2, data = simdata, index = c("id","time"),
                 method = "ife", CV = TRUE, r = c(0, 5),
                 criterion = "mspe", se = FALSE, parallel = TRUE)

out.pc <- fect(Y ~ D + X1 + X2, data = simdata, index = c("id","time"),
               method = "ife", CV = TRUE, r = c(0, 5),
               criterion = "pc", se = FALSE, parallel = TRUE)
```

```{r}
cat("criterion = 'mspe': r.cv =", out.mspe$r.cv, "\n")
cat("criterion = 'pc':   r.cv =", out.pc$r.cv, "\n")
```

::: {.callout-tip appearance="simple"}
### Selection rule
A candidate `r` is selected over a smaller value only if its criterion score improves by more than 1%. This prevents overfitting to marginal improvements.
:::

------------------------------------------------------------------------

## Post-hoc model comparison with `fect_mspe()`

After fitting multiple models, `fect_mspe()` provides a unified framework for comparing their out-of-sample prediction accuracy. Unlike the internal CV (which selects a tuning parameter within a single method), `fect_mspe()` compares across methods or specifications.

```{r mspe_demo, eval=TRUE, cache=TRUE, message=FALSE, results='hide'}
out.fe <- fect(Y ~ D + X1 + X2, data = simdata, index = c("id","time"),
               method = "fe", se = FALSE, CV = FALSE)

out.ife2 <- fect(Y ~ D + X1 + X2, data = simdata, index = c("id","time"),
                 method = "ife", r = 2, se = FALSE, CV = FALSE)
```

```{r mspe_run, eval=TRUE, cache=TRUE}
mspe.out <- fect_mspe(list(out.fe, out.ife2), seed = 1234)
print(mspe.out$summary)
```

`fect_mspe()` accepts all the same masking options (`cv.method`, `k`, `cv.prop`, `cv.nobs`, `cv.donut`, `min.T0`) and scoring options (`criterion`) as the internal CV functions. By default, it uses `cv.method = "all_units"`.

::: {.callout-note appearance="simple"}
A smaller MSPE indicates better out-of-sample prediction. The function does not perform model selection automatically — it reports scores for the user to compare.
:::

------------------------------------------------------------------------

## Availability by estimation regime

Not all `cv.method` options are available in all settings:

| | `fect_cv` (notyettreated) | nevertreated CV | `fect_mspe` (post-hoc) |
|:--|:--|:--|:--|
| `"all_units"` | Default | Available | Default |
| `"treated_units"` | Available | Default | Available |
| `"loo"` | --- | Available | --- |

The `"loo"` method is exclusive to the nevertreated regime because it requires factor estimation to be independent of the held-out treated observations. In the nevertreated setting, factors are estimated from control units only, so holding out treated pre-treatment periods does not compromise estimation. In the notyettreated setting, all units contribute to estimation, making LOO less clean; use `"treated_units"` instead.

------------------------------------------------------------------------

## Parallel cross-validation

Cross-validation can be computationally expensive, especially with `cv.method = "all_units"` in the nevertreated setting (which re-estimates the full factor model `k` times per candidate `r`). Parallel computing is enabled by default when `parallel = TRUE` in `fect()`.

For the nevertreated path, parallel CV auto-activates when the panel is large enough ($N_{co} \times T > 20{,}000$) and `cv.method = "all_units"`. The `"treated_units"` and `"loo"` methods are always sequential because the per-fold cost is too low to benefit from parallelization overhead.
