# Sensitivity Analysis {#sec-panel-sens}

Download the R code used in this chapter [here](rscript/05-sens.R).

------------------------------------------------------------------------

@rambachan2023more propose a partial identification approach that relaxes the PT assumption in the post-treatment period by allowing violations that do not exceed the magnitude of those observed in the pre-treatment period. This framework enables sensitivity analysis of estimates from **fect** or similar methods by comparing pre-treatment deviations from PT to potential post-treatment deviations.

The key intuition is that if an event study demonstrates strong post-treatment effects yet only minor PT deviations before treatment, any post-treatment departure large enough to reverse these findings must be substantially larger than those observed in the pre-treatment period. Consequently, this approach quantifies how sensitive the estimated dynamic treatment effects are to possible PT violations, using pretrend estimates as the benchmark.

Below, we illustrate how to apply this sensitivity analysis with **fect**. We focus on two restrictions from @rambachan2023more: the relative magnitude (RM) restriction and the smoothness restriction, both of which connect pre-treatment PT violations to potential post-treatment counterfactual deviations.

## Install Packages

To begin, you will need to install the necessary packages from CRAN and GitHub.

```{r message = FALSE, warning = FALSE}
# install packages from CRAN
packages <- c("dplyr", "panelView", "ggplot2") # Removed HonestDiD, doParallel
install.packages(setdiff(packages, rownames(installed.packages())))  

# install most up-to-date "fect" from Github
if ("fect" %in% rownames(installed.packages()) == FALSE) {
  devtools:: install_github("xuyiqing/fect")
}

# install forked "HonestDiD" package compatible with "fect"
if ("HonestDiDFEct" %in% rownames(installed.packages()) == FALSE) {
  devtools:: install_github("lzy318/HonestDiDFEct") # This is used by fect_sens
}
```

Load libraries:

```{r message = FALSE, warning = FALSE}
library(dplyr)
library(fect)
library(panelView)
library(ggplot2)
library(HonestDiDFEct) # Required for fect_sens to work
```

## No Treatment Reversals

We begin with an empirical example from @HH2019, who investigate the effects of indirect democracy versus direct democracy on naturalization rates in Switzerland using municipality-year panel data from 1991 to 2009. The study finds that switching from direct to indirect democracy increased naturalization rates by an average of 1.22 percentage points (Model 1, Table 1).

```{r message = FALSE, warning = FALSE}
data(fect)
data <- hh2019
head(data)
```

### Implement with Placebo Tests

To implement this method with the imputation estimator, we use the dynamic treatment effects from pre-treatment *placebo tests* to gauge PT violations and determine whether post-treatment effects remain significant under similar violations. This requires symmetric estimation of dynamic treatment effects in both pre- and post-treatment periods. As [@roth2024interpreting] notes, some estimators (e.g., CSDID without `base_period = "universal"`) may not produce symmetrical estimates.

Below, we designate placebo periods using **fect**. These placebo periods are excluded during model fitting, and counterfactuals are imputed for both placebo and post-treatment intervals to compute dynamic treatment effects, ensuring consistent estimation across all periods.

By setting `placeboTest = TRUE` and `placebo.period = c(-2, 0)`, we define three pre-treatment periods as placebo periods. Their dynamic treatment effects serve as the benchmark for PT violations in the post-treatment phase. In the code chunk below, we fit **fect** with these placebo settings, then use the `fect_sens` function to perform the sensitivity analysis. This function wraps the procedures from **HonestDiDFEct**, preparing the output for plotting. Note that not all of `Mbarvec`, `periodMbarvec`, `Mvec`, or `periodMvec` need to be specified; only the ones you want to use for the sensitivity analysis.

```{r, hh_honest_placebo, warning=FALSE, message=FALSE, cache=TRUE}
out.fect.placebo <- fect(nat_rate_ord~indirect, data = hh2019, 
                         index = c("bfs","year"),
                         method = 'fe', se = TRUE, 
                         placeboTest = TRUE, placebo.period = c(-2,0))

# Define post-treatment periods and sensitivity parameters for fect_sens
T.post <- 10 # Number of post-treatment periods based on original analysis
post_periods_vec <- 1:T.post

# Parameters for Relative Magnitude (RM) restriction
Mbar_vec_avg_rm <- seq(0, 1, by = 0.1)    # For average ATT plot
Mbar_vec_period_rm <- c(0, 0.5)          # For period-by-period ATT plot

# Parameters for Smoothness restriction
M_vec_avg_smooth <- seq(0, 0.25, by = 0.05) # For average ATT plot
M_vec_period_smooth <- c(0, 0.1)           # For period-by-period ATT plot

# Run sensitivity analysis using fect_sens
# This function augments out.fect.placebo with sensitivity results
out.fect.placebo <- fect_sens(
  fect.out      = out.fect.placebo,
  post.periods  = post_periods_vec,
  Mbarvec       = Mbar_vec_avg_rm,
  periodMbarvec = Mbar_vec_period_rm,
  Mvec          = M_vec_avg_smooth,
  periodMvec    = M_vec_period_smooth,
  parallel      = FALSE # Set to TRUE for parallel processing if desired
)
```

### RM Restriction

We first explore the Relative Magnitude (RM) restriction. Let $\delta$ represent potential PT violations for placebo and post-treatment periods. Unlike a standard event study that assumes $\delta_t=0$ for $t>0$, RM allows PT deviations as long as they do not exceed $\bar{M}$ times the maximum deviation between consecutive placebo periods.

The **fect** package, through `fect_sens`, utilizes a forked version of **HonestDiD**, called **HonestDiDFEct**, which is adapted for **fect**'s output structure. The RM restriction is defined as:

$$
\Delta^{RM}(\bar{M}) = \left\{\delta : \forall t \geq 0,\; 
\bigl|\delta_{t+1} - \delta_t\bigr| \leq \bar{M}\cdot\max\bigl(
|\delta_{-1} - \delta_{-2}|,\,|\delta_0 - \delta_{-1}|
\bigr)\right\}.
$$

Here, $max(|\delta_{-1}-\delta_{-2}|,|\delta_0-\delta_{-1}|)$ matches the largest consecutive discrepancy among the placebo periods. When $\bar{M}=0$, the PT violation observed at $t=0$ persists without change into the post-treatment window. Allowing $\bar{M}>0$ permits $\delta_t$ to vary across post-treatment periods, but the incremental changes must remain within $\bar{M}$ times the largest consecutive deviation among placebo periods.

**Robust Confidence Set for the ATT**

We begin by constructing a robust confidence set for the overall ATT. The `fect_sens` function has already computed these results, using count-based weights for the ATT by default. We specified `Mbarvec = seq(0, 1, by=0.1)` in our call to `fect_sens`. Increasing $\bar{M}$ allows proportionally larger PT violations in the post-treatment window. When $\bar{M}=0$, the resulting confidence set behaves as a "de-biased" interval that corrects post-treatment estimates based on the observed PT violation at $t=0$.

We can now plot the robust confidence intervals for the ATT using the RM restriction with `plot()` and `type = "sens"`:

```{r,hh_honest.placebo.honest, fig.width = 6, fig.height = 4.5, cache=TRUE}
plot(out.fect.placebo,
     type = "sens",
     restrict = "rm",
     main = "Relative Magnitude Restriction")
```

If the robust confidence set excludes zero at $\bar{M}=0.4$ but includes zero at $\bar{M}=0.5$, we infer that post-treatment PT violations must be at least half of the maximum observed placebo violation to overturn the estimated effect.

**Period-by-Period Robust Confidence Set**

The `fect_sens` function, with the `periodMbarvec` argument set to `c(0, 0.5)`, computes period-specific robust confidence intervals. We can visualize them using `plot()` with `type = "sens_es"`:

```{r,hh_honest.placebo.honest.gap.plot, fig.width=7, fig.height=5, cache=TRUE}
plot(out.fect.placebo,
    type = "sens_es",
    restrict = "rm",
    main = "ATTs with Robust Confidence Sets (RM)",
    ylab = "Coefficients and 95% CI",
    xlim = c(-12,10), 
    ylim = c(-6,8), 
    show.count = TRUE)
```

In the figure, different lines/bands represent the robust confidence intervals for $\bar{M}=0$ and $\bar{M}=0.5$. The interval for $\bar{M}=0$ treats the observed violation at $t=0$ as persisting into all post-treatment periods, whereas the interval for $\bar{M}=0.5$ allows added PT violations up to half of the largest placebo discrepancy. These are compared against the original confidence intervals. You can also change the colors using the `sens.colors` argument in the `plot()` function. This works for the regular `type = "sens"` plot as well, but with a vector of only one color.

```{r,hh_honest.placebo.honest.gap.plot.colors, fig.width=7, fig.height=5, cache=TRUE}
plot(out.fect.placebo,
    type = "sens_es",
    restrict = "rm",
    main = "ATTs with Robust Confidence Sets (RM)",
    ylab = "Coefficients and 95% CI",
    xlim = c(-12,10), 
    ylim = c(-6,8), 
    show.count = TRUE,
    sens.colors = c("blue", "red"))
```

### Smoothness Restriction

A second approach to bounding PT violations is the **smoothness restriction**, which prevents the post-treatment violation from diverging too sharply from a linear extrapolation of the pre-trend. This restriction is particularly relevant if we suspect a gradually varying or near-linear trend in the potential violation.

Formally, one assumes $\delta\in\Delta^{SD}(M)$ where

$$
\Delta^{SD}(M) := \bigl\{\delta : 
\bigl| (\delta_{t+1}-\delta_t) - (\delta_t -\delta_{t-1}) \bigr| \leqslant M,\ \forall t\bigr\}.
$$

The parameter $M\geq 0$ controls how quickly the slope of $\delta$ can vary. Note that $M=0$ restricts $\delta$ to be strictly linear, which does not imply $\delta=0$.

The `fect_sens` function, using the `Mvec` argument (set to `seq(0,0.25,0.05)` in our call), has already computed these results. We plot them for the overall ATT using `plot()` with `type = "sens"` and `restrict = "sm"`:

```{r,hh_honest.placebo.honest.sd,   fig.width = 7, fig.height = 5,  cache=TRUE}
plot(out.fect.placebo,
    type = "sens",
    restrict = "sm",
    main = "Smoothness Restriction")
```

Here, if the post-treatment effect is indistinguishable from zero even when $M=0$, it suggests that the estimated average treatment effect may not be robust to a strictly linear violation of PT observed in the placebo periods, let alone to variations in the slope of this linear trend.

**Period-by-Period Robust Confidence Set**

Finally, we construct period-specific robust confidence sets under the smoothness restriction. The `fect_sens` function, with `periodMvec` set to `c(0, 0.1)` in our call, calculates these. The plot is generated using `type = "sens_es"` and `restrict = "sm"`:

```{r,hh_honest.placebo.honest.gap.sd.plot, fig.height=5, fig.width=7, cache=TRUE, warning=FALSE}
plot(out.fect.placebo,
    type = "sens_es",
    restrict = "sm",
    main = "ATTs with Robust Confidence Sets (Smoothness)",
    ylab = "Coefficients and 95% CI",
    xlim = c(-12.5,10.5), # Adjusted to match original detailed plot
    ylim = c(-12,15),
    show.count = TRUE)
```

In this figure, different lines/bands will represent the robust CIs for $M=0$ (strictly linear PT violation) and $M=0.1$ (allowing a slope deviation of 0.1). These are compared against the original estimates and placebo period estimates to assess the robustness of dynamic treatment effects.

## How to Cite

Please cite the authors of the original papers for their innovations. If you find this tutorial helpful, you can cite @CLLX2025.
