# Complex Fixed Effects (CFE) {#sec-cfe}

This chapter provides a dedicated tutorial for the **Complex Fixed Effects (CFE)** estimator in **fect**. Download the R code used in this chapter [here](rscript/07-cfe.R). The CFE method generalizes the standard two-way fixed effects counterfactual estimator by allowing additional model components: extra additive fixed effects, time-invariant covariates with time-grouped coefficients, unit-specific loadings on known time trends, and interactive fixed effects via latent factors.

```{r echo = FALSE}
set.seed(1234)
rm(list = ls())
```

```{r message = FALSE, warning = FALSE}
library(fect)
data(fect)
```

------------------------------------------------------------------------

## When to use CFE

The standard FEct estimator (`method = "fe"`) assumes that the untreated potential outcome is driven by unit and time fixed effects plus observed time-varying covariates. This may be insufficient when:

-   Units belong to groups (e.g., regions, sectors) with group-level time trends,
-   Time-invariant characteristics (e.g., baseline demographics) have effects that vary across time periods,
-   Unit-specific loadings on known time trends (e.g., linear, quadratic, or spline trends) are needed,
-   Interactive fixed effects (latent factors) are also present.

The CFE estimator fits a richer model on untreated observations, imputes counterfactuals for treated observations, and then estimates the ATT.

------------------------------------------------------------------------

## Model

The CFE estimator models the untreated potential outcome as:

$$
Y_{it}^{0} = \mu + X_{it}'\beta + \sum_{k=1}^{p_\gamma} Z_{i,S_k}' \gamma_{k, g_k(t)} + \sum_{k=1}^{p_\kappa} \kappa_{h_k(i),k}' Q_{k,t} + \alpha_i + \xi_t + \omega_k + \nu_l + \sum_{m=1}^{r} \lambda_{im} f_{tm} + e_{it}
$$

where:

-   $\mu$ is the grand mean,
-   $X_{it}'\beta$ captures time-varying covariates,
-   $Z'\gamma$ terms are time-invariant covariates with time-grouped coefficients,
-   $\kappa'Q$ terms are unit/group-specific loadings on known time trends,
-   $\alpha_i, \xi_t, \omega_k, \nu_l$ are additive fixed effects (unit, time, and additional levels),
-   $\lambda_{im} f_{tm}$ are interactive fixed effects (latent factors).

The model is estimated via an EM-style block coordinate descent algorithm that iterates over the five components until convergence.

------------------------------------------------------------------------

## Basic usage: additional additive fixed effects

The simplest CFE use case adds extra additive fixed effects beyond unit and time. In the `index` argument, any elements beyond the first two are treated as additional fixed effect groupings.

```{r cfe_sfe, eval=TRUE, cache = TRUE, message = FALSE, results = 'hide'}
# Create an additional grouping variable
simdata[,"FE3"] <- sample(1:5, size = nrow(simdata), replace = TRUE)

out.cfe.basic <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time", "FE3"),
  method = "cfe", force = "two-way",
  se = TRUE, parallel = TRUE, nboots = 200)
```

```{r cfe_sfe_plot, fig.width = 6, fig.height = 4.5}
plot(out.cfe.basic, main = "CFE with Additional Fixed Effect",
     ylab = "Effect of D on Y",
     cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8)
```

::: {.callout-note appearance="simple"}
The `method = "cfe"` interface puts additional fixed effects in `index` and uses `Z`/`Q`/`gamma`/`kappa` for the richer covariate structure.
:::

------------------------------------------------------------------------

## Unit-specific time trends via `Q.type`

A common use case is to allow unit-specific time trends. The `Q.type` option automatically generates time bases (linear, quadratic, cubic, or B-spline) and estimates unit-specific loadings on them.

### Linear time trend

```{r cfe_linear, eval=TRUE, cache = TRUE, message = FALSE, results = 'hide'}
out.cfe.linear <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time"),
  method = "cfe", force = "two-way",
  Q.type = "linear",
  se = TRUE, parallel = TRUE, nboots = 200)
```

```{r cfe_linear_plot, fig.width = 6, fig.height = 4.5}
plot(out.cfe.linear, main = "CFE with Unit-Specific Linear Trend",
     ylab = "Effect of D on Y",
     cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8)
```

### Quadratic time trend

```{r cfe_quad, eval=TRUE, cache = TRUE, message = FALSE, results = 'hide'}
out.cfe.quad <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time"),
  method = "cfe", force = "two-way",
  Q.type = c("linear", "quadratic"),
  se = TRUE, parallel = TRUE, nboots = 200)
```

```{r cfe_quad_plot, fig.width = 6, fig.height = 4.5}
plot(out.cfe.quad, main = "CFE with Unit-Specific Quadratic Trend",
     ylab = "Effect of D on Y",
     cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8)
```

### B-spline time trend

For more flexible nonlinear trends, use `Q.type = "bspline"`. The degree of the B-spline can be controlled via `Q.bspline.degree`.

```{r cfe_bspline, eval=TRUE, cache = TRUE, message = FALSE, results = 'hide'}
out.cfe.bs <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time"),
  method = "cfe", force = "two-way",
  Q.type = "bspline", Q.bspline.degree = 3,
  se = TRUE, parallel = TRUE, nboots = 200)
```

```{r cfe_bspline_plot, fig.width = 6, fig.height = 4.5}
plot(out.cfe.bs, main = "CFE with B-spline Time Trend",
     ylab = "Effect of D on Y",
     cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8)
```

------------------------------------------------------------------------

## Time-invariant covariates with time-varying coefficients (`Z`/`gamma`)

When you have time-invariant unit characteristics $Z_i$ (e.g., baseline GDP, population) and want their effects to vary over time, use the `Z` and `gamma` arguments.

-   `Z`: a character vector naming time-invariant covariate columns in the data.
-   `gamma`: a character vector naming columns that define the time grouping for the $Z$ coefficients. If `gamma` is a single variable, the same time groups apply to all $Z$ columns.

```{r cfe_z_setup}
# Create a time-invariant covariate (constant within each unit)
set.seed(42)
unit_values <- rnorm(length(unique(simdata$id)))
names(unit_values) <- unique(simdata$id)
simdata$Z_baseline <- unit_values[as.character(simdata$id)]

# Create a time-grouping variable (e.g., early vs late periods)
simdata$time_group <- ifelse(simdata$time <= 17, 1, 2)
```

```{r cfe_z, eval=TRUE, cache = TRUE, message = FALSE, results = 'hide'}
out.cfe.z <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time"),
  method = "cfe", force = "two-way",
  Z = "Z_baseline", gamma = "time_group",
  se = TRUE, parallel = TRUE, nboots = 200)
```

```{r cfe_z_plot, fig.width = 6, fig.height = 4.5}
plot(out.cfe.z, main = "CFE with Time-Invariant Z",
     ylab = "Effect of D on Y",
     cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8)
```

------------------------------------------------------------------------

## CFE with interactive fixed effects

The CFE estimator can also include latent factors (interactive fixed effects) alongside the other components. Set `r` to specify the number of factors. Cross-validation (`CV = TRUE`) can help select the optimal number.

```{r cfe_ife, eval=TRUE, cache = TRUE, message = FALSE, results = 'hide'}
out.cfe.ife <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time"),
  method = "cfe", force = "two-way",
  Q.type = "linear",
  r = 2,
  se = TRUE, parallel = TRUE, nboots = 200)
```

```{r cfe_ife_plot, fig.width = 6, fig.height = 4.5}
plot(out.cfe.ife,
     main = "CFE with Linear Trend + 2 Factors",
     ylab = "Effect of D on Y",
     cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8)
```

------------------------------------------------------------------------

## Combining multiple components

The full power of CFE comes from combining several components. Below, we estimate a model with:

- Two-way fixed effects plus an additional grouping variable,
- Unit-specific linear and quadratic time trends,
- Two latent factors.

```{r cfe_full, eval=TRUE, cache = TRUE, message = FALSE, results = 'hide'}
out.cfe.full <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time", "FE3"),
  method = "cfe", force = "two-way",
  Q.type = c("linear", "quadratic"),
  r = 2,
  se = TRUE, parallel = TRUE, nboots = 200)
```

```{r cfe_full_plot, fig.width = 6, fig.height = 4.5}
plot(out.cfe.full,
     main = "CFE: Full Model",
     ylab = "Effect of D on Y",
     cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8)
```

------------------------------------------------------------------------

## Diagnostic tests

All diagnostic tests available for other estimators work with CFE as well.

### Pre-trend test

```{r cfe_pretrend, eval=TRUE, cache = TRUE, fig.width = 6, fig.height = 4.5}
plot(out.cfe.linear, type = "equiv", ylim = c(-4, 4),
     cex.legend = 0.6, main = "Pre-Trend Test (CFE)", cex.text = 0.8)
```

### Placebo test

```{r cfe_placebo, eval=TRUE, cache = TRUE, message = FALSE, results = 'hide'}
out.cfe.p <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time"),
  method = "cfe", force = "two-way",
  Q.type = "linear",
  se = TRUE, parallel = TRUE, nboots = 200,
  placeboTest = TRUE, placebo.period = c(-2, 0))
```

```{r cfe_placebo_plot, fig.width = 6, fig.height = 4.5}
plot(out.cfe.p, cex.text = 0.8,
     stats = c("placebo.p", "equiv.p"),
     main = "Placebo Test (CFE)")
```

### Carryover test

If the treatment switches on and off, we can test for carryover effects:

```{r cfe_carryover, eval=TRUE, cache = TRUE, message = FALSE, results = 'hide'}
out.cfe.c <- fect(Y ~ D + X1 + X2, data = simdata,
  index = c("id", "time"),
  method = "cfe", force = "two-way",
  Q.type = "linear",
  se = TRUE, parallel = TRUE, nboots = 200,
  carryoverTest = TRUE, carryover.period = c(1, 3))
```

```{r cfe_carryover_plot, fig.width = 6, fig.height = 4.5}
plot(out.cfe.c, type = "exit", ylim = c(-2.5, 4.5),
     cex.text = 0.8, main = "Carryover Test (CFE)")
```

------------------------------------------------------------------------

## Block assignment with `Z.param` and `Q.param`

When you have multiple $Z$ and $\gamma$ variables and want to control which $Z$ columns share which $\gamma$ coefficients, use `Z.param`. Similarly, `Q.param` controls the block assignment for $\kappa$ and $Q$.

Each parameter is a named list where:

-   **Names** = the $\gamma$ (or $\kappa$) variable names,
-   **Values** = character vectors of $Z$ (or $Q$) variable names assigned to that block.

```{r cfe_param_example, eval = FALSE}
# Example: two Z variables with different time groupings
out <- fect(Y ~ D, data = mydata,
  index = c("unit", "time"),
  method = "cfe", force = "two-way",
  Z = c("baseline_gdp", "baseline_pop"),
  gamma = c("era1", "era2"),
  Z.param = list(era1 = "baseline_gdp", era2 = "baseline_pop"))
```

------------------------------------------------------------------------

## Summary of CFE-specific arguments

| Argument | Type | Description |
|:---------|:-----|:------------|
| `method = "cfe"` | character | Activates the CFE estimator |
| `index` | character vector | `c(unit, time, ...)` — elements beyond the first two are extra additive FEs |
| `Z` | character vector | Names of time-invariant covariate columns |
| `gamma` | character vector | Names of columns defining time grouping for $Z$ coefficients |
| `Q` | character vector | Names of known time trend columns |
| `kappa` | character vector | Names of columns defining unit/group assignment for $Q$ loadings |
| `Q.type` | character vector | Auto-generate $Q$: `"linear"`, `"quadratic"`, `"cubic"`, `"bspline"` |
| `Q.bspline.degree` | integer | Degree of B-spline when `Q.type = "bspline"` |
| `Z.param` | named list | Block assignment: which $Z$ columns share which $\gamma$ |
| `Q.param` | named list | Block assignment: which $Q$ columns share which $\kappa$ |
| `r` | integer or vector | Number of latent factors (interactive FE); use `c(0, 5)` with `CV = TRUE` |
| `force` | character | Additive FE: `"none"`, `"unit"`, `"time"`, `"two-way"` |

------------------------------------------------------------------------

## Algorithm details

The CFE estimator uses an EM-style block coordinate descent procedure:

1.  **E-step (impute)**: Replace missing/treated entries with current fitted values.
2.  **Time-varying covariates** ($\beta$): OLS on the imputed panel.
3.  **$Z\gamma_t$ blocks**: For each block $k$, regress the partial residual on $Z_{S_k}$ within each time group.
4.  **$\kappa Q$ blocks**: For each block $k$, regress the partial residual on $Q_k$ within each unit group.
5.  **Additive fixed effects**: Demean by unit, time, and additional grouping variables.
6.  **Interactive fixed effects**: Rank-$r$ SVD of the partial residual.
7.  **Convergence**: Stop when the relative Frobenius norm change in each block and globally falls below `tol`.

This procedure is implemented in C++ (`src/cfe.cpp`) for computational efficiency.
