# Gsynth Program {#sec-gsynth}

```{r .common, include = FALSE}
source("_common.R")
```

This chapter demonstrates the generalized synthetic control method, or Gsynth, proposed in @Xu2017 \[<a href="https://www.cambridge.org/core/journals/political-analysis/article/generalized-synthetic-control-method-causal-inference-with-interactive-fixed-effects-models/B63A8BD7C239DD4141C67DA10CD0E4F3" target="_blank">Paper</a>\].

::: {.callout-tip appearance="simple"}
The **gsynth** package as a standalone package still exists, but algorithmically is a wrapper of the **fect** package. Click [here](https://yiqingxu.org/packages/gsynth/) if you want to use the old syntax. 
:::

Gsynth was originally implemented in the **gsynth** package but has now been fully integrated into the **fect** package. Gsynth (`method = "gsynth"`) and FEct/IFEct/MC (`method = "fe"/"ife"/"mc"`) are different in the following ways:

-   Gsynth is designed to handle block and staggered DID settings **without** treatment reversal, whereas other methods allow for treatment reversal under the assumption of limited carryover effects.
-   Gsynth is particularly suited for the Synth setting, where the number of treated units is small, including scenarios with only one treated unit. By setting `vartype = "parametric"`, we can use a two-stage parametric bootstrapping procedure to produce uncertainty estimates. In contrast, other methods rely on large samples, particularly a large number of treated units, to obtain reliable standard errors and confidence intervals using `"bootstrap"` or `"jackknife"`.
-   Compared with IFEct (`method = "ife"`), Gsynth *does not* rely on pre-treatment data from the treated units to estimate time components (e.g., factors). Hence, `time.component.from = "nevertreated"`. This approach speeds up computation, improves stability, and is more suitable for predictive inference based on row exchangeability.

Therefore, we recommend setting `method = "gsynth"` in **fect** for the synthetic control setting, where the treatment does not reverse (or is coded accordingly) and the number of treated units is small. R script used in this chapter can be downloaded [here](https://raw.githubusercontent.com/xuyiqing/fect/dev/vignettes/rscript/07-gsynth.R).

::: {.callout-important appearance="simple"}
### Key Equivalence
`fect(..., method = "gsynth")` is equivalent to `fect(..., method = "ife", time.component.from = "nevertreated")`.

Gsynth is a procedure based on an IFE model estimated using only never-treated controls. The gsynth package was developed in the synthetic control setting, where a small number of treated units are compared with never-treated controls to learn latent time effects.
:::

Use `method = "gsynth"` (or equivalently `time.component.from = "nevertreated"`) when:

-   The treatment does *not* reverse (block or staggered adoption only).
-   The number of treated units is *small*, including $N_{tr} = 1$.
-   You want *parametric bootstrap* (`vartype = "parametric"`) for uncertainty estimation — this is particularly useful with few treated units.
-   You do *not* want pre-treatment data from treated units to influence time component (i.e., factor) estimation.

In contrast, the default regime (`time.component.from = "notyettreated"`, used in [Chapter @sec-fect]) accommodates treatment reversal and uses EM to estimate time components (time fixed effects, latent factors, and temporal dynamics) from all not-yet-treated observations.

We will use two datasets, `sim_gsynth` and `turnout`, to perform analyses in block and staggered DID settings. First, we load the two datasets: `sim_gsynth` and `turnout`:

```{r setup-seed, echo = FALSE}
set.seed(1234)
```

```{r load-packages, warning=FALSE, message=FALSE}
data(sim_gsynth)
data(turnout)
ls()
```

## Basic Usage 

### Simulated Data

We start with the first example, `sim_gsynth`, a simulated dataset described in @Xu2017.

There are 5 treated units, 45 control units, and 30 time periods. The treatment kicks at Period 21 for all treated units, hence a block DID setup.

```{r head-sim_gsynth}
head(sim_gsynth)
```

Before we conduct any statistical analysis, it is helpful to visualize the data structure and spot missing values (if there are any). We can easily do so with the help of the **panelView** package. The figure below shows that: (1) there are 5 treated units and 45 control units; (2) the treated units start to be treated in period 21; and (3) there are no missing values, which is a rare case.

```{r sim-panelview-status, cache = FALSE, fig.height=7, fig.width=7, warning=FALSE}
library(panelView)
panelview(Y ~ D, data = sim_gsynth,  index = c("id","time"), pre.post = TRUE) 
```

The code chunk below visualizes the trends of outcome variable over the full panel across groups; different colors correspond to unique treatment statuses.

```{r sim-panelview-outcome, cache = FALSE,fig.height=5, fig.width=7}
panelview(Y ~ D, data = sim_gsynth,  index = c("id","time"), type = "outcome") 
```

### Estimation

We estimate the model using only the outcome variable ($Y$), treatment indicator ($D$), covariates ($X_1$) and ($X_2$), and group indicators ($id$) and ($time$).

To implement the Gsynth algorithm proposed in Xu (2017), set `method = "gsynth"`.

```{r sim2_onecore, cache = TRUE}
system.time(
out <- fect(Y ~ D + X1 + X2, data = sim_gsynth, index = c("id","time"), 
            method = "gsynth", force = "two-way", CV = TRUE, r = c(0, 5), 
            se = TRUE, nboots = 1000, vartype = 'parametric', 
            parallel = TRUE, cores = 16))
```

The first variable on the right-hand side is the binary treatment indicator, with remaining variables as controls. The `index` option designates unit and time indicators for fixed effects analysis.

The `force` option (`"none"`, `"unit"`, `"time"`, and `"two-way"`) specifies the additive component(s) of the fixed effects included in the model. The default option is `"two-way"` (including additive unit fixed effects). A cross-validation procedure is provided (when `CV = TRUE`) to select the number of unobserved factors within the interval of `r=c(0,5)`. When cross-validation is switched off, the first element in r will be set as the number of factors. Alternatively, for example, when `CV = FALSE`, r can take in a numeric value ($r = 0$).

Setting `se = TRUE`, the algorithm can produce uncertainty measurements. When the number of treated units is small, a parametric bootstrap procedure is preferred (`vartype = "parametric"`). The default number of parametric bootstrap runs is set to 200. The number of bootstrap runs can be set by `nboots`.

The algorithm prints out the results automatically. `sigma2` stands for the estimated variance of the error term; `IC` represents the Bayesian Information Criterion; and `MPSE` is the Mean Squared Prediction Error. The cross-validation procedure selects an `r*` that minimizes the MSPE.

Users can also use the `print` function to directly retrieve specific results. Here, we show three examples for the commonly requested estimates.

-   `est.att` reports the average treatment effect on the treated (ATT) by period
-   `est.avg` shows the ATT averaged over all periods
-   `beta` represents the coefficients of the time-varying covariates

```{r print-results, eval = FALSE}
print(out)
out$est.att
out$est.avg
out$beta
```

Treatment effect estimates from each bootstrap run are stored in `att.boot`, an array whose dimension = (#time periods \* #treated \* #bootstrap runs).

::: {.callout-note appearance="simple"}
**Parallel computing** will speed up both cross-validation and uncertainty estimation significantly. We recommend that users manually set the number of cores using the `cores` option. If this is not supplied or is `NULL`, we will automatically select the smaller of `8` and the number of usable system cores minus `2`, to prevent excessive use of system resources.
:::

```{r sim2, cache = TRUE, warning = FALSE}
system.time(
out <- fect(Y ~ D + X1 + X2, data = sim_gsynth, index = c("id","time"), method = "gsynth", force = "two-way", CV = TRUE, r = c(0, 5), se = TRUE, nboots = 1000, vartype = 'parametric', parallel = TRUE, cores = 16)
)
```

Gsynth in **fect** also incorporates jackknife method for uncertainty estimates.

```{r simJack, cache = TRUE, message = FALSE}
out2 <- fect(Y ~ D + X1 + X2, data = sim_gsynth,  index = c("id","time"),
               method = "gsynth", force = "two-way",
               CV = TRUE, r = c(0, 5), se = TRUE,
               vartype = "jackknife",
               parallel = TRUE, cores = 16)

```

### Cross-validation

Not all `cv.method` options are available in the Synth/Gsynth settings:

| | `fect_cv` (DID/TWFE) | CV in Gsynth | `fect_mspe` (post-hoc) |
|:--|:--|:--|:--|
| `"all_units"` | Default | Available | Default |
| `"treated_units"` | Available | Default | Available |
| `"loo"` | --- | Available | --- |

The `"loo"` method is exclusive to Gsynth (`time.component.from = "nevertreated"`) because it requires factor estimation to be independent of the held-out treated observations. In the `"nevertreated"` setting, factors are estimated from control units only, so holding out treated pre-treatment periods does not compromise estimation. In the `"notyettreated"` setting, all units contribute to estimation, making LOO less clean; use `"treated_units"` instead.

### Visualizing Results

By default, the `print` function produces a *gap* plot, equivalent to using `plot(out, type = "gap")`, and visualizes the estimated Average Treatment Effect on the Treated (ATT) by period. For reference, the true effects in `gsynth` range from 1 to 10 (with some added white noise) and take effect during periods 21 to 30.

Note that the plot() generates a **ggplot2** object, which users can save by simply calling `print` with the object's name.

```{r sim_gap1, fig.height=5, fig.width=7}
a <- plot(out) # by default, type = "gap"
print(a)
```

Setting `theme.bw = FALSE` sets the plotting style to the default *ggplot2* theme.

```{r sim_gap1a, fig.height=5, fig.width=7}
plot(out, theme.bw = FALSE) 
```

By switching on `connected`, the confidence interval of the ATT estimates will be represented by a shaded area.

```{r sim-gap-connected, fig.height=5, fig.width=7}
plot(out, connected = TRUE)
```

Moreover, by switching off `show.points`, ATT estimates can be displayed as a line plot.

```{r sim-gap-line, fig.height=5, fig.width=7}
plot(out, connected = TRUE, show.points = FALSE)
```

Aesthetic options for ggplot2 objects are compatible for all **fect** plots.

For demonstration, we use `main`, `xlim`, and `ylim` to set the plot title and axis labels.

```{r sim_gap2, fig.height=5, fig.width=7}
plot(out, type = "gap", ylim = c(-6,12), xlab = "Period", 
     main = "Estimated ATT (Gsynth)")
```

**fect** objects can generate eight types of plots for diverse demonstration purposes. The `type` option includes the following:

-   `"gap"`: Reports ATT by period (default).
-   `"counterfactual"`: Generates the observed treated outcome alongside the imputed counterfactual averages. The style can be modified with `raw = "all"` for linear or `raw = "band"` for graphic representation of all estimated counterfactual results.
-   `"factors"` and `"loadings"`: Plot the estimated factors and loadings, respectively.
-   `"status"`: Plots treatment status of all observations, similar to `panelview()`. The time labels are displayed only when `axis.lab = "time"` is set.
-   `"box"`: Displays the estimated individual treatment effects with box plot.
-   `"calendar"`: Shows how the treatment effect evolves over time.
-   `"equiv"`: Exhibits the average pre-treatment residuals with equivalence confidence intervals.

To visualize the estimated counterfactual outcomes, we use `type = "counterfactual"`. If no argument is assigned to `raw`, the plot will display the average of the *treated* and of the *estimated counterfactual* outcomes. When `method = "gsynth"` and `vartype = "parametric"`, the 95% confidence intervals around the counterfactual outcomes will be the same as for the event-study ATT (though flipped in shape).

```{r sim-counterfactual, cache = FALSE, fig.height=5, fig.width=7}
plot(out, type = "counterfactual")
```

**fect** offers two options for displaying all estimated counterfactual estimates: `raw = "all"` for linear and `raw = "band"` for a graphical (shaded band) representation.

```{r sim-counterfactual-all, cache = FALSE,fig.height=5, fig.width=7}
plot(out, type = "counterfactual", raw = "all")
```

Note that in the plot below, the shaded areas represent the 5-95% quantiles of the treated and estimated counterfactual trajectories, not their uncertainty estimates.

```{r sim-counterfactual-band, cache = FALSE,fig.height=5, fig.width=7}
plot(out, type = "counterfactual", raw = "band")
```


The figure below illustrates the treatment status of all observations. Users can specify the exact values shown on the x- and y-axes using the `xticklabels` and `yticklabels` parameters (i.e., `xticklabels=c("1","5","10", "15","20", "25", "30")`). Setting a nonexistent value to the option removes all numeric labels for that axis (i.e., `yticklabels="0"`). The `status` plot has now been incorporated into a standalone package, **panelView**.

```{r sim_status, cache = FALSE,fig.height=7, fig.width=7}
plot(out, type = "status", yticklabels="0", 
     xticklabels=c("5", "10", "15","20", "25", "30") )
```

The next two figures plot the estimated factors and factor loadings, respectively.

```{r sim_L, cache = TRUE, message = FALSE, results='hide', fig.height=7, fig.width=7}
plot(out, type = "loadings")
```

```{r sim_F, message = FALSE, results='hide', fig.height=5, fig.width=7}
plot(out, type = "factors", xlab = "Time")
```


The box plot visualizes the estimated individual treatment effects for the treated units. While these effects are not identified at the individual level, their dispersion provides insight into heterogeneous treatment effects across different time periods and informs model performance. By default, the number of total treated units is labeled on the graph. Note, if the number of treated units is small, the box plot will reduce to a scatter plot, as shown below.

We also provide the `xangle` and `yangle` options to allow users to tilt the labels for better display. We will demonstrate the functionality in later examples.

```{r sim_box, cache = FALSE,fig.height=5, fig.width=8}
plot(out, type = "box", xlab = "time",
     xticklabels=c("-19", "-15", "-10", "-5","0","5","10") )
```

If we want to focus on specific periods within the full panel, the `xlim` option is useful. Here, we narrow the time frame to fifteen periods before and ten periods after the treatment.

```{r sim_box2, eval = FALSE, fig.height=7, fig.width=7}
plot(out, type = "box", xlim = c(-15, 10), 
     xticklabels=c( "-15", "-10", "-5","0","5","10"))
```

To explore how the treatment effect evolves over time, we can set `type = "calendar"`.

In the plot below, the points represent the ATTs by calendar time. The blue curve shows a lowess fit of these estimates, with the shaded band indicating the 95% confidence interval. The red horizontal dashed line marks the overall average ATT across all time periods.

```{r calendar, cache = FALSE,fig.height=5, fig.width=7}
plot(out,type = "calendar")
```

Lastly, `type = "equiv"` helps visualize the average pre-treatment residuals with equivalence confidence intervals.

```{r sim-equiv, cache = FALSE, fig.height=5, fig.width=7}
plot(out, type = "equiv", ylim = c(-5, 5))
```

The floating legend displaying the F-test p-value and Equivalence test p-value can be removed by setting `show.stats = FALSE`.

```{r sim-equiv-no-stats, cache = FALSE, fig.height=5, fig.width=7}
plot(out, type = "equiv", show.stats =  FALSE)
```

Alternatively, it can be repositioned by providing coordinates to `stats.pos`.

```{r sim-equiv-reposition, cache = FALSE, fig.height=5, fig.width=7}
plot(out, type = "equiv", stats.pos = c(-19, 4.5), ylim = c(-5, 5))
```


------------------------------------------------------------------------

## Staggered DID

The second example examines the impact of Election-Day Registration (EDR) reforms on voter turnout in the United States, with the treatment taking effect at varying times across states. Further details about the `turnout` dataset can be found in @Xu2017 Section 5.

### Data structure

First, we use **panelView** to visualize the data structure. The following figure shows that (1) we have a balanced panel with 9 treated units and (2) the treatment starts at different time periods.

```{r turnout-panelview-status, cache = FALSE, warning=FALSE, fig.height=10, fig.width=7}
panelview(turnout ~ policy_edr, data = turnout, 
          index = c("abb","year"), pre.post = TRUE, 
          by.timing = TRUE) 
```

**panelView** can visualize the outcome variable by group, using colored lines to represent changes in treatment status.

```{r turnout-panelview-outcome, cache = FALSE, warning =FALSE, fig.height=5, fig.width=7}
panelview(turnout ~ policy_edr, data = turnout,
          index = c("abb","year"), type = "outcome", 
          main = "EDR Reform and Turnout", 
          by.group = TRUE)
```

### Estimation w/o factors

When no factor is assumed, the ATT estimates are close to what we obtain from difference-in-differences.

```{r turnout_did, cache = TRUE}
out0 <- fect(turnout ~ policy_edr + policy_mail_in + policy_motor, 
               data = turnout, index = c("abb","year"), 
               se = TRUE, method = "gsynth",
               r = 0, CV = FALSE, force = "two-way", 
               nboots = 1000, seed = 02139)
```

```{r turnout_did_gap, fig.height=5, fig.width=7}
plot(out0, type = "gap", xlim = c(-15, 5), ylim=c(-15, 10))
```

### Estimation w/ factors

We now allow the algorithm to determine the optimal number of factors that best predict the pre-treatment data with cross-validation. The `keep.sims` option must be set to `TRUE` if the user wishes to visualize subsets of the treated units in the "counterfactual" plot.

```{r turnout_est, cache = TRUE}
out_turnout <- fect(turnout ~ policy_edr + policy_mail_in + policy_motor, 
              data = turnout,  index = c("abb","year"), 
              se = TRUE, method = "gsynth", vartype = "parametric",
              r = c(0, 5), CV = TRUE, force = "two-way", 
              nboots = 1000, seed = 02139, keep.sims = TRUE)
```

### Implied weights

`out$wgt.implied` ($N_{co}\times N_{tr}$) stores the implied weights of all control units for each treated unit. Different from the synthetic control method, the weights can take both positive and negative values. Below we show the control unit weights for Wisconsin.

```{r turnout-implied-weights}
dim(out_turnout$wgt.implied)
sort(out_turnout$wgt.implied[,8])
```

### Visualizing results

Like we have demonstrated with the multiperiod DID analysis, we also use the gap (default) plot to visualize the ATT.

```{r turnout_gap, fig.height=5, fig.width=7}
plot(out_turnout, xlim = c(-10, 5), ylim=c(-15, 10))
```

For staggered DID, a status plot can be generated after estimation. Here, we use `xlab`, `ylab`, and `main` to modify the axis labels and the graph title, respectively. Additionally, `xangle` and `yangle` can tilt the numeric labels to improve readability. Note that the values provided for these options specify the degree of tilt.

```{r turnout-status-plot, fig.height=12, fig.width=7}
plot(out_turnout, type = "status",xlab = "Year", ylab = "State", main = "Treatment Status", 
     xticklabels=c(1920, 1928, 1936, 1944, 1952, 1960, 
                   1968, 1976, 1984, 1992, 2000, 2008), xangle=10)
```

We can also visualize this using the "counterfactual" plot.

```{r turnout_counterfactual, fig.height=5, fig.width=7}
plot(out_turnout, type = "counterfactual")
```

If we want to zoom in to examine the treatment effect on a specific state, such as Wisconsin, we can do so by setting the `id` option accordingly. The `id` option can also accept a vector if one wishes to plot multiple treated units. This plot will only display confidence intervals for the counterfactual estimates if `keep.sims = TRUE` is set in the estimation step and `vartype = "parametric"`.

```{r turnout_gap2, fig.height=5, fig.width=7}
plot(out_turnout, type = "counterfactual", id = "WI", main = "Wisconsin")
```

As we have seen in the last example, due to the small number of treated units, the box plot reduces to a scatter plot.

```{r turnout_box, fig.height=4, fig.width=8}
plot(out_turnout, type = "box", 
     xticklabels=c("-20", "-15", "-10", "-5","0","5","10"))
```

Here, we visualize the estimated average treatment effects by calendar time.

```{r turnout_calendar, fig.height=5, fig.width=7}
plot(out_turnout, type = "calendar", ylim = c(-15,15))
```

Estimated factors and factor loadings are plotted below.

```{r turnout_F, message = FALSE, results = 'hide', fig.height=5, fig.width=7, warning=FALSE}
plot(out_turnout, type = "factors", xlab = "Year")
```

```{r turnout_L, message = FALSE, results = 'hide', fig.height=7, fig.width=7, warning=FALSE}
plot(out_turnout, type = "loadings")
```

------------------------------------------------------------------------

## Unbalanced panels

Gsynth in **fect** can accommodate unbalanced panels. To illustrate how it works, we randomly remove 50 observations as well as the first 15 observations of Wyoming from the turnout dataset and then re-estimate the model.

```{r create-unbalanced-data}
set.seed(123456)
turnout.ub <- turnout[-c(which(turnout$abb=="WY")[1:15], 
                         sample(1:nrow(turnout),50,replace=FALSE)),]
```

Again, before running any regressions, we first plot the data structure and visualize missing values. In the following plot, white cells represent missing values.

```{r turnout_ub_panelview_miss, cache = FALSE,fig.height=7, fig.width=7}
panelview(turnout ~ policy_edr + policy_mail_in + policy_motor, 
          data = turnout.ub,  index = c("abb","year"), 
          pre.post = TRUE) 
```

### Estimation

```{r turnout_ub_est, cache = TRUE, message = FALSE}
out_ub <- fect(turnout ~ policy_edr + policy_mail_in + policy_motor, 
              data = turnout.ub,  index = c("abb","year"), 
              se = TRUE, method = "gsynth", 
              r = c(0, 5), CV = TRUE, force = "two-way", 
              parallel = TRUE, cores = 16, min.T0 = 8,
              nboots = 1000, seed = 02139)
```

### Visualizing results

The status plot visualizes the regression data and differs from the **panelView** plot. Wyoming's data is shown in gray, as its observations are excluded due to an insufficient number of pre-treatment periods.

The missing data matrix can also be accessed via `out_ub$obs.missing`, where values 0, 1, 2, 3, and 4 represent missing values, control observations, treated (pre-treatment) observations, treated (post-treatment) observations, and treated observations removed from the sample, respectively.

```{r, turnout_ub_panelview_miss2, fig.height=12, fig.width=7}
plot(out_ub, type = "status",
     xticklabels=c(1920, 1928, 1936, 1944, 1952, 1960, 
                   1968, 1976, 1984, 1992, 2000, 2008),
     xangle=10)
```

Like in **panelView**, we can plot specific observations by setting `id` and adjust labels on x-axis as needed.

```{r turnout_ub_obs_2, fig.height=7, fig.width=7}
plot(out_ub, type = "status", xlab = "Year", ylab = "State",
     main = "Treatment Status", id = out_ub$id[out_ub$tr],
     xlim = c(1920,2012), 
     xticklabels=c(1920, 1928, 1936, 1944, 1952, 1960,
                   1968, 1976, 1984, 1992, 2000, 2008))

```

We re-produce the gap plot with the unbalanced panel, here, we set the range of y in between $(-10,20)$.

```{r turnout_ub_gap, fig.height=5, fig.width=7}
plot(out_ub, type = "gap", ylim = c(-10, 20))
```


------------------------------------------------------------------------

## CFE with GSynth

Gsynth is not limited to the IFE method. The complex fixed effects (CFE) estimator can also be used with `time.component.from = "nevertreated"`. This is useful when the data generating process involves not only latent interactive fixed effects but also additional additive fixed effects (e.g., region, industry), unit-specific time trends (via `Q.type`), or time-invariant covariates with time-varying coefficients (`Z`/`gamma`). See [Chapter @sec-cfe] for the full CFE tutorial.

### Example: CFE + nevertreated

```{r cfe_nt_demo, eval=TRUE, cache=TRUE, message=FALSE, results='hide'}
out.cfe.nt <- fect(Y ~ D + X1 + X2, data = sim_gsynth, index = c("id","time"),
                   method = "cfe", force = "two-way",
                   time.component.from = "nevertreated",
                   Q.type = "linear",
                   se = FALSE, CV = TRUE, r = c(0, 5),
                   parallel = TRUE, cores = 16)
```

```{r cfe-nt-summary}
cat("CFE + nevertreated: r.cv =", out.cfe.nt$r.cv,
    ", ATT =", round(out.cfe.nt$att.avg, 3), "\n")
```

::: {.callout-note appearance="simple"}
Under `time.component.from = "nevertreated"`, the CFE estimator uses only never-treated controls for factor estimation — the same principle as gsynth. The additional CFE components (extra FE, time trends, Z/gamma) are estimated on the control panel and projected onto treated units alongside the interactive factors. See [Chapter @sec-cfe] for the full CFE tutorial.
:::

### Comparing gsynth and CFE + nevertreated

Both approaches share the same estimation regime (never-treated controls only), but CFE can add structural components on top. Whether the extra flexibility helps depends on the data generating process. We compare three specifications using `fect_mspe()`:

```{r cfe_nt_vs_gsynth, eval=TRUE, cache=TRUE, message=FALSE, results='hide'}
# Model 1: gsynth (pure IFE, r = 2)
out.gsynth.comp <- fect(Y ~ D + X1 + X2, data = sim_gsynth, index = c("id","time"),
                        method = "gsynth", force = "two-way",
                        r = 2, se = FALSE, CV = FALSE)

# Model 2: CFE + nevertreated with r = 2 only (equivalent to gsynth)
out.cfe.nt.comp <- fect(Y ~ D + X1 + X2, data = sim_gsynth, index = c("id","time"),
                        method = "cfe", force = "two-way",
                        time.component.from = "nevertreated",
                        r = 2, se = FALSE, CV = FALSE)

# Model 3: CFE + nevertreated with r = 2 and linear trend (overspecified)
out.cfe.nt.lin <- fect(Y ~ D + X1 + X2, data = sim_gsynth, index = c("id","time"),
                       method = "cfe", force = "two-way",
                       time.component.from = "nevertreated",
                       Q.type = "linear", r = 2, se = FALSE, CV = FALSE)
```

```{r cfe_nt_mspe, eval=TRUE, cache=TRUE}
mspe.comp <- fect_mspe(list(gsynth_r2 = out.gsynth.comp,
                            CFE_r2 = out.cfe.nt.comp,
                            CFE_linear_r2 = out.cfe.nt.lin), seed = 1234)
print(mspe.comp$summary[, c("Model", "MSPE", "RMSE", "MAD")])
```

Since `sim_gsynth` follows a pure IFE data generating process with two factors, Models 1 and 2 should produce identical MSPE --- confirming that CFE with `time.component.from = "nevertreated"` and no additional structure is numerically equivalent to gsynth. Model 3, which adds unnecessary linear trends, should produce similar or slightly worse MSPE because the extra parameters add noise without benefit when the true DGP has no unit-specific trends.


## Additional Notes

1.  **Unbalanced Panels**: Running Gsynth within **fect** on unbalanced panels will take significantly more time compared to balanced panels, often by a factor of 100:1 or more. This is because the EM algorithm, which fills in missing values (implemented in C++), requires many more iterations to converge. To reduce run-time, users can remove units or time periods with extensive missing values. Understanding the data structure before running any regressions is always helpful. Note that this approach differs from setting `method = "ife"`, as no pre-treatment data from the treated units is used to impute $\hat{Y}(0)$.

2.  **Adding Covariates**: Including covariates in the model will significantly slow down the algorithm, as the IFE/MC model requires more time to converge. Users should be aware of this trade-off when incorporating covariates.

3.  **Setting `min.T0`**: Setting `min.T0` to a positive value helps. The algorithm will automatically exclude treated units with too few pre-treatment periods. A larger $T_0$ reduces bias in causal estimates and minimizes the risk of severe extrapolation. When running cross-validation to select the number of factors, `min.T0` must be equal to or greater than (`r.max` + 2). Errors frequently occur when there are too few pre-treatment periods, so ensuring adequate $T_0$ (e.g. setting `min.T0 = 5`) is crucial.


## How to Cite

If you find this method helpful, you can cite @Xu2017.

```bibtex
@article{Xu2017,
  title = {Generalized Synthetic Control Method: Causal Inference with Interactive Fixed Effects Models},
  author = {Xu, Yiqing},
  journal = {Political Analysis},
  volume = {25},
  number = {1},
  pages = {57--76},
  year = {2017}
}
```
