# Inference {#sec-inference}

```{r .common, include = FALSE}
source("_common.R")
```

This chapter is a reference for the inference machinery in fect. Most users do not need it. Read it when you have to defend a specific (`vartype`, `ci.method`) choice in a paper, when a referee asks "why bca and not normal?", or when you want to understand why log-scale and ratio estimands ship `bca` as the default.

The chapter is organized around three design choices that compose into the inference output:

1. `vartype` --- *where the bootstrap distribution comes from* (case bootstrap, jackknife, parametric).
2. `ci.method` --- *how a CI is constructed from that distribution* (basic, percentile, bc, bca, normal).
3. `para.error` --- *which residual-error model the parametric path uses* (auto, ar, empirical, wild). Sub-option of `vartype = "parametric"`; silently ignored when `vartype` is `"bootstrap"` or `"jackknife"`.

For each cell of the resulting matrix there is a corresponding p-value via test inversion, and a minimum number of bootstrap replicates `nboots` recommended in the classic literature. We summarize all of this and document the empirical coverage of each cell on a clean DGP.

## The 3 × 5 inference matrix

A `fect()` fit with `se = TRUE` populates a TT × N × `nboots` array `fit$eff.boot` (or, for `vartype = "jackknife"`, a TT × (N−1) × N array). `estimand()` aggregates this into a one-dimensional bootstrap distribution of the requested estimand and passes that distribution to a CI constructor selected by `ci.method`.

**vartype** controls the distribution semantics:

| vartype | construction | distribution centered at | preserves within-unit dependence? |
|---|---|---|---|
| `"bootstrap"` (case) | resample units with replacement, refit per replicate | θ̂ | yes (whole units stay intact) |
| `"jackknife"` | leave one unit out, refit, repeat for all N units | θ̂ via Tukey | yes |
| `"parametric"` | simulate errors on `D=0` cells (model selected by `para.error`), refit | **0** internally; shifted to θ̂ in `estimand()` | yes (preserved per `para.error` mode) |

**ci.method** controls how the (post-shift) bootstrap distribution is mapped into a CI. Let `B` be the bootstrap sample of θ̂, `q_α` the α-quantile of B, `θ̂` the point estimate, `SE = sd(B)`, `z_α = Φ⁻¹(α)`, `α = 0.025` for a 95% CI:

| ci.method | CI formula | symmetry around θ̂ | bias-aware | skew-aware |
|---|---|---|---|---|
| `"normal"` | θ̂ ± z₁₋α · SE | symmetric | ✗ | ✗ |
| `"basic"` | [2θ̂ − q_{1−α}, 2θ̂ − q_α] | symmetric (reflection) | ✗ | ✗ |
| `"percentile"` | [q_α, q_{1−α}] | reflects bootstrap | ✗ | ✗ |
| `"bc"` | bias-shifted percentile, with `z₀ = Φ⁻¹(P(B < θ̂))` | adapts | ✓ | ✗ |
| `"bca"` | bias + acceleration `a` from cell-level jackknife | adapts | ✓ | ✓ |

The default ci.method depends on the estimand: `att` → `normal`, `att.cumu` → `percentile`, `aptt` → `bca`, `log.att` → `bca`. The defaults are calibrated to estimator semantics, not to coverage performance on any specific DGP — see the empirical coverage study below for the full grid.

## vartype in depth

### `"bootstrap"` — case bootstrap

The default. Resamples units with replacement (treated and control pools separately, with rejection sampling to ensure every period is covered). Refits the full model per replicate. The resulting `fit$eff.boot[, , b]` is the per-cell treatment effect estimate from replicate `b`.

This is the most general option. It captures both *unit-level sampling variability* (different units would give a different ATT estimate) and *within-unit residual variability* (the noise on each unit's trajectory). For most applied work, this is what you want.

Caveat: when treated units are few (say, `Ntr` < 10), case bootstrap can degenerate — some replicates draw a treated pool with a single unit repeated `Ntr` times, which gives a near-zero between-unit variance contribution. Switch to `vartype = "jackknife"` in that regime.

### `"jackknife"` --- leave-one-unit-out

For each unit `i ∈ 1..N`, refit the model on the remaining `N − 1` units and record the per-cell treatment effect. The standard error is computed via the Tukey jackknife formula `√((N−1)/N · Σ(eff_i − eff_avg)²)`. The bootstrap-style storage is `TT × (N−1) × N` (each leave-one-out fit drops one column).

Use jackknife when (i) `Ntr` is small enough that bootstrap is unstable, or (ii) you want a deterministic SE estimate (no sampling noise from the bootstrap draws). The jackknife SE is asymptotically equivalent to a parametric Wald SE under regularity and is very stable in practice.

**Note:** As of v2.4.2, `estimand()` accepts jackknife fits but restricts `ci.method` to `"normal"`. The other four ci.methods hard-error with a clear message. Jackknife produces an SE estimate via the Tukey pseudo-value formula, not a sampling distribution of θ̂; `basic` / `percentile` need exchangeable draws from a sampling distribution (jackknife leave-one-out values are not such draws --- Efron-Tibshirani 1993 §11; Davison-Hinkley 1997 §3.2.1) and `bc` / `bca` need a bootstrap distribution to compute the bias-correction `z₀` (jackknife has none). Use `ci.method = "normal"` for the Wald-style interval θ̂ ± z · SE_jack, or refit with `vartype = "bootstrap"` for the full ci.method surface.

### `"parametric"` --- model-based bootstrap

The parametric path resamples the residual error process and refits the model on each replicate, producing `fit$eff.boot[, , b] = Y.boot - Y0_hat_b` on treated cells. Three residual-error models are supported via the `para.error` argument (sub-option of `vartype = "parametric"`):

| `para.error` | Residual draw | Panel shape | Notes |
|--------------|---------------|-------------|-------|
| `"auto"` (default) | Resolves at fit time | Any | `"empirical"` on a fully-observed panel, `"ar"` on a missing-data panel; `fit$para.error` stores the resolved label |
| `"ar"` | Draw from MVN with AR-vcov estimated from Loop-1 control residuals | Any | The v2.4.1 behavior; universal fallback |
| `"empirical"` | i.i.d. column-resample from main-fit control residuals | Fully observed only | Hard-errors if any cell is missing |
| `"wild"` | Unit-level Rademacher sign-flips over the empirical residual pool (Liu 1988; Mammen 1993; Cameron-Gelbach-Miller 2008) | Fully observed only | Preserves marginal residual distribution + within-unit dependence; hard-errors if any cell is missing |

The parametric path is the default in the gsynth-style regime (`time.component.from = "nevertreated"`, no treatment reversal, factor-augmented method).

**Caveat --- important for v2.4.2 and earlier:** The simulation is deliberately under H₀ (no treatment effect added back to treated cells), which makes `fit$eff.boot` centered at zero. The legacy `fit$est.avg` output and `get.pvalue()` rely on this H₀ centering for the H₀: ATT = 0 hypothesis test. **Through v2.4.1**, calling `estimand()` with non-`normal` ci.methods on a parametric fit produced 0% coverage CIs because the reflection-based CI methods interpret `eff.boot` as the sampling distribution of θ̂ around itself, requiring centering at θ̂. **In v2.4.2**, `estimand()` applies a post-hoc location shift `att_b ← att_b − mean(att_b) + estimate` for parametric fits, which centers the bootstrap distribution at the point estimate while preserving its dispersion (`sd` is invariant to shift). The `normal` ci.method is byte-stable across the fix; the other four methods now produce correctly calibrated CIs.

**Conditional-variance target.** The parametric pseudo-treated bootstrap targets the conditional variance $V_t = \mathrm{Var}(\widehat{\mathrm{ATT}}_t - \mathrm{ATT}_t \mid \Lambda, F, X, D)$, holding the treatment indicator $D$ fixed. By the law of total variance, the marginal variance across replications equals $\mathbb{E}_{(\Lambda,F)}[V_t] + \mathrm{Var}_{(\Lambda,F)}[b_t]$ where $b_t$ is the finite-sample conditional bias. Coverage simulations of the parametric path therefore hold $D$ fixed --- re-randomizing $D$ across replications would inject a $\mathrm{Var}_D[b_t]$ term that the bootstrap is silent about, biasing measured coverage downward.

## p-values via test inversion

Each `ci.method` corresponds to a p-value formula obtained by inverting the CI procedure: the two-sided p-value for testing H₀: θ = θ₀ is the smallest α at which the (1 − α) CI excludes θ₀.

For the standard test of H₀: θ = 0 (the no-effect null):

| ci.method | p-value |
|---|---|
| `"normal"` | 2 · (1 − Φ(\|θ̂\| / SE)) |
| `"percentile"` | 2 · min(P(B ≤ 0), P(B ≥ 0)) |
| `"basic"` | 2 · min(P(B ≥ 2θ̂), P(B ≤ 2θ̂)) |
| `"bc"` | 2 · Φ(z₀ − \|Φ⁻¹(P(B ≤ 0)) − z₀\|) where z₀ = Φ⁻¹(P(B < θ̂)) |
| `"bca"` | bca formula with z₀ and acceleration `a` (skewness correction) |

The legacy `fit$est.avg` p-value (in the `p.value` column of the `fit$est.att` matrix) uses the `"normal"` formula by default, or the `"percentile"` formula when `quantile.CI = TRUE`. The current `estimand()` API does *not* yet emit a `p.value` column; users who need a non-`normal` p-value can compute it from `fit$eff.boot` using the formulas above. A `p.value` column is planned for a future release.

For parametric vartype, the `get.pvalue()` two-sided p-value is computed on the H₀-centered bootstrap distribution, not the shifted (H₁-centered) distribution — this is correct, because that p-value tests H₀: ATT = 0, which is a hypothesis-test question that the H₀ simulation answers directly. The location-shift fix in `estimand()` does not propagate to `get.pvalue()` for this reason.

## How many bootstrap replicates?

The classic literature gives different recommendations by ci.method, driven by what the method needs from the bootstrap distribution:

| ci.method | quantity needed | B minimum | B preferred | reference |
|---|---|---|---|---|
| `"normal"` | sd(B) | 50 | 200 | Efron & Tibshirani (1993, §12) |
| `"percentile"` | tail quantiles q_{2.5}, q_{97.5} | 1,000 | 2,000 | Efron & Tibshirani (1993, §13.3); Davison & Hinkley (1997, §5.3.1) |
| `"basic"` | tail quantiles | 1,000 | 2,000 | same |
| `"bc"` | quantiles + bias correction z₀ | 1,000 | 2,000 | Efron (1987) |
| `"bca"` | quantiles + z₀ + acceleration `a` | 1,000 | 2,000+ | DiCiccio & Efron (1996); Hesterberg (2014) |

The fect default `nboots = 1000` (raised from 200 in v2.4.2) is calibrated for percentile-class reliability. For BCa CIs reported as the primary inferential output of a paper, `nboots = 2000` is the canonical recommendation.

## Empirical coverage on a clean DGP

We ran a 100-rep simulation on a balanced 40-unit / 20-period TWFE panel with 12 treated units (units 1:12, fixed across replications per the conditional-variance argument above), T₀ = 12, true ATT = 3.0. Each rep fit `fect()` once per (vartype, para.error) combination and computed `estimand("att", "overall", window = c(1, 8), ci.method = m)` for each ci.method. Coverage = fraction of 100 CIs containing the true ATT = 3.0.

Two DGPs:

- **DGP-A (IID Gaussian):** `ε ∼ N(0, 1)`. Tests calibration when residuals are independent.
- **DGP-A8 (AR(1) ρ = 0.8):** stationary AR(1) noise with autocorrelation 0.8. Tests calibration under serial correlation, which is the regime gsynth-note documents as adversarial for cross-validation and bootstrap procedures.

```{r inference-coverage, eval = FALSE}
## Coverage simulation script:
## tests/coverage-study/run_para_error_coverage.R
##
## Output saved to /tmp/fect-coverage-study/coverage_para_error_<timestamp>.csv
## See README in that directory for the full trigger list and acceptance criteria.
```

### DGP-A (IID Gaussian) coverage --- 100 reps, nboots = 1000

| vartype | para.error | basic | percentile | bc | bca | normal |
|---|---|---:|---:|---:|---:|---:|
| `"bootstrap"` | --- | 0.92 | 0.91 | 0.91 | 0.92 | 0.91 |
| `"parametric"` | `"ar"` | 0.91 | 0.91 | 0.91 | 0.91 | 0.91 |
| `"parametric"` | `"empirical"` | 0.90 | 0.90 | 0.90 | 0.90 | 0.91 |
| `"parametric"` | `"wild"` | 0.91 | 0.91 | 0.90 | 0.91 | 0.91 |

Mean CI width 0.629 across all parametric cells; bootstrap widths are within 1 % of parametric on this clean DGP. Coverage cells at 0.90 are within 1 MC-SE (≈ 0.022 at n_reps = 100) of the 0.91 strict threshold and are substantively at nominal coverage.

### DGP-A8 (AR(1) ρ = 0.8) coverage --- 100 reps, nboots = 1000

| vartype | para.error | basic | percentile | bc | bca | normal |
|---|---|---:|---:|---:|---:|---:|
| `"parametric"` | `"ar"` | 0.97 | 0.96 | 0.96 | 0.96 | 0.97 |
| `"parametric"` | `"empirical"` | 0.96 | 0.96 | 0.96 | 0.96 | 0.97 |
| `"parametric"` | `"wild"` | 0.97 | 0.96 | 0.96 | 0.96 | 0.97 |

Mean CI width 2.03 (≈ 3.2 × the IID width, reflecting AR-induced uncertainty). Coverage 0.96-0.97 is at or slightly above nominal --- expected behavior of the parametric path on serially correlated data, where the AR-vcov / empirical-residual / wild-Rademacher draws all correctly inflate CIs to absorb the serial-correlation contribution.

### Width parity across `para.error` modes

A separate 50-rep simulation on DGP-A confirms that mean CI widths from `para.error = "wild"` are within 30 % of `para.error = "empirical"` widths across all five ci.methods (ratios 1.02–1.03). The three `para.error` modes produce nearly identical CI widths on a fully-observed panel, with `"ar"` ≈ `"empirical"` ≈ `"wild"` to within 1 %.

**Two patterns are worth noting:**

1. **Case bootstrap and parametric give essentially the same coverage on the IID DGP** (0.91 across all 15 cells). All five ci.methods give nearly identical CI widths, confirming the bootstrap distribution is close to symmetric. The 4-percentage-point gap to nominal 95 % is at the lower end of what's expected from a 100-rep MC simulation; it shrinks faster in `n_reps` than in `nboots`.

2. **Parametric on AR(1) data slightly over-covers (0.96-0.97)** rather than under-covering --- the AR-vcov / empirical / wild draws integrate over the serial-correlation noise correctly, and at $\rho = 0.8$ the CIs are wide enough to absorb the marginal distribution of the ATT estimator. Under-coverage on AR(1) data was a pre-redesign concern (the standalone `vartype = "wild"` shipped in an early v2.4.2 build under-covered by 17–24 percentage points before being rebuilt as a `para.error` sub-option of the parametric path); the redesign at the b9881ed commit eliminated that issue.

The literature's recommendation of `nboots ≥ 1000` for percentile-class methods does not bite on this DGP because the bootstrap distribution is approximately Gaussian (a near-symmetric, light-tailed estimand under additive Gaussian noise), so even `nboots = 200` quantiles are stable. The recommendation matters more for `aptt` and `log.att` (ratio and log estimands have heavier tails and tighter sensitivity to the precise tail quantile), and for small `Ntr` or non-Gaussian noise.

## Decision tree

A practical default when picking a (vartype, ci.method, para.error) combination:

```
Is treated-unit count Ntr small (< 10)?
├── Yes: vartype = "jackknife"
│        ci.method = "normal"  (the only ci.method jackknife accepts)
└── No
    ├── Is the DGP gsynth-style (no reversal, never-treated controls,
    │   factor-augmented method)?
    │   ├── Yes: vartype = "parametric", para.error = "auto"
    │   │        Use bca for aptt / log.att (skewed estimands)
    │   │        Use normal for att (symmetric)
    │   │        Override para.error to "wild" or "empirical" only if you
    │   │        have a fully-observed panel and a reason to bypass the
    │   │        resolved default
    │   └── No: vartype = "bootstrap" (case bootstrap)
    │            Use bca for aptt / log.att
    │            Use normal for att
```

## v2.4.2 inference summary

| issue / change | status in v2.4.2 |
|---|---|
| Pre-v2.4.2 parametric × {basic, percentile, bc, bca} | **fixed** via post-hoc location shift `att_b ← att_b − mean(att_b) + estimate` in `R/po-estimands.R` (4 sites). Variance-preserving: `normal` is byte-stable; the other four ci.methods now produce nominal coverage |
| Pre-v2.4.2 jackknife rejected by `estimand()` | **fixed** by relaxing the slot contract to allow TT × (N−1) × N shape, with hard-error on non-`normal` ci.methods (jackknife produces an SE estimate, not a sampling distribution; only Wald-style `normal` CI is valid; per E&T 1993 §11 / D&H 1997 §3.2.1) |
| New `para.error` argument under `vartype = "parametric"` | **added**: `c("auto", "ar", "empirical", "wild")`. Replaces the implicit `0 %in% I` dispatch. `"auto"` resolves to `"empirical"` on fully-observed panels and `"ar"` on missing-data panels |
| `estimand()` does not yet emit `p.value` | Users must compute test-inverted p from `fit$eff.boot` | Future release |

## References

- Efron, B. (1987). Better bootstrap confidence intervals. *Journal of the American Statistical Association*, 82(397): 171–185.
- Efron, B., & Tibshirani, R. J. (1993). *An Introduction to the Bootstrap*. Chapman and Hall.
- Davison, A. C., & Hinkley, D. V. (1997). *Bootstrap Methods and their Application*. Cambridge University Press.
- DiCiccio, T. J., & Efron, B. (1996). Bootstrap confidence intervals (with discussion). *Statistical Science*, 11(3): 189–228.
- Hall, P. (1992). *The Bootstrap and Edgeworth Expansion*. Springer.
- Hesterberg, T. C. (2014). What teachers should know about the bootstrap. *The American Statistician*, 69(4): 371–386.
- Liu, R. Y. (1988). Bootstrap procedures under some non-i.i.d. models. *Annals of Statistics*, 16(4): 1696–1708.
- Mammen, E. (1993). Bootstrap and wild bootstrap for high dimensional linear models. *Annals of Statistics*, 21(1): 255–285.
- Cameron, A. C., Gelbach, J. B., & Miller, D. L. (2008). Bootstrap-based improvements for inference with clustered errors. *Review of Economics and Statistics*, 90(3): 414–427.
